---
title: "20210224_DEG_heatmap"
author: "Gijsje"
date: "2/24/2021"
output: 
  rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---


```{r packages, echo=FALSE, message=FALSE, results='hide'}
library(readxl)
library(edgeR)
library(limma)
library(DESeq2)
library(tidyverse)
library(data.table)
library(shiny)
library(DT)
library(pheatmap)
library(ggfortify)
library(ggplot2)
library(UpSetR)
library(VennDiagram)
library(knitr) 
library(tidyverse)
library(dplyr)
library(tidyr)
library(tibble)
library(purrr)
library(corrplot)
library(ggplot2)
library(ggpubr)


knitr::opts_chunk$set( 
  warning=FALSE,
  message=FALSE,
  results = 'asis',
  error = FALSE,
  tidy = FALSE,
  fig.show = "hold")
```


# DEG across stimulations FDR % log FC > 1 or -1   
#1204 genes 

```{r load tables, echo = TRUE, results = 'hide'}
LPS <- read.csv("~/Documents/MiGASti/Databases/res_name_LPS_LFC1.txt", sep="")
IFNy <- read.csv("~/Documents/MiGASti/Databases/res_name_IFNy_LFC1.txt", sep="")
R848 <- read.csv("~/Documents/MiGASti/Databases/res_name_R848_LFC1.txt", sep="")
TNFa <- read.csv("~/Documents/MiGASti/Databases/res_name_TNFa_LFC1.txt", sep="")
DEX <- read.csv("~/Documents/MiGASti/Databases/res_name_DEX_LFC1.txt", sep="")
IL4 <- read.csv("~/Documents/MiGASti/Databases/res_name_IL4_LFC1.txt", sep="")
ATP <- read.csv("~/Documents/MiGASti/Databases/res_name_ATP_LFC1.txt", sep="")

genes_LPS <- LPS$ensembl
genes_ATP <- ATP$ensembl
genes_R848 <- R848$ensembl
genes_IFNy <- IFNy$ensembl
genes_DEX <- DEX$ensembl
genes_IL4 <- IL4$ensembl
genes_TNFa <- TNFa$ensembl
    
# create a vector with all genes listed
x <- c(genes_LPS, genes_ATP, genes_R848, genes_IFNy, genes_DEX, genes_IL4, genes_TNFa)
# make a dataframe
as.data.frame(x) #1325 genes 
# check if there are genes duplicate 
duplicated(x) 
unique(x)
#remove duplicated
ensembl <- x[!duplicated(x)] #1204 genes 

genelist <- as.data.frame(ensembl)
ensembl <- genelist$ensembl[!duplicated(genelist$ensembl)]
genelist <- as.data.frame(ensembl) 
genelist %>% distinct(ensembl, .keep_all = TRUE)

load("~/Documents/MiGASti/res_name_LPS2.Rdata")
load("~/Documents/MiGASti/docs/res_name_IFNy2.Rdata")
load("~/Documents/MiGASti/Databases/res_name_TNFa2.Rdata")
load("~/Documents/MiGASti/Databases/res_name_R8482.Rdata")
load("~/Documents/MiGASti/Databases/res_name_IL42.Rdata")
load("~/Documents/MiGASti/Databases/res_name_ATP2.Rdata")
load("~/Documents/MiGASti/Databases/res_name_DEX2.Rdata")

DEG_genes_LPS = merge(res_name_LPS2, genelist, by ="ensembl")
DEG_genes_IFNy = merge(res_name_IFNy2, genelist, by = "ensembl")
DEG_genes_IL4 = merge(res_name_IL42, genelist, by = "ensembl")
DEG_genes_R848 = merge(res_name_R8482, genelist, by = "ensembl")
DEG_genes_ATP = merge(res_name_ATP2, genelist, by = "ensembl")
DEG_genes_DEX = merge(res_name_DEX2, genelist, by = "ensembl")
DEG_genes_TNFa = merge(res_name_TNFa2, genelist, by = "ensembl")

Expression_stimulations <- cbind(DEG_genes_LPS, DEG_genes_IFNy, DEG_genes_IL4, DEG_genes_R848, DEG_genes_ATP, DEG_genes_DEX, DEG_genes_TNFa)

as.data.frame(Expression_stimulations) 
# changes names in data frame
names(Expression_stimulations)[2] <- "LPS"
names(Expression_stimulations)[10] <- "IFNy"
names(Expression_stimulations)[18] <- "IL4"
names(Expression_stimulations)[26] <- "R848"
names(Expression_stimulations)[34] <- "ATP"
names(Expression_stimulations)[42] <- "DEX"
names(Expression_stimulations)[50] <- "TNFa"

DF_log2FC <- Expression_stimulations %>% select(1, 2, 10, 18, 26, 34, 42, 50)
DF <- as.data.frame(DF_log2FC)
DF <- t(DF)
head( DF )[,1:10]

# fill NA with 0
DF_log2FC[is.na(DF_log2FC)] = 0
#save(DF_log2FC, file = "Kmeans1204gsupport.Rdata")

gencode <- read.delim("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
gencode$ensembl = gencode$gene_id
DF_log2FC <- merge(gencode, DF_log2FC, by = "ensembl")
DF_log2FC$symbol = DF_log2FC$GeneSymbol

# remove the first column
df_num = as.matrix(DF_log2FC[,4:10])
# scale the data 
df_num_scale = scale(df_num)
save(df_num, file = "df_num_logFC1stim.Rdata")
#save(DF_log2FC, file = "LOGFC1_stimulations.Rdata")
#save(df_num_scale, file = "df_num_scale_LOGFC1stim.Rdata")
```

### Overlap 1204 genes with Patir

```{r overlap, echo = TRUE}
gencode <- read.delim("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
gencode$ensembl = gencode$gene_id
genelist <- merge(gencode, genelist, by = "ensembl")
genelist$symbol = genelist$GeneSymbol
Patir <- read.csv("~/Documents/MiGASti/Databases/Patir.txt", sep="")
names(Patir)[1] <- "symbol"
overlap1 <- merge(Patir, genelist, by = "symbol")
list(overlap1$symbol)
```

### Heatmap DEG
#1204 genes (scaling data and additional row scaling)

```{r heatmap_GFM_cor, echo = TRUE}
rownames(DF_log2FC) = DF_log2FC$symbol
genelist <- c("ADORA3", "AIF1", "C2", "ALOX5AP", "ARHGAP25", "C2", "C3AR1", "CD300A", "CIITA", "EBI3", "FCGR1A", "FCGR1B", "FGD2", "FGL2", "FPR1", "GIMAP2", "GPR183", "HAMP", "IFI30", "IL13RA1", "IRF5", "LILRB2", "LY86", "LYL1", "MPEG1", "MS4A4A", "P2RY12", "RHBDF2", "SAMSN1", "SELPLG", "SIGLEC10", "SIRPB2", "SLC2A5", "SLC37A2", "SUCNR1", "SYTL3", "TLR2", "TNFSF8", "TRIM22")

labels <- rownames(DF_log2FC)
labels[!labels %in% genelist] <- "" 
pheatmap(df_num_scale, scale="row", labels_row = labels, fontsize_row = 6, color = colorRampPalette(c("navy", "white", "firebrick3"))(50))
```

### Correlation plot DEG (Spearman)
#1204 genes

```{r corrplot_GFM_cor, echo=TRUE, message=FALSE}
com = DF_log2FC[,4:10]
cc = cor(com, method = "spearman")
pheatmap(cc)
```

### PCA across stimulations

```{r PCA, echo=TRUE, message=FALSE}
dat <- data.frame(df_num_scale)
stimulation <- colnames(dat)
as.data.frame(stimulation)
dat <- data.frame(dat)
pca <- prcomp(t(dat), scale=TRUE)
PoV <- round((pca$sdev^2/sum(pca$sdev^2))*100)
#Grab covariates to plot:
library(ggplot2)
library(ggfortify)


#Plot:
ggplot(pca, aes(PC1, PC2, color=stimulation)) +
  geom_point(size=2) +
  xlab(paste0("PC1: ",PoV[1],"% variance")) +
  ylab(paste0("PC2: ",PoV[2],"% variance")) +
  theme_classic()

```


```{r upset_LPS1, echo=TRUE}
list_filters = list(LPS = genes_LPS, IFNy = genes_IFNy, R848 = genes_R848, TNF = genes_TNFa, ATP = genes_ATP, DEX = genes_DEX, IL4 = genes_IL4)
#list_filters
overlap = unlist(list_filters)
```

### Upset plot
 
```{r upset_LPS2, echo=TRUE}
upset(fromList(list_filters), order.by = "freq")
```

# DEG across stimulations FDR 5% in 2 stimulations 
#3161 genes  

```{r load tables1, echo = TRUE, results = 'hide'}
DEG_LPS <- subset(res_name_LPS2, adj.P.Val < 0.05)
DEG_IFNy <- subset(res_name_IFNy2, adj.P.Val < 0.05)
DEG_TNFa <- subset(res_name_TNFa2, adj.P.Val < 0.05)
DEG_R848 <- subset(res_name_R8482, adj.P.Val < 0.05)
DEG_ATP <- subset(res_name_ATP2, adj.P.Val < 0.05)
DEG_DEX <- subset(res_name_DEX2, adj.P.Val < 0.05)
DEG_IL4 <- subset(res_name_IL42, adj.P.Val < 0.05)

genes_LPS <- DEG_LPS$ensembl
genes_ATP <- DEG_ATP$ensembl
genes_R848 <- DEG_R848$ensembl
genes_IFNy <- DEG_IFNy$ensembl
genes_DEX <- DEG_DEX$ensembl
genes_IL4 <- DEG_IL4$ensembl
genes_TNFa <- DEG_TNFa$ensembl

# create a vector with all genes listed
x <- c(genes_LPS, genes_ATP, genes_R848, genes_IFNy, genes_DEX, genes_IL4, genes_TNFa)
# make a dataframe
as.data.frame(x) #12590 genes 
# check if there are genes duplicate 
duplicated(x) 
x[duplicated(x)]
unique(x)
#continue only with duplicated
ensembl <- x[duplicated(x)] #4051 genes

genelist <- as.data.frame(ensembl)
#remove duplicated
ensembl <- genelist$ensembl[!duplicated(genelist$ensembl)]
genelist <- as.data.frame(ensembl) #3159 genes

DEG_genes_LPS = merge(res_name_LPS2, genelist, by ="ensembl")
DEG_genes_IFNy = merge(res_name_IFNy2, genelist, by = "ensembl")
DEG_genes_IL4 = merge(res_name_IL42, genelist, by = "ensembl")
DEG_genes_R848 = merge(res_name_R8482, genelist, by = "ensembl")
DEG_genes_ATP = merge(res_name_ATP2, genelist, by = "ensembl")
DEG_genes_DEX = merge(res_name_DEX2, genelist, by = "ensembl")
DEG_genes_TNFa = merge(res_name_TNFa2, genelist, by = "ensembl")

Expression_stimulations <- cbind(DEG_genes_LPS, DEG_genes_IFNy, DEG_genes_IL4, DEG_genes_R848, DEG_genes_ATP, DEG_genes_DEX, DEG_genes_TNFa)

as.data.frame(Expression_stimulations) 
# changes names in data frame
names(Expression_stimulations)[2] <- "LPS"
names(Expression_stimulations)[10] <- "IFNy"
names(Expression_stimulations)[18] <- "IL4"
names(Expression_stimulations)[26] <- "R848"
names(Expression_stimulations)[34] <- "ATP"
names(Expression_stimulations)[42] <- "DEX"
names(Expression_stimulations)[50] <- "TNFa"

DF_log2FC <- Expression_stimulations %>% select(1, 2, 10, 18, 26, 34, 42, 50)
DF <- as.data.frame(DF_log2FC)
DF <- t(DF)
head( DF )[,1:10]

DF_log2FC <- merge(gencode, DF_log2FC, by = "ensembl")
DF_log2FC$symbol = DF_log2FC$GeneSymbol

# fill NA with 0
DF_log2FC[is.na(DF_log2FC)] = 0
#save(DF_log2FC, file = "kmeans3151gsupport.Rdata")

# remove the first column
df_num = as.matrix(DF_log2FC[,4:10])
# scale the data 
df_num_scale = scale(df_num)
# make rownames as function
rownames(df_num) = sapply(DF_log2FC$symbol,function(x) 
strsplit(as.character(x),split = "\\\\")[[1]][1])
save(df_num, file = "df_num_FDR0.52stim_ensembl.Rdata")
#save(df_num_scale, file = "df_num_scale_FDR0.52stim_ensembl.Rdata")
```

### Overlap 3161 genes with Patir

```{r overlap2, echo=TRUE}
gencode <- read.delim("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
gencode$ensembl = gencode$gene_id
genelist <- merge(gencode, genelist, by = "ensembl")
genelist$symbol = genelist$GeneSymbol
Patir <- read.csv("~/Documents/MiGASti/Databases/Patir.txt", sep="")
names(Patir)[1] <- "symbol"
overlap2 <- merge(Patir, genelist, by = "symbol")
list(overlap2$symbol)
```

### Heatmap DEG
#3161 genes

```{r heatmap_cor_3161, echo=TRUE}
rownames(DF_log2FC) = DF_log2FC$symbol
genelist <- c("C1QB", "C1QC", "C3AR1", "CASP1", "CCR1","CD33", "MS4A4A", "TREM2", "TRIM22", "TMEM119", "HLA-DPA1", "HLA-DQB1", "HLA-DRA",  "HLA-DRB1", "IL6R")

labels <- rownames(DF_log2FC)
labels[!labels %in% genelist] <- "" 
pheatmap(df_num_scale, scale="row", labels_row = labels, fontsize_row = 6, color = colorRampPalette(c("navy", "white", "firebrick3"))(50))

pheatmap(df_num_scale, scale = "row", main = "Differentially expressed genes in 2 or more stimulations")
```

### Correlation plot DEG (Spearman)
#3161 genes

```{r corrplot_3161, echo=TRUE, message=FALSE}
com = DF_log2FC[,4:10]
cc = cor(com, method = "spearman")
pheatmap(cc)
```

### PCA across stimulations

```{r PCA_stim, echo=TRUE, message=FALSE}
dat <- data.frame(df_num_scale)
stimulation <- colnames(dat)
as.data.frame(stimulation)
dat <- data.frame(dat)
pca <- prcomp(t(dat), scale=TRUE)
PoV <- round((pca$sdev^2/sum(pca$sdev^2))*100)
#Grab covariates to plot:
library(ggplot2)
library(ggfortify)


#Plot:
ggplot(pca, aes(PC1, PC2, color=stimulation)) +
  geom_point(size=2) +
  xlab(paste0("PC1: ",PoV[1],"% variance")) +
  ylab(paste0("PC2: ",PoV[2],"% variance")) +
  theme_classic()
```


```{r upset_LP4, echo=TRUE}
list_filters = list(LPS = DEG_genes_LPS$symbol, IFNy = DEG_genes_IFNy$symbol, R848 = DEG_genes_R848$symbol, TNF = DEG_genes_TNFa$symbol, ATP = DEG_genes_ATP$symbol, DEX = DEG_genes_DEX$symbol, IL4 = DEG_genes_IL4$symbol)
#list_filters
overlap = unlist(list_filters)
```

### Upset plot
 
```{r upset_LPS3, echo=TRUE}
upset(fromList(list_filters), order.by = "freq")
```


# DEG across stimulations FDR 5%
#8523 genes 

```{r load tables20, echo = TRUE, results = 'hide'}
#check duplicated genes in list 
#[1] "ATF7"    "TMSB15B" "POLR2J4" "MATR3"   "HSPA14"  "SCO2" "TBCE"    "ATXN7"   "ABCF2"   "POLR2J3" "GGT1"  

DEG_LPS <- subset(res_name_LPS2, adj.P.Val < 0.05)
DEG_IFNy <- subset(res_name_IFNy2, adj.P.Val < 0.05)
DEG_TNFa <- subset(res_name_TNFa2, adj.P.Val < 0.05)
DEG_R848 <- subset(res_name_R8482, adj.P.Val < 0.05)
DEG_ATP <- subset(res_name_ATP2, adj.P.Val < 0.05)
DEG_DEX <- subset(res_name_DEX2, adj.P.Val < 0.05)
DEG_IL4 <- subset(res_name_IL42, adj.P.Val < 0.05)

genes_LPS <- DEG_LPS$ensembl
genes_ATP <- DEG_ATP$ensembl
genes_R848 <- DEG_R848$ensembl
genes_IFNy <- DEG_IFNy$ensembl
genes_DEX <- DEG_DEX$ensembl
genes_IL4 <- DEG_IL4$ensembl
genes_TNFa <- DEG_TNFa$ensembl

# create a vector with all genes listed
x <- c(genes_LPS, genes_ATP, genes_R848, genes_IFNy, genes_DEX, genes_IL4, genes_TNFa)
# make a dataframe
as.data.frame(x) #12590 genes 
# check if there are genes duplicate 
duplicated(x) 
unique(x)
#remove duplicated
ensembl <- x[!duplicated(x)] #8506 genes

genelist <- as.data.frame(ensembl)
ensembl <- genelist$ensembl[!duplicated(genelist$ensembl)]
genelist <- as.data.frame(ensembl) #8506 genes 
genelist %>% distinct(ensembl, .keep_all = TRUE)

DEG_genes_LPS = merge(res_name_LPS2, genelist, by ="ensembl")
DEG_genes_IFNy = merge(res_name_IFNy2, genelist, by = "ensembl")
DEG_genes_IL4 = merge(res_name_IL42, genelist, by = "ensembl")
DEG_genes_R848 = merge(res_name_R8482, genelist, by = "ensembl")
DEG_genes_ATP = merge(res_name_ATP2, genelist, by = "ensembl")
DEG_genes_DEX = merge(res_name_DEX2, genelist, by = "ensembl")
DEG_genes_TNFa = merge(res_name_TNFa2, genelist, by = "ensembl")

Expression_stimulations <- cbind(DEG_genes_LPS, DEG_genes_IFNy, DEG_genes_IL4, DEG_genes_R848, DEG_genes_ATP, DEG_genes_DEX, DEG_genes_TNFa)

as.data.frame(Expression_stimulations) 
# changes names in data frame
names(Expression_stimulations)[2] <- "LPS"
names(Expression_stimulations)[10] <- "IFNy"
names(Expression_stimulations)[18] <- "IL4"
names(Expression_stimulations)[26] <- "R848"
names(Expression_stimulations)[34] <- "ATP"
names(Expression_stimulations)[42] <- "DEX"
names(Expression_stimulations)[50] <- "TNFa"

DF_log2FC <- Expression_stimulations %>% select(1, 2, 10, 18, 26, 34, 42, 50)
DF <- as.data.frame(DF_log2FC)
DF <- t(DF)
head( DF )[,1:10]


# fill NA with 0
DF_log2FC[is.na(DF_log2FC)] = 0
#save(DF_log2FC, file = "kmeans8523ensembl.Rdata")

DF_log2FC <- merge(gencode, DF_log2FC, by = "ensembl")
DF_log2FC$symbol = DF_log2FC$GeneSymbol

#remove rownames that are duplicates (GGT1, HSPA14)
DF_log2FC = DF_log2FC[-c(1122, 6885),]

# remove the first column
df_num = as.matrix(DF_log2FC[,4:10])
# scale the data 
df_num_scale = scale(df_num)
# make rownames as function
rownames(df_num) = sapply(DF_log2FC$symbol,function(x) 
strsplit(as.character(x),split = "\\\\")[[1]][1])
save(df_num, file = "df_num_FDR0.5stim_allgenes.ensembl.Rdata")
#save(DF_log2FC, file = "FDR0.5_stimulations_allgenes.Rdata")
#save(df_num_scale, file = "df_num_scale_FDR0.5stim_allgenes_ensembl.Rdata")
```

### Overlap 8523 genes with Patir

```{r overlap_patir, echo=TRUE}
gencode <- read.delim("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
gencode$ensembl = gencode$gene_id
genelist <- merge(gencode, genelist, by = "ensembl")
genelist$symbol = genelist$GeneSymbol
Patir <- read.csv("~/Documents/MiGASti/Databases/Patir.txt", sep="")
names(Patir)[1] <- "symbol"
overlap2 <- merge(Patir, genelist, by = "symbol")
list(overlap2$symbol)
```

### Heatmap DEG
#8521 genes

```{r heatmap_DEG_all, echo=TRUE}

rownames(DF_log2FC) = DF_log2FC$symbol
genelist <- c("CD33", "MS4A4A", "TREM2", "TYROBP", "TLR1", "HLA-DPA1", "HLA-DQB1", "HLA-DRA",  "HLA-DRB1", "IL6R", "TLR6", "FCGR2A", "AIF1", "ITGAX", "ITGAM", "HLA-DRB5" , "CD84")

labels <- rownames(DF_log2FC)
labels[!labels %in% genelist] <- "" 
pheatmap(df_num_scale, scale="row", labels_row = labels, fontsize_row = 6, color = colorRampPalette(c("navy", "white", "firebrick3"))(50))

pheatmap(df_num_scale, scale = "row", main = "Differentially expressed genes")
```


### Correlation plot DEG (Spearman)
#8523

```{r corr_all, echo=TRUE, message=FALSE}
com = DF_log2FC[,4:10]
cc = cor(com, method = "spearman")
pheatmap(cc)
```


### PCA across stimulations

```{r PCA_all, echo=TRUE, message=FALSE}
dat <- data.frame(df_num_scale)
stimulation <- colnames(dat)
as.data.frame(stimulation)
dat <- data.frame(dat)
pca <- prcomp(t(dat), scale=TRUE)
PoV <- round((pca$sdev^2/sum(pca$sdev^2))*100)
#Grab covariates to plot:
library(ggplot2)
library(ggfortify)


#Plot:
ggplot(pca, aes(PC1, PC2, color=stimulation)) +
  geom_point(size=2) +
  xlab(paste0("PC1: ",PoV[1],"% variance")) +
  ylab(paste0("PC2: ",PoV[2],"% variance")) +
  theme_classic()

```


```{r upset_LPS_all, echo=TRUE}
list_filters = list(LPS = genes_LPS, IFNy = genes_IFNy, R848 = genes_R848, TNF = genes_TNFa, ATP = genes_ATP, DEX = genes_DEX, IL4 = genes_IL4)
#list_filters
overlap = unlist(list_filters)
```

### Upset plot
 
```{r upset_LPS2_all, echo=TRUE}
upset(fromList(list_filters), order.by = "freq")
```



# Scatterplots 
### LogFC LPS vs IFNy 
#18981 genes 

```{r scat_LPS_ifny, echo=TRUE}
names(res_name_LPS2)[2] <- "LPS"
names(res_name_IFNy2)[2] <- "IFNy"
DEG_genes_LPS_IFNy = merge(res_name_LPS2, res_name_IFNy2, by ="symbol")

LPS <- DEG_genes_LPS_IFNy[,3]
IFNy <- DEG_genes_LPS_IFNy[,12]
df = data.frame(LPS, IFNy)
p <- ggplot(df, aes(x=LPS, y=IFNy) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  scale_fill_continuous(type = "viridis") +
  theme_bw()
p + stat_cor(method = "spearman")
```


### LogFC LPS vs R848 

```{r scat_LPS_R848, echo=TRUE}
names(res_name_LPS2)[2] <- "LPS"
names(res_name_R8482)[2] <- "R848"
DEG_genes_LPS_R848 = merge(res_name_LPS2, res_name_R8482, by ="symbol")

LPS <- DEG_genes_LPS_R848[,3]
R848 <- DEG_genes_LPS_R848[,12]
df = data.frame(LPS, R848)
p <- ggplot(df, aes(x=LPS, y=R848) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  scale_fill_continuous(type = "viridis") +
  theme_bw()
p + stat_cor(method = "spearman")
```

### LogFC LPS vs TNFa 

```{r scat_LPS_TNFa, echo=TRUE}
names(res_name_LPS2)[2] <- "LPS"
names(res_name_TNFa2)[2] <- "TNFa"
DEG_genes_LPS_TNFa = merge(res_name_LPS2, res_name_TNFa2, by ="symbol")

LPS <- DEG_genes_LPS_TNFa[,3]
TNFa <- DEG_genes_LPS_TNFa[,12]
df = data.frame(LPS, TNFa)
p <- ggplot(df, aes(x=LPS, y=TNFa) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  scale_fill_continuous(type = "viridis") +
  theme_bw()
p + stat_cor(method = "spearman")
```

### LogFC LPS vs DEX 

```{r scat_LPS_DEX, echo=TRUE}
names(res_name_LPS2)[2] <- "LPS"
names(res_name_DEX2)[2] <- "DEX"
DEG_genes_LPS_DEX = merge(res_name_LPS2, res_name_DEX2, by ="symbol")

LPS <- DEG_genes_LPS_DEX[,3]
DEX <- DEG_genes_LPS_DEX[,12]
df = data.frame(LPS, DEX)
p <- ggplot(df, aes(x=LPS, y=DEX) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  scale_fill_continuous(type = "viridis") +
  theme_bw()
p + stat_cor(method = "spearman")
```

### LogFC LPS vs IL4 

```{r scat_LPS2_IL4, echo=TRUE}
names(res_name_LPS2)[2] <- "LPS"
names(res_name_IL42)[2] <- "IL4"
DEG_genes_LPS_IL4 = merge(res_name_LPS2, res_name_IL42, by ="symbol")

LPS <- DEG_genes_LPS_IL4[,3]
IL4 <- DEG_genes_LPS_IL4[,12]
df = data.frame(LPS, IL4)
p <- ggplot(df, aes(x=LPS, y=IL4) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  scale_fill_continuous(type = "viridis") +
  theme_bw()
p + stat_cor(method = "spearman")
```

### LogFC LPS vs ATP 

```{r scat_LPS_ATP, echo=TRUE}
names(res_name_LPS2)[2] <- "LPS"
names(res_name_ATP2)[2] <- "ATP"
DEG_genes_LPS_ATP = merge(res_name_LPS2, res_name_ATP2, by ="symbol")

LPS <- DEG_genes_LPS_ATP[,3]
ATP <- DEG_genes_LPS_ATP[,12]
df = data.frame(LPS, ATP)
p <- ggplot(df, aes(x=LPS, y=ATP) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  scale_fill_continuous(type = "viridis") +
  theme_bw()
p + stat_cor(method = "spearman")
```




