---
title: "Differential expression analysis LPS and IFNy"
author: "Gijsje"
date: "2/16/2021"
output: 
 rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---

```{r load.packages, echo=FALSE, message=FALSE, results='hide'}
library(readxl)
library(edgeR)
library(limma)
library(DESeq2)
library(tidyverse)
library(data.table)


knitr::opts_chunk$set( 
  warning=FALSE,
  message=FALSE,
  results = 'asis',
  error = FALSE,
  tidy = FALSE,
  fig.show = "hold")
```


```{r data_load, echo = TRUE, results = 'hide'}
load("~/Documents/MiGASti/Databases/gene_matrix.RData")
metadata <- read.table("~/Documents/MiGASti/Databases/metadata.txt")
#set rownames to Sample
row.names(metadata) <- metadata$Sample 
setwd("~/Documents/MiGASti/Databases")
#exclude samples that did not pass QC filtering
exclude <- read.table("samples2remove.txt")
exclude <- exclude$x
genes_counts_filt = genes_counts[, !colnames(genes_counts) %in% exclude] 
#Excludes the samples from filters. 
#dim(genes_counts_filt)
metadata_filt = metadata[ !(rownames(metadata) %in% exclude), ]
length(metadata_filt)
gencode_30 = read.table("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
colnames(gencode_30) = c("ensembl","symbol")
#remove low count genes
cpm <- cpm(genes_counts_filt) 
# CPM >= 1 in at least 50% of the samples
keep.exp <- rowSums(cpm > 1) >= (0.5 * ncol(genes_counts_filt) )
genes_counts_filt1 <- genes_counts_filt[ keep.exp, ] #18997 genes 
#order metadata and genes counts
genes_counts_ordered <- genes_counts_filt1[,rownames(metadata_filt)]
#head(genes_counts_ordered)
all(rownames(metadata_filt) == colnames (genes_counts_ordered)) #TRUE
```


```{r data_load_outline, echo = TRUE}
# Counts, TPM, voom and metadata filtered 
# 496 samples!  
# Filter for genes not expressed: 50 % of the samples! 
# dim(genes_counts_ordered) #18997   496
# dim(metadata_filt) #496  38
```

# Preparing the samples for DEG

```{r preparing, echo = TRUE}
#remove uncultured samples
metadata_cultured <- metadata_filt[metadata_filt$Stimulation != "ununstim",]
#check numbers
#dim(metadata_cultured)
#check numbers per stimulation
#table(metadata_filt$Stimulation)
#select only GFM samples in metadata
metadata_GFM = metadata_cultured[metadata_cultured$Region=="GFM",]
#select only GFM samples in genes counts
genes_counts_GFM <- genes_counts_ordered[,metadata_GFM$Sample]
#order metadata and genes_counts
genes_counts_GFM_ordered <- genes_counts_GFM[,rownames(metadata_GFM)]
#check ordering
all(rownames(metadata_GFM) == colnames (genes_counts_GFM_ordered)) #TRUE

# select the condition to compare
pheno_LPS <- subset(metadata_GFM, Stimulation == "unstim" | Stimulation == "LPS")
exp_LPS <- genes_counts_GFM_ordered[, rownames(pheno_LPS)]
identical(rownames(pheno_LPS), colnames(exp_LPS))

pheno_IFNy <- subset(metadata_GFM, Stimulation == "unstim" | Stimulation == "IFNy")
exp_IFNy <- genes_counts_GFM_ordered[, rownames(pheno_IFNy)]
identical(rownames(pheno_IFNy), colnames(exp_IFNy))

#round counts; deseq2 can only handle integers
exp_LPS <- round(exp_LPS, digits=0)
exp_IFNy <- round(exp_IFNy, digits=0)

#make sure covariate variables are the right format 
#cannot create dds object with numeric values
pheno_LPS$Donor_id <- as.factor(pheno_LPS$Donor_id)
pheno_LPS$age <- as.integer(pheno_LPS$age)
pheno_LPS$Stimulation <- as.factor(pheno_LPS$Stimulation)
pheno_LPS$picard_pct_ribosomal_bases = scale(pheno_LPS$picard_pct_ribosomal_bases)
pheno_LPS$picard_pct_mrna_bases = scale(pheno_LPS$picard_pct_mrna_bases)
pheno_LPS$picard_pct_pf_reads_aligned = scale(pheno_LPS$picard_pct_pf_reads_aligned)
pheno_LPS$picard_pct_percent_duplication = scale(pheno_LPS$picard_pct_percent_duplication)


pheno_IFNy$age <- as.integer(pheno_IFNy$age)
pheno_IFNy$Stimulation <- as.factor(pheno_IFNy$Stimulation)
pheno_IFNy$picard_pct_mrna_bases <- as.integer(pheno_IFNy$picard_pct_mrna_bases)
pheno_IFNy$picard_pct_ribosomal_bases <- as.integer(pheno_IFNy$picard_pct_ribosomal_bases)

#adjust for: ~ age + (1|donor_id) + picard_pct_ribosomal_bases + picard_pct_mrna_bases + (1|sex)  + picard_pct_percent_duplication + picard_pct_pf_reads_aligned 
```

# DESEQ2 for LPS 

```{r data_load_LPS, echo = TRUE}
#createDeSEQ2 object for LPS
dds <- DESeqDataSetFromMatrix(countData = exp_LPS,
                              colData = pheno_LPS,
                              design = ~ age + picard_pct_ribosomal_bases + picard_pct_mrna_bases + picard_pct_percent_duplication + picard_pct_pf_reads_aligned + Stimulation) 
#variable of interest at end of the formula

#Make sure that control group is set as the reference group
dds$Stimulation <- relevel(dds$Stimulation, ref="unstim")
summary(dds$Stimulation) # 38 unstim #35 LPS
head(dds)

#Run Differential Expression Analysis 
dds <- DESeq(dds)
res <- results(dds, contrasts())
sum(res$padj < 0.05, na.rm=TRUE)
resOrdered <- res[order(res$pvalue),] 
resOrdered <- as.data.frame(resOrdered)
#write.table(resOrdered, "DEG_LPS_GFM.txt")

#PCA
vsd <- vst(dds, blind=FALSE) 
plotPCA(vsd, intgroup = "Stimulation")
```

# MA plot 
#In DESeq2, the function plotMA shows the log2 fold changes attributable to a given variable over the mean of normalized counts for all the samples in the DESeqDataSet. Points will be colored red if the adjusted p value is less than 0.1. Points which fall out of the window are plotted as open triangles pointing either up or down.

```{r data_load_MA, echo = TRUE}
plotMA(res, ylim=c(-2,2))
```




# DESEQ2 for IFNy

```{r data_load_IFNy, echo = TRUE}
#createDeSEQ2 object for LPS
dds <- DESeqDataSetFromMatrix(countData = exp_IFNy,
                              colData = pheno_IFNy,
                              design = ~ age + picard_pct_mrna_bases + picard_pct_ribosomal_bases + Stimulation) 
#variable of interest at end of the formula

#Make sure that control group is set as the reference group
dds$Stimulation <- relevel(dds$Stimulation, ref="unstim")
summary(dds$Stimulation) # 38 unstim #28 IFNy
head(dds)

#Run Differential Expression Analysis 
dds <- DESeq(dds)
res <- results(dds)
sum(res$padj < 0.05, na.rm=TRUE)
resOrdered <- res[order(res$pvalue),] 
resOrdered <- as.data.frame(resOrdered)
write.table(resOrdered, "DEG_IFNy_GFM.txt")

#PCA
vsd <- vst(dds, blind=FALSE) 
plotPCA(vsd, intgroup = "Stimulation")
dev.off()
```




