---
title: "202100413_Kmeans_clustering_all"
output: 
rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true 
---

```{r setu11p, echo=TRUE}
library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization
library(gProfileR)
```

# Treshold FDR 5% log FC > 1 or < -1 (210 genes) only LPS and IFNy

```{r data_load, echo=TRUE}
load("~/df_num_LPS_IFNy_LOGFC1.Rdata")
load("~/Documents/MiGASti/docs/Kmeans210gsupport_LPS_IFNy.Rdata")
df <- df_num
#scale data:As we don?t want the clustering algorithm to depend to an arbitrary variable unit, we start by scaling/standardizing 
df3 <- scale(df)
```

### Elbow method:

```{r elbow2, echo=TRUE}
set.seed(123)
fviz_nbclust(df3, kmeans, method = "wss")
```

### Average silhouette approach 

```{r silhoutte2, echo = TRUE}
fviz_nbclust(df3, kmeans, method = "silhouette")
```

### Gap statistic clustering 

```{r gap1, echo=TRUE}
# compute gap statistic
set.seed(123)
gap_stat <- clusGap(df3, FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
# Print the result
print(gap_stat, method = "firstmax")
fviz_gap_stat(gap_stat)
```

# summary 
### 2 clusters

```{r gap2, echo=TRUE}
k2 <- kmeans(df3, centers = 2, nstart = 25)
str(k2)
fviz_cluster(k2, data = df3)
```

### summary 3 clusters

```{r gap3, echo=TRUE}
k3 <- kmeans(df3, centers = 3, nstart = 25)
str(k3)
fviz_cluster(k3, data = df3)
```

### summary 4 clusters

```{r gap545, echo=TRUE}
k4 <- kmeans(df3, centers = 4, nstart = 25)
str(k4)
fviz_cluster(k4, data = df3)
```


### summary 5 clusters

```{r gap5, echo=TRUE}
k5 <- kmeans(df3, centers = 5, nstart = 25)
str(k5)
fviz_cluster(k5, data = df3)
```

### summary 6 clusters

```{r gap47, echo=TRUE}
k6 <- kmeans(df3, centers = 6, nstart = 25)
str(k6)
fviz_cluster(k6, data = df3)
```

### Compare 2 clusters
#spearman correlation coefficient between 2 clusters

```{r gap20, echo=TRUE}
#first run
k <- kmeans(df3, centers = 2, nstart = 25)
kmeans <- k$cluster
kmean_download <- as.data.frame(kmeans)
kmeanscor <- cbind(kmean_download, DF_log2FC)
gencode_30 <- read.delim("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
colnames(gencode_30) = c("ensembl","symbol")
kmeanscor1 = merge(kmeanscor, gencode_30, by="ensembl")

#second run
k <- kmeans(df3, centers = 2, nstart = 25)
kmeans <- k$cluster
kmean_download <- as.data.frame(kmeans)
kmeanscor <- cbind(kmean_download, DF_log2FC)
kmeanscor2 = merge(kmeanscor, gencode_30, by="ensembl")

res <- cor.test(kmeanscor1$kmeans, kmeanscor2$kmeans, method = "pearson")
res 
```

### comparison between 2 clusters based on gprofiler results 

```{r gap12, echo=TRUE}
cluster1 <- kmeanscor1[kmeanscor1$kmeans == 1,]
cluster2 <- kmeanscor1[kmeanscor1$kmeans == 2,]


gprofiler_results_1 <- gprofiler(query = cluster1$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_2 <- gprofiler(query = cluster2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")



head(gprofiler_results_1)
head(gprofiler_results_2)


cluster1_2 <- kmeanscor2[kmeanscor2$kmeans == 1,]
cluster2_2 <- kmeanscor2[kmeanscor2$kmeans == 2,]


gprofiler_results_1_2 <- gprofiler(query = cluster1_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_2_2 <- gprofiler(query = cluster2_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")



head(gprofiler_results_1_2$term.name)
head(gprofiler_results_2_2$term.name)
```


### Compare 3 clusters
#spearman correlation coefficient between 3 clusters

```{r gap10230, echo=TRUE}
k <- kmeans(df3, centers = 3, nstart = 25)
kmeans <- k$cluster
kmean_download <- as.data.frame(kmeans)
kmeanscor <- cbind(kmean_download, DF_log2FC)
kmeanscor1 = merge(kmeanscor, gencode_30, by="ensembl")

k <- kmeans(df3, centers = 3, nstart = 25)
kmeans <- k$cluster
kmean_download <- as.data.frame(kmeans)
kmeanscor <- cbind(kmean_download, DF_log2FC)
kmeanscor2 = merge(kmeanscor, gencode_30, by="ensembl")

res <- cor.test(kmeanscor1$kmeans, kmeanscor2$kmeans, method = "pearson")
res 
```

### Compare 4 clusters
#spearman correlation coefficient between 4 clusters

```{r gap1034550, echo=TRUE}
k <- kmeans(df3, centers = 4, nstart = 25)
kmeans <- k$cluster
kmean_download <- as.data.frame(kmeans)
kmeanscor <- cbind(kmean_download, DF_log2FC)
kmeanscor1 = merge(kmeanscor, gencode_30, by="ensembl")

k <- kmeans(df3, centers = 4, nstart = 25)
kmeans <- k$cluster
kmean_download <- as.data.frame(kmeans)
kmeanscor <- cbind(kmean_download, DF_log2FC)
kmeanscor2 = merge(kmeanscor, gencode_30, by="ensembl")

res <- cor.test(kmeanscor1$kmeans, kmeanscor2$kmeans, method = "pearson")
res 
```

### comparison between 4 clusters based on gprofiler results 

```{r gap1221294, echo=TRUE}
cluster1 <- kmeanscor1[kmeanscor1$kmeans == 1,]
cluster2 <- kmeanscor1[kmeanscor1$kmeans == 2,]
cluster3 <- kmeanscor1[kmeanscor1$kmeans == 3,]
cluster4 <- kmeanscor1[kmeanscor1$kmeans == 4,]


gprofiler_results_1 <- gprofiler(query = cluster1$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_2 <- gprofiler(query = cluster2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_3 <- gprofiler(query = cluster3$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_4 <- gprofiler(query = cluster4$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")


head(gprofiler_results_1$term.name)
head(gprofiler_results_2$term.name)
head(gprofiler_results_3$term.name)
head(gprofiler_results_4$term.name)


cluster1_2 <- kmeanscor2[kmeanscor2$kmeans == 1,]
cluster2_2 <- kmeanscor2[kmeanscor2$kmeans == 2,]
cluster3_2 <- kmeanscor2[kmeanscor2$kmeans == 3,]
cluster4_2 <- kmeanscor2[kmeanscor2$kmeans == 4,]



gprofiler_results_1_2 <- gprofiler(query = cluster1_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_2_2 <- gprofiler(query = cluster2_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_3_2 <- gprofiler(query = cluster3_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_4_2 <- gprofiler(query = cluster4_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")


head(gprofiler_results_1_2$term.name)
head(gprofiler_results_2_2$term.name)
head(gprofiler_results_3_2$term.name)
head(gprofiler_results_4_2$term.name)
```

### Compare 6 clusters
#spearman correlation coefficient between 6 clusters

```{r gap1012, echo=TRUE}
# first run
k <- kmeans(df3, centers = 6, nstart = 25)
kmeans <- k$cluster
kmean_download <- as.data.frame(kmeans)
kmeanscor <- cbind(kmean_download, DF_log2FC)
kmeanscor1 = merge(kmeanscor, gencode_30, by="ensembl")

# second run 
k <- kmeans(df3, centers = 6, nstart = 25)
kmeans <- k$cluster
kmean_download <- as.data.frame(kmeans)
kmeanscor <- cbind(kmean_download, DF_log2FC)
kmeanscor2 = merge(kmeanscor, gencode_30, by="ensembl")

res <- cor.test(kmeanscor1$kmeans, kmeanscor2$kmeans, method = "pearson")
res 
```

### comparison between 6 clusters based on gprofiler results 

```{r gap1322, echo=TRUE}
cluster1 <- kmeanscor1[kmeanscor1$kmeans == 1,]
cluster2 <- kmeanscor1[kmeanscor1$kmeans == 2,]
cluster3 <- kmeanscor1[kmeanscor1$kmeans == 3,]
cluster4 <- kmeanscor1[kmeanscor1$kmeans == 4,]
cluster5 <- kmeanscor1[kmeanscor1$kmeans == 5,]
cluster6 <- kmeanscor1[kmeanscor1$kmeans == 6,]

gprofiler_results_1 <- gprofiler(query = cluster1$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_2 <- gprofiler(query = cluster2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_3 <- gprofiler(query = cluster3$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_4 <- gprofiler(query = cluster4$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_5 <- gprofiler(query = cluster5$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_6 <- gprofiler(query = cluster6$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

head(gprofiler_results_1$term.name)
head(gprofiler_results_2$term.name)
head(gprofiler_results_3$term.name)
head(gprofiler_results_4$term.name)
head(gprofiler_results_5$term.name)
head(gprofiler_results_6$term.name)

cluster1_2 <- kmeanscor2[kmeanscor2$kmeans == 1,]
cluster2_2 <- kmeanscor2[kmeanscor2$kmeans == 2,]
cluster3_2 <- kmeanscor2[kmeanscor2$kmeans == 3,]
cluster4_2 <- kmeanscor2[kmeanscor2$kmeans == 4,]
cluster5_2 <- kmeanscor2[kmeanscor2$kmeans == 5,]
cluster6_2 <- kmeanscor2[kmeanscor2$kmeans == 6,]


gprofiler_results_1_2 <- gprofiler(query = cluster1_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_2_2 <- gprofiler(query = cluster2_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_3_2 <- gprofiler(query = cluster3_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_4_2 <- gprofiler(query = cluster4_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_5_2 <- gprofiler(query = cluster5_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_6_2 <- gprofiler(query = cluster6_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

head(gprofiler_results_1_2$term.name)
head(gprofiler_results_2_2$term.name)
head(gprofiler_results_3_2$term.name)
head(gprofiler_results_4_2$term.name)
head(gprofiler_results_5_2$term.name)
head(gprofiler_results_6_2$term.name)
```

### Compare 7 clusters
#spearman correlation coefficient between 7 clusters

```{r gap13, echo=TRUE}
#first run 
k <- kmeans(df3, centers = 7, nstart = 25)
kmeans <- k$cluster
kmean_download <- as.data.frame(kmeans)
kmeanscor <- cbind(kmean_download, DF_log2FC)
kmeanscor1 = merge(kmeanscor, gencode_30, by="ensembl")

#second run 
k <- kmeans(df3, centers = 7, nstart = 25)
kmeans <- k$cluster
kmean_download <- as.data.frame(kmeans)
kmeanscor <- cbind(kmean_download, DF_log2FC)
kmeanscor2 = merge(kmeanscor, gencode_30, by="ensembl")

res <- cor.test(kmeanscor1$kmeans, kmeanscor2$kmeans, method = "pearson")
res 
```

### comparison between 7 clusters based on gprofiler results 

```{r gap12289, echo=TRUE}
cluster1 <- kmeanscor1[kmeanscor1$kmeans == 1,]
cluster2 <- kmeanscor1[kmeanscor1$kmeans == 2,]
cluster3 <- kmeanscor1[kmeanscor1$kmeans == 3,]
cluster4 <- kmeanscor1[kmeanscor1$kmeans == 4,]
cluster5 <- kmeanscor1[kmeanscor1$kmeans == 5,]
cluster6 <- kmeanscor1[kmeanscor1$kmeans == 6,]
cluster7 <- kmeanscor1[kmeanscor1$kmeans == 7,]

gprofiler_results_1 <- gprofiler(query = cluster1$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_2 <- gprofiler(query = cluster2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_3 <- gprofiler(query = cluster3$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_4 <- gprofiler(query = cluster4$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_5 <- gprofiler(query = cluster5$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_6 <- gprofiler(query = cluster6$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")
gprofiler_results_7 <- gprofiler(query = cluster7$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

head(gprofiler_results_1$term.name)
head(gprofiler_results_2$term.name)
head(gprofiler_results_3$term.name)
head(gprofiler_results_4$term.name)
head(gprofiler_results_5$term.name)
head(gprofiler_results_6$term.name)
head(gprofiler_results_7$term.name)

cluster1_2 <- kmeanscor2[kmeanscor2$kmeans == 1,]
cluster2_2 <- kmeanscor2[kmeanscor2$kmeans == 2,]
cluster3_2 <- kmeanscor2[kmeanscor2$kmeans == 3,]
cluster4_2 <- kmeanscor2[kmeanscor2$kmeans == 4,]
cluster5_2 <- kmeanscor2[kmeanscor2$kmeans == 5,]
cluster6_2 <- kmeanscor2[kmeanscor2$kmeans == 6,]
cluster7_2 <- kmeanscor2[kmeanscor2$kmeans == 7,]

gprofiler_results_1_2 <- gprofiler(query = cluster1_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_2_2 <- gprofiler(query = cluster2_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_3_2 <- gprofiler(query = cluster3_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_4_2 <- gprofiler(query = cluster4_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_5_2 <- gprofiler(query = cluster5_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_6_2 <- gprofiler(query = cluster6_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

gprofiler_results_7_2 <- gprofiler(query = cluster7_2$symbol,                                   organism = "hsapiens",
                                  ordered_query = F, 
                                  exclude_iea = F, 
                                  max_p_value = 0.05, 
                                  max_set_size = 0,
                                  correction_method = "fdr",
                                  hier_filtering = "none", 
                                  domain_size = "annotated")

head(gprofiler_results_1_2$term.name)
head(gprofiler_results_2_2$term.name)
head(gprofiler_results_3_2$term.name)
head(gprofiler_results_4_2$term.name)
head(gprofiler_results_5_2$term.name)
head(gprofiler_results_6_2$term.name)
head(gprofiler_results_7_2$term.name)
```




