---
title: "20210224_DEG_heatmap"
author: "Gijsje"
date: "2/24/2021"
output: 
  rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---


```{r packages, echo=FALSE, message=FALSE, results='hide'}
library(readxl)
library(edgeR)
library(limma)
library(DESeq2)
library(tidyverse)
library(data.table)
library(shiny)
library(DT)
library(pheatmap)
library(ggfortify)
library(ggplot2)
library(UpSetR)
library(VennDiagram)
library(knitr) 
library(tidyverse)
library(dplyr)
library(tidyr)
library(tibble)
library(purrr)
library(corrplot)
library(ggplot2)
library(ggpubr)


knitr::opts_chunk$set( 
  warning=FALSE,
  message=FALSE,
  results = 'asis',
  error = FALSE,
  tidy = FALSE,
  fig.show = "hold")
```


# DEG across stimulations FDR % log FC > 1 or -1   
#210 genes 

```{r load tables, echo = TRUE, results = 'hide'}
LPS <- read.csv("~/Documents/MiGASti/Databases/res_name_LPS_LFC1.txt", sep="")
IFNy <- read.csv("~/Documents/MiGASti/Databases/res_name_IFNy_LFC1.txt", sep="")

genes_LPS <- LPS$ensembl
genes_IFNy <- IFNy$ensembl
    
# create a vector with all genes listed
x <- c(genes_LPS, genes_IFNy)
# make a dataframe
as.data.frame(x) 
# check if there are genes duplicate 
duplicated(x) 
unique(x)
#remove duplicated
ensembl <- x[!duplicated(x)] #1204 genes 

genelist <- as.data.frame(ensembl)
ensembl <- genelist$ensembl[!duplicated(genelist$ensembl)]
genelist <- as.data.frame(ensembl) 
genelist %>% distinct(ensembl, .keep_all = TRUE)

load("~/Documents/MiGASti/res_name_LPS2.Rdata")
load("~/Documents/MiGASti/docs/res_name_IFNy2.Rdata")

DEG_genes_LPS = merge(res_name_LPS2, genelist, by ="ensembl")
DEG_genes_IFNy = merge(res_name_IFNy2, genelist, by = "ensembl")

Expression_stimulations <- cbind(DEG_genes_LPS, DEG_genes_IFNy)

as.data.frame(Expression_stimulations) 
# changes names in data frame
names(Expression_stimulations)[2] <- "LPS"
names(Expression_stimulations)[10] <- "IFNy"

DF_log2FC <- Expression_stimulations %>% select(1, 2, 10)
DF <- as.data.frame(DF_log2FC)
DF <- t(DF)
head( DF )[,1:10]

# fill NA with 0
DF_log2FC[is.na(DF_log2FC)] = 0
save(DF_log2FC, file = "Kmeans210gsupport_LPS_IFNy.Rdata")

gencode <- read.delim("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
gencode$ensembl = gencode$gene_id
DF_log2FC <- merge(gencode, DF_log2FC, by = "ensembl")
DF_log2FC$symbol = DF_log2FC$GeneSymbol

# remove the first column
df_num = as.matrix(DF_log2FC[,4:5])
# scale the data 
df_num_scale = scale(df_num)
save(df_num, file = "df_num_LPS_IFNy_LOGFC1.Rdata")
#save(DF_log2FC, file = "LOGFC1_stimulations_LPS_IFNy.Rdata")
save(df_num_scale, file = "df_num_scale_LOGFC1stim_LPS_IFNy.Rdata")
```

### Overlap 210 genes with Patir

```{r overlap, echo = TRUE}
gencode <- read.delim("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
gencode$ensembl = gencode$gene_id
genelist <- merge(gencode, genelist, by = "ensembl")
genelist$symbol = genelist$GeneSymbol
Patir <- read.csv("~/Documents/MiGASti/Databases/Patir.txt", sep="")
names(Patir)[1] <- "symbol"
overlap1 <- merge(Patir, genelist, by = "symbol")
list(overlap1$symbol)
```

### Heatmap DEG
#210 genes (scaling data)

```{r heatmap_GFM_cor, echo = TRUE}
rownames(DF_log2FC) = DF_log2FC$symbol
genelist <- c("C1QB",    "C2",      "CD300A",  "CIITA",   "EBI3",    "FCGR1A", "FCGR1B",  "FGL2",    "GIMAP2",  "GPR183", "HAMP",    "IFI30", "IRF5",    "LILRB2",  "MPEG1",   "SLC37A2", "TRIM22" )

labels <- rownames(DF_log2FC)
labels[!labels %in% genelist] <- "" 
pheatmap(df_num_scale, labels_row = labels, fontsize_row = 6)
```

### Correlation between log FC LPS and IFNy (Spearman)
#1204 genes

```{r corrplot_GFM_cor, echo=TRUE, message=FALSE}
df <- as.data.frame(df_num_scale)
res <- cor.test(df$LPS, df$IFNy, method = "spearman")
res
```

### PCA across stimulations

```{r PCA, echo=TRUE, message=FALSE}
dat <- data.frame(df_num_scale)
stimulation <- colnames(dat)
as.data.frame(stimulation)
dat <- data.frame(dat)
pca <- prcomp(t(dat), scale=TRUE)
PoV <- round((pca$sdev^2/sum(pca$sdev^2))*100)
#Grab covariates to plot:
library(ggplot2)
library(ggfortify)


#Plot:
ggplot(pca, aes(PC1, PC2, color=stimulation)) +
  geom_point(size=2) +
  xlab(paste0("PC1: ",PoV[1],"% variance")) +
  ylab(paste0("PC2: ",PoV[2],"% variance")) +
  theme_classic()

```


```{r upset_LPS1, echo=TRUE}
list_filters = list(LPS = genes_LPS, IFNy = genes_IFNy)
#list_filters
overlap = unlist(list_filters)
```

### Upset plot
 
```{r upset_LPS2, echo=TRUE}
upset(fromList(list_filters), order.by = "freq")
```

# DEG across stimulations FDR 5% in 2 stimulations 
#348 genes 

```{r load tables1, echo = TRUE, results = 'hide'}
DEG_LPS <- subset(res_name_LPS2, adj.P.Val < 0.05)
DEG_IFNy <- subset(res_name_IFNy2, adj.P.Val < 0.05)

genes_LPS <- DEG_LPS$ensembl
genes_IFNy <- DEG_IFNy$ensembl

# create a vector with all genes listed
x <- c(genes_LPS, genes_IFNy)
# make a dataframe
as.data.frame(x) #12590 genes 
# check if there are genes duplicate 
duplicated(x) 
x[duplicated(x)]
unique(x)
#continue only with duplicated
ensembl <- x[duplicated(x)] 

genelist <- as.data.frame(ensembl)
#remove duplicated
ensembl <- genelist$ensembl[!duplicated(genelist$ensembl)]
genelist <- as.data.frame(ensembl) #3159 genes

DEG_genes_LPS = merge(res_name_LPS2, genelist, by ="ensembl")
DEG_genes_IFNy = merge(res_name_IFNy2, genelist, by = "ensembl")

Expression_stimulations <- cbind(DEG_genes_LPS, DEG_genes_IFNy)

as.data.frame(Expression_stimulations) 
# changes names in data frame
names(Expression_stimulations)[2] <- "LPS"
names(Expression_stimulations)[10] <- "IFNy"

DF_log2FC <- Expression_stimulations %>% select(1, 2, 10)
DF <- as.data.frame(DF_log2FC)
DF <- t(DF)
head( DF )[,1:10]

DF_log2FC <- merge(gencode, DF_log2FC, by = "ensembl")
DF_log2FC$symbol = DF_log2FC$GeneSymbol

# fill NA with 0
DF_log2FC[is.na(DF_log2FC)] = 0
save(DF_log2FC, file = "kmeans348gsupport_IFNy_LPS.Rdata")

# remove the first column
df_num = as.matrix(DF_log2FC[,4:5])
# scale the data 
df_num_scale = scale(df_num)
# make rownames as function
rownames(df_num) = sapply(DF_log2FC$symbol,function(x) 
strsplit(as.character(x),split = "\\\\")[[1]][1])
save(df_num, file = "df_num_FDR0.53stim_ensembl_LPS_IFNy.Rdata")
save(df_num_scale, file = "df_num_scale_FDR0.52stim_ensembl_LPS_IFNy.Rdata")
```

### Overlap 348 genes with Patir

```{r overlap2, echo=TRUE}
gencode <- read.delim("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
gencode$ensembl = gencode$gene_id
genelist <- merge(gencode, genelist, by = "ensembl")
genelist$symbol = genelist$GeneSymbol
Patir <- read.csv("~/Documents/MiGASti/Databases/Patir.txt", sep="")
names(Patir)[1] <- "symbol"
overlap2 <- merge(Patir, genelist, by = "symbol")
list(overlap2$symbol)
```

### Heatmap DEG
#348 genes

```{r heatmap_cor_3161, echo=TRUE}
rownames(DF_log2FC) = DF_log2FC$symbol
genelist <- c("ADAM28",   "ATP8B4",   "C1QB",    "CASP1",   "CD53",     "CD69", "CD86",   "CLEC17A",  "CSF1R",    "CTSC",  "FCGR1A",   "FGL2", "GPR183",   "HAMP",  "HLA-DPA1", "HLA-DQB1", "HLA-DRA",  "HLA-DRB1", "OLR1", "PTPN6",    "PTPRC", "SIGLEC9",  "SP140", "SP140L", "TFEC", "TNFRSF1B", "TRIM22")

labels <- rownames(DF_log2FC)
labels[!labels %in% genelist] <- "" 
pheatmap(df_num_scale, labels_row = labels, main = "Differentially expressed genes in the two stimulations", fontsize_row = 6)

```

### Correlation between logFC LPS and IFNy (Spearman)
#348 genes

```{r corrplot_3161, echo=TRUE, message=FALSE}
df <- as.data.frame(df_num_scale)
res <- cor.test(df$LPS, df$IFNy, method = "spearman")
res
```

### PCA across stimulations

```{r PCA_stim, echo=TRUE, message=FALSE}
dat <- data.frame(df_num_scale)
stimulation <- colnames(dat)
as.data.frame(stimulation)
dat <- data.frame(dat)
pca <- prcomp(t(dat), scale=TRUE)
PoV <- round((pca$sdev^2/sum(pca$sdev^2))*100)
#Grab covariates to plot:
library(ggplot2)
library(ggfortify)


#Plot:
ggplot(pca, aes(PC1, PC2, color=stimulation)) +
  geom_point(size=2) +
  xlab(paste0("PC1: ",PoV[1],"% variance")) +
  ylab(paste0("PC2: ",PoV[2],"% variance")) +
  theme_classic()
```


```{r upset_LP4, echo=TRUE}
list_filters = list(LPS = DEG_genes_LPS$symbol, IFNy = DEG_genes_IFNy$symbol)
#list_filters
overlap = unlist(list_filters)
```

### Upset plot
 
```{r upset_LPS3, echo=TRUE}
upset(fromList(list_filters), order.by = "freq")
```

# DEG across stimulations FDR 5%
#6571 genes 

```{r load tables20, echo = TRUE, results = 'hide'}
#check duplicated genes in list 
#[1] "ATF7"    "TMSB15B" "POLR2J4" "MATR3"   "HSPA14"  "SCO2" "TBCE"    "ATXN7"   "ABCF2"   "POLR2J3" "GGT1"  

DEG_LPS <- subset(res_name_LPS2, adj.P.Val < 0.05)
DEG_IFNy <- subset(res_name_IFNy2, adj.P.Val < 0.05)

genes_LPS <- DEG_LPS$ensembl
genes_IFNy <- DEG_IFNy$ensembl

# create a vector with all genes listed
x <- c(genes_LPS, genes_IFNy)
# make a dataframe
as.data.frame(x) #12590 genes 
# check if there are genes duplicate 
duplicated(x) 
unique(x)
#remove duplicated
ensembl <- x[!duplicated(x)] 

genelist <- as.data.frame(ensembl)
ensembl <- genelist$ensembl[!duplicated(genelist$ensembl)]
genelist <- as.data.frame(ensembl)
genelist %>% distinct(ensembl, .keep_all = TRUE)

DEG_genes_LPS = merge(res_name_LPS2, genelist, by ="ensembl")
DEG_genes_IFNy = merge(res_name_IFNy2, genelist, by = "ensembl")

Expression_stimulations <- cbind(DEG_genes_LPS, DEG_genes_IFNy)

as.data.frame(Expression_stimulations) 
# changes names in data frame
names(Expression_stimulations)[2] <- "LPS"
names(Expression_stimulations)[10] <- "IFNy"

DF_log2FC <- Expression_stimulations %>% select(1, 2, 10)
DF <- as.data.frame(DF_log2FC)
DF <- t(DF)
head( DF )[,1:10]


# fill NA with 0
DF_log2FC[is.na(DF_log2FC)] = 0
save(DF_log2FC, file = "kmeans8523ensembl_LPS_IFNy.Rdata")

DF_log2FC <- merge(gencode, DF_log2FC, by = "ensembl")
DF_log2FC$symbol = DF_log2FC$GeneSymbol

#remove rownames that are duplicates (HSPA14)
DF_log2FC = DF_log2FC[-c(5404),]


# remove the first column
df_num = as.matrix(DF_log2FC[,4:5])
# scale the data 
df_num_scale = scale(df_num)
# make rownames as function
rownames(df_num) = sapply(DF_log2FC$symbol,function(x) 
strsplit(as.character(x),split = "\\\\")[[1]][1])
save(df_num, file = "df_num_FDR0.5stim_allgenes_ensembl_LPS_IFNy.Rdata")
save(df_num_scale, file = "df_num_scale_FDR0.5stim_allgenes_ensembl_LPS_IFNy.Rdata")
```

### Overlap 6571 genes with Patir

```{r overlap_patir, echo=TRUE}
gencode <- read.delim("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
gencode$ensembl = gencode$gene_id
genelist <- merge(gencode, genelist, by = "ensembl")
genelist$symbol = genelist$GeneSymbol
Patir <- read.csv("~/Documents/MiGASti/Databases/Patir.txt", sep="")
names(Patir)[1] <- "symbol"
overlap2 <- merge(Patir, genelist, by = "symbol")
list(overlap2$symbol)
```

### Heatmap DEG
#6571 genes

```{r heatmap_DEG_all, echo=TRUE}
rownames(DF_log2FC) = DF_log2FC$symbol
genelist <- c("RGS1", "TLR6", "TYROBP", "TRIM22", "CD86", "TLR3"   )

labels <- rownames(DF_log2FC)
labels[!labels %in% genelist] <- "" 
pheatmap(df_num_scale, labels_row = labels, main = "Differentially expressed genes", fontsize_row =6)
```


### Correlation logFC between LPS and IFNy (spearman)
#6571 genes

```{r corr_all, echo=TRUE, message=FALSE}
df <- as.data.frame(df_num_scale)
res <- cor.test(df$LPS, df$IFNy, method = "spearman")
res
```


### PCA across stimulations

```{r PCA_all, echo=TRUE, message=FALSE}
dat <- data.frame(df_num_scale)
stimulation <- colnames(dat)
as.data.frame(stimulation)
dat <- data.frame(dat)
pca <- prcomp(t(dat), scale=TRUE)
PoV <- round((pca$sdev^2/sum(pca$sdev^2))*100)
#Grab covariates to plot:
library(ggplot2)
library(ggfortify)


#Plot:
ggplot(pca, aes(PC1, PC2, color=stimulation)) +
  geom_point(size=2) +
  xlab(paste0("PC1: ",PoV[1],"% variance")) +
  ylab(paste0("PC2: ",PoV[2],"% variance")) +
  theme_classic()

```


```{r upset_LPS_all, echo=TRUE}
list_filters = list(LPS = genes_LPS, IFNy = genes_IFNy)
#list_filters
overlap = unlist(list_filters)
```

### Upset plot
 
```{r upset_LPS2_all, echo=TRUE}
upset(fromList(list_filters), order.by = "freq")
```


# Scatterplots 
### LogFC LPS vs IFNy 
#18981 genes 

```{r scat_LPS_ifny, echo=TRUE}
names(res_name_LPS2)[2] <- "LPS"
names(res_name_IFNy2)[2] <- "IFNy"
DEG_genes_LPS_IFNy = merge(res_name_LPS2, res_name_IFNy2, by ="symbol")

LPS <- DEG_genes_LPS_IFNy[,3]
IFNy <- DEG_genes_LPS_IFNy[,12]
df = data.frame(LPS, IFNy)
p <- ggplot(df, aes(x=LPS, y=IFNy) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  scale_fill_continuous(type = "viridis") +
  theme_bw()
p + stat_cor(method = "spearman")
```




