---
title: "Metadata"
author: "Gijsje"
date: "2/10/2021"
output: 
rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---

```{r load.packages, echo=FALSE, message=FALSE, results='hide'}
library(readxl)
library(knitr)
library(kableExtra)
library(ggplot2)
library(RColorBrewer)


knitr::opts_chunk$set( 
  warning=FALSE,
  message=FALSE,
  results = 'asis',
  error = FALSE,
  tidy = FALSE,
  fig.show = "hold")
```

# Preparing metadata
### Create metadata file 

```{r load_file, echo=FALSE}
metadata = "~/Documents/MiGASti/Old versions databases/Metadata.xlsx"
metadata = read_excel(metadata, col_names = TRUE) 
metadata = as.data.frame(metadata)
setwd("~/Documents/MiGASti/Databases")
general_stats <- read.table("multiqc_general_stats_533s.txt", header=T, stringsAsFactors = FALSE)

metadata <- merge(metadata, general_stats, by = "Sample")
#set rownames to Sample
row.names(metadata) <- metadata$Sample 
```

# Number of donors before QC

```{r load_file2, echo=FALSE}
#remove uncultured samples
metadata_cultured <- metadata[metadata$Stimulation != "ununstim",]
#check numbers
length(unique(metadata_cultured$Donor_id)) 
```

### Change names

```{r metadata1, echo=FALSE, results = 'hide'}
# change names in metadata and set to correct data types 
names(metadata)[names(metadata) == "Picard_mqc.generalstats.picard.PCT_MRNA_BASES"] <- "picard_pct_mrna_bases"
names(metadata)[names(metadata) == "Picard_mqc.generalstats.picard.PERCENT_DUPLICATION"] <- "picard_pct_percent_duplication"
names(metadata)[names(metadata) == "Picard_mqc.generalstats.picard.summed_mean"] <- "picard_summed_mean"
names(metadata)[names(metadata) == "Picard_mqc.generalstats.picard.summed_median"] <- "picard_summed_median"
names(metadata)[names(metadata) == "Picard_mqc.generalstats.picard.PCT_RIBOSOMAL_BASES"] <- "picard_pct_ribosomal_bases"
names(metadata)[names(metadata) == "Picard_mqc.generalstats.picard.PCT_PF_READS_ALIGNED"] <- "picard_pct_pf_reads_aligned"
names(metadata)[names(metadata) == "Picard_mqc.generalstats.picard.PCT_PF_READS_ALIGNED"] <- "picard_pct_pf_reads_aligned"
names(metadata)[names(metadata) == "featureCounts_mqc.generalstats.featurecounts.Assigned"] <- "featurecounts_assigned"
names(metadata)[names(metadata) == "STAR_mqc.generalstats.star.uniquely_mapped"] <- "star_uniquely_mapped"
names(metadata)[names(metadata) == "STAR_mqc.generalstats.star.uniquely_mapped_percent"] <- "star_uniquely_mapped_percent"
names(metadata)[names(metadata) == "STAR_mqc.generalstats.star.total_reads"] <- "star_total_reads"
names(metadata)[names(metadata) == "Trimmomatic_mqc.generalstats.trimmomatic.dropped_pct"] <- "trimmomatic_dropped_pct"
names(metadata)[names(metadata) == "Picard_mqc.generalstats.picard.PCT_INTERGENIC_BASES"] <- "picard_pct_intergenic_bases"
names(metadata)[names(metadata) == "Picard_mqc.generalstats.picard.PCT_INTRONIC_BASES"] <- "picard_pct_intronic_bases"
```

### Change to correct data format

```{r metadata2, echo=FALSE, results = 'hide'}
metadata$picard_pct_mrna_bases <- as.numeric(metadata$picard_pct_mrna_bases)
length(metadata$picard_pct_mrna_bases)

metadata$picard_pct_percent_duplication <- as.numeric(metadata$picard_pct_percent_duplication)
length(metadata$picard_pct_percent_duplication)

metadata$picard_summed_mean <- as.numeric(metadata$picard_summed_mean)
length(metadata$picard_summed_mean)

metadata$picard_summed_median <- as.numeric(metadata$picard_summed_median)
length(metadata$picard_summed_median)

metadata$picard_pct_ribosomal_bases <- as.numeric(metadata$picard_pct_ribosomal_bases)
length(metadata$picard_pct_ribosomal_bases)

metadata$picard_pct_pf_reads_aligned <- as.numeric(metadata$picard_pct_pf_reads_aligned)
length(metadata$picard_pct_pf_reads_algined)

metadata$featurecounts_assigned <- as.numeric(metadata$featurecounts_assigned)
length(metadata$featurecounts_assigned)

metadata$star_uniquely_mapped <- as.numeric(metadata$star_uniquely_mapped)
length(metadata$star_uniquely_mapped)

metadata$star_uniquely_mapped_percent <- as.numeric(metadata$star_uniquely_mapped_percent)
length (metadata$star_uniquely_mapped_percent)

metadata$star_total_reads <- as.numeric(metadata$star_total_reads)
length(metadata$star_total_reads)

metadata$trimmomatic_dropped_pct <- as.numeric(metadata$trimmomatic_dropped_pct)
length(metadata$trimmomatic_dropped_pct)

metadata$picard_pct_ribosomal_bases <- as.numeric(metadata$picard_pct_ribosomal_bases)
length(metadata$picard_pct_ribosomal_bases)

metadata$picard_pct_intronic_bases <- as.numeric(metadata$picard_pct_intronic_bases)
length(metadata$picard_pct_intronic_bases)

metadata$picard_pct_intergenic_bases <- as.numeric(metadata$picard_pct_intergenic_bases)
length(metadata$picard_pct_intergenic_bases)

metadata$featurecounts_assigned <- as.numeric(metadata$featurecounts_assigned)

metadata$Region <- as.factor(metadata$Region)
metadata$Stimulation <- as.factor(metadata$Stimulation)
metadata$Average_library_size <- as.numeric(metadata$Average_library_size)
metadata$Qubit_conc <- as.numeric(metadata$Qubit_conc)
metadata$Diagnosis <- as.factor (metadata$Diagnosis)
metadata$Main_diagnosis <- as.factor (metadata$Main_diagnosis)
```

### QC: name samples TRUE or FALSE (outlier = TRUE) based on QC filtering 

```{r metadata_QC, echo=FALSE, results = 'hide'}
setwd("~/Documents/MiGASti/Databases")
exclude <- read.table("~/Documents/MiGASti/Databases/samples2remove.txt")
samplenames <- exclude
metadata$outlier <- NA
metadata[metadata$Sample %in% samplenames$x,]$outlier <- "TRUE"
metadata[!metadata$Sample %in% samplenames$x,]$outlier <- "FALSE"
#write.table(metadata, "metadata.txt")
```

# Number of unique donors of all cultured + stimulated samples 

```{r metadata_QC2, echo=TRUE}
#set rownames to Sample
row.names(metadata) <- metadata$Sample 
exclude <- exclude$x
metadata_filt = metadata[ !(rownames(metadata) %in% exclude), ]
#remove uncultured samples
metadata_cultured <- metadata_filt[metadata_filt$Stimulation != "ununstim",]
#check numbers
length(unique(metadata_cultured$Donor_id)) 
```

# Number of stimulations 

```{r metadata_stims1, echo=TRUE}
metadata_filt <- metadata_cultured
length(unique(metadata_filt$Stimulation))
```

### Number of samples per stimulation

```{r metadata_sample_stim, echo=TRUE}
as.data.frame(t(as.matrix(unclass(table(metadata_filt$Stimulation, useNA = "ifany")  )))) %>%
kable(row.names = F) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

### Number of donors per stimulation

```{r metadata_sample_stims2, echo=TRUE}
donor_stimuli = unique(metadata_filt[,c("Donor_id", "Stimulation")])

as.data.frame(t(as.matrix(unclass(  table(donor_stimuli$Stimulation, useNA = "ifany")  )))) %>%
kable(row.names = F) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

### Number of samples by region

```{r metadata_sample_stim20, echo=TRUE}
as.data.frame(t(as.matrix(unclass(table(metadata_filt$Region, useNA = "ifany")  )))) %>%
kable(row.names = F) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

### Number of donors by region 

```{r metadata_sample_stim3, echo=TRUE}
donor_region = unique(metadata_filt[,c("Donor_id", "Region")])
as.data.frame(t(as.matrix(unclass(table(donor_region$Region, useNA = "ifany")  )))) %>%
kable(row.names = F) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```


# Number of donors per sex

```{r metadata_sample_stim7, echo=TRUE}
sexByDonor = unique(metadata_filt[,c("Donor_id", "sex")])
#createDT(sexByDonor)
as.data.frame(t(as.matrix(unclass(table(sexByDonor$sex, useNA = "ifany"))))) %>%
 kable(row.names = F) %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```


# Age
### Maximum age 

```{r metadata_sample_stim8, echo=TRUE}
ageByDonor = unique(metadata_filt[,c("Donor_id", "age", "sex")])
#createDT(ageByDonor)
max(ageByDonor$age, na.rm = T)
```

### Minimum age

```{r metadata_sample_stim21, echo=TRUE}
min(ageByDonor$age, na.rm = T)
```

### Distribution plot age/sex by donor 

```{r metadata_sample_stim10, echo=TRUE}
mean_f = mean(ageByDonor[ageByDonor$sex == "f", "age"], na.rm = T)
mean_m = mean(ageByDonor[ageByDonor$sex == "m", "age"], na.rm = T)

ggplot(ageByDonor, aes(x=age, fill=sex)) +
  geom_histogram(bins = 25, colour='black', position = "stack") +
  labs(x="Age", y="Donors") +
  scale_y_continuous(breaks = (1:20)) +
  scale_x_continuous(breaks=seq(20,120,10)) + 
  geom_vline(xintercept=mean_f, color = "red", linetype="dashed") +
  geom_vline(xintercept=mean_m, color = "blue", linetype="dashed") +
  theme_classic()
```

# Main diagnosis by donor 

```{r metadata_sample_stim24, echo=TRUE, results = 'hide'}
Main_diagnosisByDonnor = unique(metadata_filt[,c("Donor_id", "Main_diagnosis")])
table(Main_diagnosisByDonnor)
```

### Number of different main diagnosis 

```{r metadata_sample_stim9, echo=TRUE}
length(unique(metadata$Main_diagnosis))
```

### Number of different main diagnosis 

```{r metadata_sample_stim11, echo=TRUE}
Main_diagnosisByDonor = unique(metadata_filt[,c("Donor_id", "Main_diagnosis")])
as.data.frame(t(as.matrix(unclass(table(Main_diagnosisByDonor$Main_diagnosis, useNA = "ifany")  )))) %>%
 kable(row.names = F) %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

### Pie chart main_diagnosis 

```{r metadata_sample_stim12, echo=TRUE}
pie_table = table(Main_diagnosisByDonor$Main_diagnosis)

myPalette <- brewer.pal(10, "Set3") 
pie(pie_table, border = "white", col = myPalette) +
  coord_polar("y", start = 0) 
```

### Main diagnosis by sample

```{r metadata_sample_stim13, echo=TRUE}
as.data.frame(table(metadata_filt$Main_diagnosis), useNA = "ifany"  ) %>%
 kable(row.names = F) %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

### Samples by main diagnosis and region 

```{r metadata_sample_stim14, echo=TRUE}
as.data.frame(table(metadata_filt[,c("Main_diagnosis", "Region")]), useNA = "ifany"  ) %>% tidyr::spread(Region, Freq) %>%
 kable(row.names = F) %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

### Samples by stimulation and region 

```{r metadata_sample_stim100, echo=TRUE}
as.data.frame(table(metadata_filt[,c("Stimulation", "Region")]), useNA = "ifany"  ) %>% tidyr::spread(Region, Freq) %>%
 kable(row.names = F) %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

# Cause of death 

### Categories

```{r metadata_sample_stim15, echo=TRUE}
DonorByDeath = unique(metadata_filt[,c("Donor_id", "Cause_of_death_categories")])
#createDT(DonorByDeath)
as.data.frame(t(as.matrix(unclass(  table(DonorByDeath$Cause_of_death_categories, useNA = "ifany")  )))) %>%
 kable(row.names = F) %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

### Chart

```{r metadata_sample_stim16, echo=TRUE}
df = as.data.frame(table(DonorByDeath$Cause_of_death_categories, useNA = "ifany"))
colnames(df) <- c("Cause_of_death", "Frequency")

ggplot(data=df, aes(x=Cause_of_death, y=Frequency)) +
  geom_bar(stat="identity", fill="#FA8072")+
  geom_text(aes(label=Frequency), vjust=0.5, hjust=1.6, color="white", size=3.5)+
  coord_flip() + 
  theme_classic()
```


# Number of unique donors of all cultured + stimulated samples  (minus TNFa)

```{r metadata_QC4, echo=TRUE}
#remove uncultured samples
metadata_cultured2 <- metadata_cultured[metadata_cultured$Stimulation != "TNFa",]
#check numbers
length(unique(metadata_cultured2$Donor_id)) 
```




