---
title: "PCA before filtering"
author: "Gijsje"
date: "2/9/2021"
output: 
rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---

```{r load.packages, echo=FALSE, message=FALSE, results='hide'}
library(devtools)
library(factoextra)
library(ggfortify)
library(ggplot2)
library(dplyr)
library(broom)
library(variancePartition)
library(pheatmap)
library(edgeR)
library(limma)
library(DESeq2)
library(Hmisc)
library(matrixStats)


knitr::opts_chunk$set( 
  warning=FALSE,
  message=FALSE,
  results = 'asis',
  error = FALSE,
  tidy = FALSE,
  fig.show = "hold")
```


```{r load_data, echo=TRUE, results = 'hide'}
load("~/Documents/MiGASti/Databases/gene_matrix.RData")
metadata <- read.table("~/Documents/MiGASti/Databases/metadata.txt")
#set rownames to Sample
row.names(metadata) <- metadata$Sample 
#remove low count genes
cpm <- cpm(genes_counts) 
# CPM >= 1 in at least 50% of the samples
keep.exp <- rowSums(cpm > 1) >= (0.5 * ncol(genes_counts) )
genes_counts5 <- genes_counts[ keep.exp, ] #18625 genes 
#performing voom normalisation on data
counts_voom <- limma::voom(genes_counts5)
genes_counts_voom <- counts_voom$E
#order metadata and genes counts
#rownames(metadata)
#colnames(genes_counts_voom)
genes_counts_ordered <- genes_counts_voom[,rownames(metadata)]
#head(genes_counts_ordered)
all(rownames(metadata) == colnames (genes_counts_ordered)) #TRUE
```

# PCAs
### PCA no scaling and centering of all 533 samples 

```{r PCA_noscale, echo=TRUE}
res.pca = prcomp(t(genes_counts_ordered))
autoplot(res.pca, colour = "Stimulation", data = metadata)
```

### PCA with scaling and centering of all 533 samples (also uncultured samples)

```{r PCA_scale, echo=TRUE}
pca = prcomp(t(genes_counts_ordered), scale. = TRUE, center = TRUE)
autoplot(pca, colour = "Stimulation", data = metadata)
```

### PCA with scaling and centering + labeling of all 533 samples

```{r PCA_labels, echo=TRUE}
pca = prcomp(t(genes_counts_ordered), scale. = TRUE, center = TRUE)
autoplot(pca, data= metadata, colour = 'Stimulation', shape = FALSE, label.size = 2)
```

### Variances per PC of all samples 

```{r variances, echo=TRUE}
plot(pca)
```


# Boxplots of all samples
### PC1 

```{r boxplots_PC1, echo=TRUE}
PC1 <- pca$x[,1]
Stimulation <- metadata$Stimulation
df = data.frame(Stimulation, PC1)
ggplot(data = df, mapping = aes(x = Stimulation, y = PC1)) +
geom_boxplot()
```

### PC2  

```{r boxplots PC2, echo=TRUE}
PC2 <- pca$x[,2]
Stimulation <- metadata$Stimulation
df = data.frame(Stimulation, PC2)
ggplot(data = df, mapping = aes(x = Stimulation, y = PC2)) +
geom_boxplot()
```


# PCAs of cultured samples
### PCA of cultured samples with centering and scaling 

```{r PCA_cultured, echo=TRUE}
#label ununstim samples in metadata
metadata_cultured <- metadata[metadata$Stimulation != "ununstim",]
#check numbers
dim(metadata_cultured)
#check numbers per stimulation
table(metadata$Stimulation)
#remove samples in genes counts datafile 
genes_counts_cultured <- genes_counts_ordered[,metadata_cultured$Sample]
pca = prcomp(t(genes_counts_cultured), scale. = TRUE, center = TRUE)
autoplot(pca, colour = "Stimulation", data = metadata_cultured)
```


### PCA of cultured samples with labeling of samples 

```{r PCA, labeling, echo=TRUE}
autoplot(pca, data= metadata_cultured, colour = 'Stimulation', shape = FALSE, label.size = 2)
```

### Variances per PC of cultured samples 

```{r variances_cultured, echo=TRUE}
plot(pca)
```

# Boxplots of cultured samples 
### PC1

```{r boxplots_1PC, echo=TRUE}
PC1 <- pca$x[,1]
Stimulation <- metadata_cultured$Stimulation
df = data.frame(Stimulation, PC1)
ggplot(data = df, mapping = aes(x = Stimulation, y = PC1)) +
  geom_boxplot()
```

### PC2 

```{r boxplots_PC2, echo=TRUE}
PC2 <- pca$x[,2]
Stimulation <- metadata_cultured$Stimulation
df = data.frame(Stimulation, PC2)
ggplot(data = df, mapping = aes(x = Stimulation, y = PC2)) +
  geom_boxplot()
```

#### PC3

```{r boxplots_PC3, echo=TRUE}
PC3 <- pca$x[,3]
Stimulation <- metadata_cultured$Stimulation
df = data.frame(Stimulation, PC3)
ggplot(data = df, mapping = aes(x = Stimulation, y = PC3)) +
  geom_boxplot()
```


### PC5 

```{r boxplots_PC5, echo=TRUE}
PC5 <- pca$x[,5]
Stimulation <- metadata_cultured$Stimulation
df = data.frame(Stimulation, PC5)
ggplot(data = df, mapping = aes(x = Stimulation, y = PC5)) +
  geom_boxplot()
```

### PC7

```{r boxplots_PC7, echo=TRUE}
PC7 <- pca$x[,7]
Stimulation <- metadata_cultured$Stimulation
df = data.frame(Stimulation, PC7)
ggplot(data = df, mapping = aes(x = Stimulation, y = PC7)) +
  geom_boxplot()
```

# Scatterplot PC1 + PC2 

```{r scatterplot, echo=TRUE}
PCdf <- data.frame(sample = rownames(pca$x), pc1 = pca$x[,1], pc2 = pca$x[,2], Stimulation = metadata_cultured$Stimulation)
ggplot(data = PCdf, mapping = aes(x = PC1, y = PC2, colour = Stimulation)) +
  geom_point()
```
