---
title: "DREAM_monocytes"
author: "Gijsje"
date: "5/13/2021"
output: 
 rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---

```{r setup, include=FALSE}
library('variancePartition')
library('edgeR')
library(dplyr)
library(tidyr)
library(readxl)
library(edgeR)
library(limma)
library(DESeq2)
library(tidyverse)
library(data.table)
library(shiny)
library(pheatmap)
library(ggfortify)
library(ggplot2)
library(UpSetR)
library(VennDiagram)
library(knitr) 
library(tidyverse)
library(dplyr)
library(tidyr)
library(tibble)
library(purrr)
library(corrplot)
library(ggplot2)
library(ggpubr)
```

```{r Helper Functions, echo = FALSE}
createDT <- function(DF, caption="", scrollY=500){
data <- DT::datatable(DF, caption=caption,
extensions = 'Buttons',
options = list( dom = 'Bfrtip',
buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
scrollY = scrollY, scrollX=T, scrollCollapse = T, paging = F,
columnDefs = list(list(className = 'dt-center', targets = "_all"))
)
)
return(data)
}
```

# Data load

```{r kable, echo=FALSE, message=FALSE, warnings=FALSE, results='hide'}
#load gene matrix
load("~/Downloads/gene_matrix_monocytes.RData")
#import data metadata
merged_meta_mono <- read.csv("~/Downloads/merged_meta_mono.csv", header=FALSE)
colnames(merged_meta_mono) <- merged_meta_mono[1,]
merged_meta_mono <- merged_meta_mono[-1,]
rownames(merged_meta_mono) = merged_meta_mono$sample_id
#remove low count genes
cpm <- cpm(genes_counts)
# CPM >= 1 in at least 30% of the samples
keep.exp <- rowSums(cpm > 1) >= (0.3 * ncol(genes_counts))
genes_counts_filt1 <- genes_counts[ keep.exp, ] 
#order metadata and genes counts
genes_counts_ordered <- genes_counts_filt1[,rownames(merged_meta_mono)]

merged_meta_mono$PCT_CODING_BASES <- as.numeric(merged_meta_mono$PCT_CODING_BASES)
merged_meta_mono$MEDIAN_3PRIME_BIAS <- as.numeric(merged_meta_mono$MEDIAN_3PRIME_BIAS)
merged_meta_mono$gender <- as.factor(merged_meta_mono$gender)
merged_meta_mono$RNA_cc <- as.numeric(merged_meta_mono$RNA_cc)
merged_meta_mono$age <- as.numeric(merged_meta_mono$age)
merged_meta_mono$C1 <- as.numeric(merged_meta_mono$C1)
merged_meta_mono$C2<- as.numeric(merged_meta_mono$C2)
merged_meta_mono$C3 <- as.numeric(merged_meta_mono$C3)
merged_meta_mono$C4 <- as.numeric(merged_meta_mono$C4)
merged_meta_mono$PCT_INTERGENIC_BASES <- as.numeric(merged_meta_mono$PCT_INTERGENIC_BASES)
merged_meta_mono$treatment <- as.factor(merged_meta_mono$treatment)
merged_meta_mono$donor <- as.factor(merged_meta_mono$donor)

form <- ~ 0 + treatment + (1 | donor) + PCT_CODING_BASES + MEDIAN_3PRIME_BIAS + gender + RNA_cc + age + C1 + C2 + C3 + C4 + PCT_INTERGENIC_BASES 

# estimate weights using linear mixed model of dream
#vobjDream = voomWithDreamWeights ( genes_counts_ordered, form, merged_meta_mono)
```

# DREAM model 

```{r setup2, echo=TRUE}
#L1 = getContrast( vobjDream, form, merged_meta_mono, c("treatmentLPS", "treatmentBasal"))
#L2 = getContrast( vobjDream, form, merged_meta_mono, c("treatmentIFNg", "treatmentBasal"))

#L = cbind(L1, L2)
#fit = dream(vobjDream, form, merged_meta_mono, L)
#res_LPS_mono <- data.frame(topTable(fit, coef='L1', number=nrow(genes_counts_filt1), sort.by = "p"), check.names = F)
#res_IFNy_mono <- data.frame(topTable(fit, coef='L2', number=nrow(genes_counts_filt1), sort.by = "p"), check.names = F)
gencode_30 = read.table("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
colnames(gencode_30) = c("ensembl","symbol")
#res_LPS_mono <- tibble::rownames_to_column(res_LPS_mono, "ensembl")
#res_LPS_monocytes <- merge(res_LPS_mono, gencode_30, by = "ensembl")
#res_IFNy_mono <- tibble::rownames_to_column(res_IFNy_mono, "ensembl")
#res_IFNy_monocytes <- merge(res_IFNy_mono, gencode_30, by = "ensembl")
#save(res_LPS_monocytes, file = "res_LPS_monocytes.Rdata")
#save(res_IFNy_monocytes, file = "res_IFNy_monocytes.Rdata")
```

# LPS

### DE genes 15% LPS 

```{r DE15, echo=TRUE}
load("~/Documents/MiGASti/Databases/res_LPS_monocytes.Rdata")
sign_LPS <- subset(res_LPS_monocytes, adj.P.Val < 0.15)
length(rownames(sign_LPS))
```

### DE genes 10% LPS 

```{r DE10, echo=TRUE}
sign_LPS <- subset(res_LPS_monocytes, adj.P.Val < 0.10)
length(rownames(sign_LPS))
```

### DE genes 5% LPS 

```{r DE5, echo=TRUE}
sign_LPS <- subset(res_LPS_monocytes, adj.P.Val < 0.05)
length(rownames(sign_LPS))
```

### FDR distibution

```{r FDRdistribution, echo = TRUE}
res = res_LPS_monocytes
p = ggplot(res, aes(P.Value))
p + geom_density(color="darkblue", fill="lightblue") +
theme_classic() +
ggtitle("FDR Distribution")
```

### Fold change distribution

```{r foldchangedistribution, echo = TRUE}
p = ggplot(res, aes(logFC))
p + geom_density(color = "darkblue", fill = "lightblue") +
theme_classic() +
ggtitle("Fold Change Distribution")
```

### MA plot 

```{r MAplot, echo = TRUE}
plot.data = res
plot.data$id = rownames(plot.data)
data = data.frame(plot.data)
data$P.Value = -log10(data$P.Value)
data$fifteen = as.factor(abs(data$adj.P.Val < 0.05))
ma = ggplot(data, aes(AveExpr, logFC, color = fifteen))
ma + geom_point() +
scale_color_manual(values = c("black", "red"), labels = c ("> 0.05", "< 0.05")) +
labs(title = "MA plot", color = "labels") +
theme_classic()
#theme(plot.title = element_text(hjust = 0.5)) + ylim (-10,10) + xlim(-4,22)
```


### Volcano plot 

```{r Volcano, echo = TRUE}
vp = ggplot(data, aes(logFC, P.Value, color = fifteen))
vp + geom_point() +
scale_color_manual(values = c("black", "red"), labels = c("> 0.05", "< 0.05")) +
labs(title = "Gene Level Volcano Plot", color = "FDR") +
#theme(plot.title = element_text(hjust = 0.5)) +
theme_classic() +
xlim(-5,5) + ylim(0, 30) + ylab("-log10 pvalue")
load("~/Documents/MiGASti/docs/res_name_LPS2.Rdata")
```


### Data table for download 

```{r download, echo = TRUE}
res_LPS_diff_top = res_LPS_monocytes[, c("ensembl", "symbol", "logFC", "AveExpr", "t", "P.Value", "adj.P.Val", "z.std")]
createDT(res_LPS_diff_top)
```

# IFNy

### DE genes 15% IFNy

```{r DE15_ifny, echo=TRUE}
load("~/Documents/MiGASti/Databases/res_IFNy_monocytes.Rdata")
sign_IFNy <- subset(res_LPS_monocytes, adj.P.Val < 0.15)
length(rownames(sign_IFNy))
```


### DE genes 10% IFNy 

```{r DE10IFNy, echo=TRUE}
sign_IFNy <- subset(res_IFNy_monocytes, adj.P.Val < 0.10)
length(rownames(sign_IFNy))
```


### DE genes 5% IFNy

```{r DE5_IFNy, echo=TRUE}
sign_IFNy <- subset(res_IFNy_monocytes, adj.P.Val < 0.05)
length(rownames(sign_IFNy))
```


### FDR distibution

```{r FDRdistribution_IFNy, echo = TRUE}
res = res_IFNy_monocytes
p = ggplot(res, aes(P.Value))
p + geom_density(color="darkblue", fill="lightblue") +
theme_classic() +
ggtitle("FDR Distribution")
```


### Fold change distribution

```{r foldchangedistribution_IFNy, echo = TRUE}
p = ggplot(res, aes(logFC))
p + geom_density(color = "darkblue", fill = "lightblue") +
theme_classic() +
ggtitle("Fold Change Distribution")
```


### MA plot 

```{r MAplot_IFNy, echo = TRUE}
plot.data = res
plot.data$id = rownames(plot.data)
data = data.frame(plot.data)
data$P.Value = -log10(data$P.Value)
data$fifteen = as.factor(abs(data$adj.P.Val < 0.05))
ma = ggplot(data, aes(AveExpr, logFC, color = fifteen))
ma + geom_point() +
scale_color_manual(values = c("black", "red"), labels = c ("> 0.05", "< 0.05")) +
labs(title = "MA plot", color = "labels") +
theme_classic()
#theme(plot.title = element_text(hjust = 0.5)) + ylim (-10,10) + xlim(-4,22)
```


### Volcano plot 

```{r Volcano_IFNy, echo = TRUE}
vp = ggplot(data, aes(logFC, P.Value, color = fifteen))
vp + geom_point() +
scale_color_manual(values = c("black", "red"), labels = c("> 0.05", "< 0.05")) +
labs(title = "Gene Level Volcano Plot", color = "FDR") +
#theme(plot.title = element_text(hjust = 0.5)) +
theme_classic() +
xlim(-5,5) + ylim(0, 30) + ylab("-log10 pvalue")
load("~/Documents/MiGASti/docs/res_name_LPS2.Rdata")
```

### Data table for download 

```{r download_IFNy, echo = TRUE}
res_IFNy_diff_top = res_IFNy_monocytes[, c("ensembl", "symbol", "logFC", "AveExpr", "t", "P.Value", "adj.P.Val", "z.std")]
createDT(res_IFNy_diff_top)
```

# Overlap with significant microglia LPS genes FDR 5%

```{r overlap_LPS_FDR5, echo = TRUE}
load("~/Documents/MiGASti/res_name_LPS2.Rdata")
sign_microglia_LPS <- subset(res_name_LPS2, adj.P.Val < 0.05)
sign_monocytes_LPS <- subset(res_LPS_monocytes, adj.P.Val < 0.05)
overlap <- merge(sign_microglia_LPS, sign_monocytes_LPS, by = "symbol")
```

```{r enriched_FDR5, echo = TRUE}
set1 <- sign_microglia_LPS
set2 <- sign_monocytes_LPS

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```

### Overlap with significant microglia IFNy genes FDR 5% 

```{r overlap_IFNy_FDR5, echo = TRUE}
load("~/Documents/MiGASti/docs/res_name_IFNy2.Rdata")
sign_microglia_IFNy <- subset(res_name_IFNy2, adj.P.Val < 0.05)
sign_monocytes_IFNy <- subset(res_IFNy_monocytes, adj.P.Val < 0.05)
overlap <- merge(sign_microglia_IFNy, sign_monocytes_IFNy, by = "symbol")
```

```{r enriched_IFNy_FDR5, echo = TRUE}
set1 <- sign_microglia_IFNy
set2 <- sign_monocytes_IFNy

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```

### Overlap with significant microglia LPS genes FDR 5% log FC > 1 microglia and monocytes > 5

```{r overlap_LPS_LF_sets2, echo = TRUE}
sign_microglia_LPS <- subset(res_name_LPS2, adj.P.Val < 0.05)
sign_microglia_LPS_LF <- subset(sign_microglia_LPS, logFC > 1)
sign_monocytes_LPS <- subset(res_LPS_monocytes, adj.P.Val < 0.05)
sign_monocytes_LPS_LF <- subset(sign_monocytes_LPS, logFC > 5)
overlap <- merge(sign_monocytes_LPS_LF, sign_monocytes_LPS_LF, by = "symbol")
```

```{r enriched_LPS_FC_sets, echo = TRUE}
set1 <- sign_microglia_LPS_LF
set2 <- sign_monocytes_LPS_LF

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```

### Overlap with significant microglia LPS genes FDR 5% log FC < -1 microglia and monocytes -5 

```{r overlap_LPS_LF_1, echo = TRUE}
sign_microglia_LPS <- subset(res_name_LPS2, adj.P.Val < 0.05)
sign_microglia_LPS_LF <- subset(sign_microglia_LPS, logFC < -1)
sign_monocytes_LPS <- subset(res_LPS_monocytes, adj.P.Val < 0.05)
sign_monocytes_LPS_LF <- subset(sign_monocytes_LPS, logFC < -5)
overlap <- merge(sign_monocytes_LPS_LF, sign_monocytes_LPS_LF, by = "symbol")
```

```{r enriched_LPS_FC_check, echo = TRUE}
set1 <- sign_microglia_LPS_LF
set2 <- sign_monocytes_LPS_LF

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```

### Overlap with significant microglia IFNy genes FDR 5% log FC > 1 and monocytes 5

```{r overlap_IFNy_LF, echo = TRUE}
sign_microglia_IFNy <- subset(res_name_IFNy2, adj.P.Val < 0.05)
sign_microglia_IFNy_LF <- subset(sign_microglia_IFNy, logFC > 1)
sign_monocytes_IFNy <- subset(res_IFNy_monocytes, adj.P.Val < 0.05)
sign_monocytes_IFNy_LF <- subset(sign_monocytes_IFNy, logFC > 5)
overlap <- merge(sign_monocytes_IFNy_LF, sign_monocytes_IFNy_LF, by = "symbol")
```

```{r enriched_LPS_FC222, echo = TRUE}
set1 <- sign_microglia_IFNy_LF
set2 <- sign_monocytes_IFNy_LF

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```

### Overlap with significant microglia IFNy genes FDR 5% log FC < -1 microglia and monocytes -5

```{r overlap_IFNy_LF_sets22340, echo = TRUE}
sign_microglia_IFNy <- subset(res_name_IFNy2, adj.P.Val < 0.05)
sign_microglia_IFNy_LF <- subset(sign_microglia_IFNy, logFC < -1)
sign_monocytes_IFNy <- subset(res_IFNy_monocytes, adj.P.Val < 0.05)
sign_monocytes_IFNy_LF <- subset(sign_monocytes_IFNy, logFC < -5)
overlap <- merge(sign_monocytes_IFNy_LF, sign_monocytes_IFNy_LF, by = "symbol")
```

```{r enriched_LPS_FC2_yes, echo = TRUE}
set1 <- sign_microglia_IFNy_LF
set2 <- sign_monocytes_IFNy_LF

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```

### Overlap with significant microglia LPS genes FDR 5% log FC > 1 + INFy monocytes > 5

```{r overlap_IFNy_LF_sets220, echo = TRUE}
sign_microglia_LPS <- subset(res_name_LPS2, adj.P.Val < 0.05)
sign_microglia_LPS_LF <- subset(sign_microglia_LPS, logFC > 1)
sign_monocytes_IFNy <- subset(res_IFNy_monocytes, adj.P.Val < 0.05)
sign_monocytes_IFNy_LF <- subset(sign_monocytes_IFNy, logFC > 5)
overlap <- merge(sign_microglia_LPS_LF, sign_monocytes_IFNy_LF, by = "symbol")
```


```{r enriched_LPS_IFNY_FC2_yes, echo = TRUE}
set1 <- sign_microglia_LPS_LF
set2 <- sign_monocytes_IFNy_LF

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```

### Overlap with significant microglia LPS genes FDR 5% log FC > 1 + monocytes < -5

```{r overlap_IFNy_LdddF_sets220, echo = TRUE}
sign_microglia_LPS <- subset(res_name_LPS2, adj.P.Val < 0.05)
sign_microglia_LPS_LF <- subset(sign_microglia_LPS, logFC > 1)
sign_monocytes_IFNy <- subset(res_IFNy_monocytes, adj.P.Val < 0.05)
sign_monocytes_IFNy_LF <- subset(sign_monocytes_IFNy, logFC < -5)
overlap <- merge(sign_microglia_LPS_LF, sign_monocytes_IFNy_LF, by = "symbol")
```

```{r enriched_LPS_IFNY_FC2_no, echo = TRUE}
set1 <- sign_microglia_LPS_LF
set2 <- sign_monocytes_IFNy_LF

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```


# Overlap with significant microglia IFNy genes FDR 5% log FC > 1 + LPS monocytes < -5

```{r overlap_IFNy_LF_sets2ss20_2, echo = TRUE}
sign_microglia_IFNy <- subset(res_name_IFNy2, adj.P.Val < 0.05)
sign_microglia_IFNy_LF <- subset(sign_microglia_IFNy, logFC > 1)
sign_monocytes_LPS <- subset(res_LPS_monocytes, adj.P.Val < 0.05)
sign_monocytes_LPS_LF <- subset(sign_monocytes_LPS, logFC < -5)
overlap <- merge(sign_microglia_IFNy_LF, sign_monocytes_LPS_LF, by = "symbol")
```


```{r enriched_LP22S_IFNY_FC2_yes_2, echo = TRUE}
set1 <- sign_microglia_IFNy_LF
set2 <- sign_monocytes_LPS_LF

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```

### Overlap with significant microglia LPS genes FDR 5% log FC > 1 + monocytes < -5

```{r overlap_IFNy_LF_sets220sxx, echo = TRUE}
sign_microglia_LPS <- subset(res_name_LPS2, adj.P.Val < 0.05)
sign_microglia_LPS_LF <- subset(sign_microglia_LPS, logFC > 1)
sign_monocytes_IFNy <- subset(res_IFNy_monocytes, adj.P.Val < 0.05)
sign_monocytes_IFNy_LF <- subset(sign_monocytes_IFNy, logFC < -5)
overlap <- merge(sign_microglia_IFNy_LF, sign_monocytes_IFNy_LF, by = "symbol")
```

```{r enriched_LPS_IFNY_FC2_yesxx, echo = TRUE}
set1 <- sign_microglia_LPS_LF
set2 <- sign_monocytes_IFNy_LF

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```

### Overlap with significant microglia LPS genes FDR 5% log FC > 1 + LPS monocytes < -5 

```{r overlap_IFNy_LF_sets2ss20, echo = TRUE}
sign_microglia_LPS <- subset(res_name_LPS2, adj.P.Val < 0.05)
sign_microglia_LPS_LF <- subset(sign_microglia_LPS, logFC > 1)
sign_monocytes_LPS <- subset(res_LPS_monocytes, adj.P.Val < 0.05)
sign_monocytes_LPS_LF <- subset(sign_monocytes_LPS, logFC < -5)
overlap <- merge(sign_microglia_LPS_LF, sign_monocytes_LPS_LF, by = "symbol")
```


```{r enriched_LP22S_LPS2_FC2_yes, echo = TRUE}
set1 <- sign_microglia_LPS_LF
set2 <- sign_monocytes_LPS_LF

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```


### Overlap with significant microglia LPS genes FDR 5% log FC < - 1 + LPS monocytes > 5 

```{r overlap_IFNy_LF_sets2ss20_sec, echo = TRUE}
sign_microglia_LPS <- subset(res_name_LPS2, adj.P.Val < 0.05)
sign_microglia_LPS_LF <- subset(sign_microglia_LPS, logFC < -1)
sign_monocytes_LPS <- subset(res_LPS_monocytes, adj.P.Val < 0.05)
sign_monocytes_LPS_LF <- subset(sign_monocytes_LPS, logFC > 5)
overlap <- merge(sign_microglia_LPS_LF, sign_monocytes_LPS_LF, by = "symbol")
```


```{r enriched_LP22S_LPS2_FC2_yes_sex, echo = TRUE}
set1 <- sign_microglia_LPS_LF
set2 <- sign_monocytes_LPS_LF

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```


### Overlap with significant microglia IFNy genes FDR 5% log FC > 1 + IFNy monocytes < -5 

```{r overlap_IFNy_LF_sets2ss20_third, echo = TRUE}
sign_microglia_IFNy <- subset(res_name_IFNy2, adj.P.Val < 0.05)
sign_microglia_IFNy_LF <- subset(sign_microglia_IFNy, logFC > 1)
sign_monocytes_IFNy <- subset(res_IFNy_monocytes, adj.P.Val < 0.05)
sign_monocytes_IFNy_LF <- subset(sign_monocytes_IFNy, logFC < -5)
overlap <- merge(sign_microglia_IFNy_LF, sign_monocytes_IFNy_LF, by = "symbol")
```


```{r enriched_LP22S_LPS2_FC2_yes_third, echo = TRUE}
set1 <- sign_microglia_IFNy_LF
set2 <- sign_monocytes_IFNy_LF

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```


# Overlap with significant microglia LPS genes FDR 5% log FC < - 1 + IFNy monocytes > 5

```{r overlap_IFNy_LF_sets2ss20_fourth_x_def2, echo = TRUE}
sign_microglia_LPS <- subset(res_name_LPS2, adj.P.Val < 0.05)
sign_microglia_LPS_LF <- subset(sign_microglia_LPS, logFC < -1)
sign_monocytes_IFNy <- subset(res_IFNy_monocytes, adj.P.Val < 0.05)
sign_monocytes_IFNy_LF <- subset(sign_monocytes_IFNy, logFC > 5)
overlap <- merge(sign_microglia_LPS_LF, sign_monocytes_IFNy_LF, by = "symbol")
```


```{r enriched_LP22S_LPS2_FC2_yes_fourth_x24, echo = TRUE}
set1 <- sign_microglia_LPS_LF
set2 <- sign_monocytes_IFNy_LF

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```

### Overlap with significant microglia LPS genes FDR 5% log FC < - 1 + IFNy monocytes < - 5 

```{r overlap_IFNy_LF_sets2ss20_fourth_x_def, echo = TRUE}
sign_microglia_LPS <- subset(res_name_LPS2, adj.P.Val < 0.05)
sign_microglia_LPS_LF <- subset(sign_microglia_LPS, logFC < -1)
sign_monocytes_IFNy <- subset(res_IFNy_monocytes, adj.P.Val < 0.05)
sign_monocytes_IFNy_LF <- subset(sign_monocytes_IFNy, logFC < - 5)
overlap <- merge(sign_microglia_LPS_LF, sign_monocytes_IFNy_LF, by = "symbol")
```


```{r enriched_LP22S_LPS2_FC2_yes_fourth_x2, echo = TRUE}
set1 <- sign_microglia_LPS_LF
set2 <- sign_monocytes_IFNy_LF

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```



# Scatterplot LogFC microglia vs monocytes LPS
#12812 genes 

```{r scatter_LPS, echo = TRUE}
genes = merge(res_name_LPS2, res_LPS_monocytes, by ="symbol")

microglia <- genes[,3]
monocytes <- genes[,10]
df = data.frame(microglia, monocytes)
p <- ggplot(df, aes(x=microglia, y=monocytes) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  scale_fill_continuous(type = "viridis") +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  theme_bw()
p + stat_cor(method = "spearman")
```

# Scatterplot LogFC microglia vs monocytes IFNy
#12812 genes 

```{r scatter_IFNy, echo = TRUE}
genes = merge(res_name_IFNy2, res_IFNy_monocytes, by ="symbol")

microglia <- genes[,3]
monocytes <- genes[,10]
df = data.frame(microglia, monocytes)
p <- ggplot(df, aes(x=microglia, y=monocytes) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  scale_fill_continuous(type = "viridis") +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  theme_bw()
p + stat_cor(method = "spearman")
```




