---
title: "DTU_genes_overlap_stimulation"
author: "Gijsje"
date: "5/11/2021"
output: 
 rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---

```{r setup, include=FALSE}
library('variancePartition')
library('edgeR')
library(dplyr)
library(tidyr)
library(readxl)
```

### Data load

```{r load data, echo = TRUE}
load("~/Documents/MiGASti/docs/2nd_pass/res_R848_name.Rdata")
load("~/Documents/MiGASti/docs/2nd_pass/res_LPS_name.Rdata")
load("~/Documents/MiGASti/docs/2nd_pass/res_IFNy_name.Rdata")
load("~/Documents/MiGASti/docs/2nd_pass/res_IL4_name.Rdata")
load("~/Documents/MiGASti/docs/2nd_pass/res_DEX_name.Rdata")
load("~/Documents/MiGASti/docs/2nd_pass/res_ATP_name.Rdata")

LPS_S = "~/Documents/MiGASti/Databases/LPS_DTU.xlsx"
LPS_S = read_excel(LPS_S, col_names = TRUE) 
LPS_S = as.data.frame(LPS_S)

TNFa_S = "~/Documents/MiGASti/Databases/TNFa_DTU.xlsx"
TNFa_S = read_excel(TNFa_S, col_names = TRUE) 
TNFa_S = as.data.frame(TNFa_S)

ATP_S = "~/Documents/MiGASti/Databases/ATP_DTU.xlsx"
ATP_S = read_excel(ATP_S, col_names = TRUE) 
ATP_S = as.data.frame(ATP_S)

IFNy_S = "~/Documents/MiGASti/Databases/DTU_IFNy.xlsx"
IFNy_S = read_excel(IFNy_S, col_names = TRUE) 
IFNy_S = as.data.frame(IFNy_S)

R848_S = "~/Documents/MiGASti/Databases/R848_DTU.xlsx"
R848_S = read_excel(R848_S, col_names = TRUE) 
R848_S = as.data.frame(R848_S)

IL4_S = "~/Documents/MiGASti/Databases/IL4_DTU.xlsx"
IL4_S = read_excel(IL4_S, col_names = TRUE) 
IL4_S = as.data.frame(IL4_S)

#subset genes FDR < 0.05
res_name_LPS2 <- subset(res_LPS_name, adj.P.Val < 0.05)
res_name_IFNy2 <- subset(res_IFNy_name, adj.P.Val < 0.05)
res_name_R8482 <- subset(res_R848_name, adj.P.Val < 0.05)
res_name_ATP2 <- subset(res_ATP_name, adj.P.Val < 0.05)
res_name_DEX2 <- subset(res_DEX_name, adj.P.Val < 0.05)
res_name_IL42 <- subset(res_IL4_name, adj.P.Val < 0.05)
```

### LPS Overlap DTU transcripts and genes 
#How many unique genes and significance?

```{r LPS overlap, echo = TRUE}
overlap_LPS <- merge(LPS_S, res_name_LPS2, by = "symbol")
unique(overlap_LPS$symbol)
set1 <- LPS_S
set2 <- res_name_LPS2

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```

### IFNy Overlap DTU transcripts and genes 
#How many unique genes and significance?

```{r IFNy overlap, echo = TRUE}
overlap_IFNy <- merge(IFNy_S, res_name_IFNy2, by = "symbol")
unique(overlap_IFNy$symbol)
set1 <- IFNy_S
set2 <- res_name_IFNy2

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```


### R848 Overlap DTU transcripts and genes 
#How many unique genes and significance?

```{r R848 overlap, echo = TRUE}
overlap_R848 <- merge(R848_S, res_name_R8482, by = "symbol")
unique(overlap_R848$symbol)
set1 <- R848_S
set2 <- res_name_R8482

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```

### IL4 Overlap DTU transcripts and genes 
#How many unique genes and significance?

```{r IL4 overlap, echo = TRUE}
overlap_IL4 <- merge(IL4_S, res_name_IL42, by = "symbol")
unique(overlap_IL4$symbol)
set1 <- IL4_S
set2 <- res_name_IL42

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```


### ATP Overlap DTU transcripts and genes 
#How many unique genes and significance?

```{r ATP overlap, echo = TRUE}
overlap_ATP <- merge(ATP_S, res_name_ATP2, by = "symbol")
unique(overlap_ATP$symbol)
set1 <- ATP_S
set2 <- res_name_ATP2

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```



