---
title: "Genelists_overlap_R848"
author: "Gijsje"
date: "5/13/2021"
output: 
 rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---

```{r setup, include=FALSE}
library('variancePartition')
library('edgeR')
library(dplyr)
library(tidyr)
library(readxl)
library(edgeR)
library(limma)
library(DESeq2)
library(tidyverse)
library(data.table)
library(shiny)
library(pheatmap)
library(ggfortify)
library(ggplot2)
library(UpSetR)
library(VennDiagram)
library(knitr) 
library(tidyverse)
library(dplyr)
library(tidyr)
library(tibble)
library(purrr)
library(corrplot)
library(ggplot2)
library(ggpubr)
library(EnhancedVolcano)
library(raster)
library(dplyr)
library(ggrastr)
```

# Enrichment analyses 

```{r function_setenrichment, echo = TRUE}
# Function set enrichment
setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}
```

# Data load

```{r Patir, echo = TRUE}
Patir <- read.csv("~/Documents/MiGASti/Databases/Patir.txt", sep="")
names(Patir)[1] <- "symbol"

AD = "~/Documents/MiGASti/Databases/Alzheimers_disease.xlsx"
AD = read_excel(AD, col_names = TRUE) 
AD = as.data.frame(AD)

MS = "~/Documents/MiGASti/Databases/Multiple_sclerosis.xlsx"
MS = read_excel(MS, col_names = TRUE) 
MS = as.data.frame(MS)

DAM_down = "~/Documents/MiGASti/Databases/DAM_DOWN.xlsx"
DAM_down = read_excel(DAM_down, col_names = TRUE) 
DAM_down = as.data.frame(DAM_down)

DAM_up = "~/Documents/MiGASti/Databases/DAM_up.xlsx"
DAM_up = read_excel(DAM_up, col_names = TRUE) 
DAM_up = as.data.frame(DAM_up)

HAM_up = "~/Documents/MiGASti/Databases/HAM_up.xlsx"
HAM_up = read_excel(HAM_up, col_names = TRUE) 
HAM_up = as.data.frame(HAM_up)

HAM_down = "~/Documents/MiGASti/Databases/HAM_DOWN.xlsx"
HAM_down = read_excel(HAM_down, col_names = TRUE) 
HAM_down = as.data.frame(HAM_down)

TWAS_PD = "~/Documents/MiGASti/Databases/TWAS_PD.xlsx"
TWAS_PD = read_excel(TWAS_PD, col_names = TRUE) 
TWAS_PD = as.data.frame(TWAS_PD)

TWAS_AD = "~/Documents/MiGASti/Databases/TWAS_AD.xlsx"
TWAS_AD = read_excel(TWAS_AD, col_names = TRUE) 
TWAS_AD = as.data.frame(TWAS_AD)

#Functions:

NFkB = "~/Documents/MiGASti/Databases/NfkB_activation.xlsx"
NFkB = read_excel(NFkB, col_names = TRUE) 
NFkB = as.data.frame(NFkB)

INFy_GO = "~/Documents/MiGASti/Databases/IFNy_GO.xlsx"
INFy_GO = read_excel(INFy_GO, col_names = TRUE) 
INFy_GO = as.data.frame(INFy_GO)

PIG = "~/Documents/MiGASti/Databases/PIG.xlsx"
PIG = read_excel(PIG, col_names = TRUE) 
PIG = as.data.frame(PIG)

Sensome = "~/Documents/MiGASti/Databases/Sensome_human.xlsx"
Sensome = read_excel(Sensome, col_names = TRUE) 
Sensome = as.data.frame(Sensome)

HOM = "~/Documents/MiGASti/Databases/Homeostatic_genes.xlsx"
HOM = read_excel(HOM, col_names = TRUE) 
HOM = as.data.frame(HOM)

Phago = "~/Documents/MiGASti/Databases/Phagocytosis.xlsx"
Phago = read_excel(Phago, col_names = TRUE) 
Phago = as.data.frame(Phago)

load("~/Documents/MiGASti/docs/2nd_pass/AD_bulk_3datasets.Rdata")
AD_bulk <- genes3

AD_eQTL = "~/Documents/MiGASti/Databases/AD_eQTL.xlsx"
AD_eQTL = read_excel(AD_eQTL, col_names = TRUE) 
AD_eQTL = as.data.frame(AD_eQTL)

PD_eQTL = "~/Documents/MiGASti/Databases/PD_eQTL.xlsx"
PD_eQTL = read_excel(PD_eQTL, col_names = TRUE) 
PD_eQTL = as.data.frame(PD_eQTL)

AD_GWAS_21 = "~/Documents/MiGASti/Databases/GWAS_AD_2021.xlsx"
AD_GWAS_21 = read_excel(AD_GWAS_21, col_names = TRUE) 
AD_GWAS_21 = as.data.frame(AD_GWAS_21)

MS_genes = "~/Documents/MiGASti/Databases/TWAS_MS.xlsx"
MS_genes = read_excel(MS_genes, col_names = TRUE) 
MS_genes = as.data.frame(MS_genes)

AD_genes = rbind(AD_GWAS_21, AD_eQTL, TWAS_AD)
AD_genes2 = unique(AD_genes)

PD_genes = rbind(PD_eQTL, TWAS_PD)
PD_genes2= unique(PD_genes)

HAM = rbind(HAM_up, HAM_down)
HAM2 = unique(HAM)

DAM = rbind(DAM_up, DAM_down)
DAM2 = unique(DAM)
```

# R848 UP FDR 5% logFC > 0.5 

```{r vul_R848, echo = TRUE}
load("~/Documents/MiGASti/docs/2nd_pass/res_R848_name.Rdata")
res_R848 <- res_R848_name

sign <- subset(res_R848, adj.P.Val < 0.05)
R848_UP <- subset(sign, logFC > 0.5)

set1 <- R848_UP
set1_v <- set1$symbol
```

### + Patir

```{r R848_Patir, echo = TRUE}
set2 <- Patir
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + NFkB GO 

```{r R848_NFkB, echo = TRUE}
set2 <- NFkB
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + INFy GO

```{r R848_INFy_GO, echo = TRUE}
set2 <- INFy_GO
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + Plaque induced genes 

```{r R848_PIG, echo = TRUE}
set2 <- PIG
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + Sensome genes 

```{r R848_Sensome_check, echo = TRUE}
set2 <- Sensome
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + phagocytosis 

```{r R848_Sensome1, echo = TRUE}
set2 <- Phago
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + homeostatic genes

```{r R848_HOM2, echo = TRUE}
set2 <- HOM
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + AD genes snRNAseq 

```{r R848_AD, echo = TRUE}
set2 <- AD
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + TWAS AD genes

```{r R848_AD_TWAS, echo = TRUE}
set2 <- TWAS_AD
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + AD genes bulk brain 

```{r R848_AD_bulk5, echo = TRUE}
set2 <- AD_bulk
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + AD eQTL

```{r R848_AD_eQTL, echo = TRUE}
set2 <- AD_eQTL
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + AD GWAS Bellenguez

```{r R848_AD_GWAS_21, echo = TRUE}
set2 <- AD_GWAS_21
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + Combined AD 

```{r R848_AD_combined1, echo = TRUE}
set2 <- AD_genes2
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + MS human snRNAseq 

```{r R848_MS, echo = TRUE}
set2 <- MS
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + TWAS MS

```{r R848_MS_twas, echo = TRUE}
set2 <- MS_genes
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```


### + TWAS PD

```{r R848_TWAS_PD, echo = TRUE}
set2 <- TWAS_PD
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + eQTL PD 

```{r R848_eQTL_PD, echo = TRUE}
set2 <- PD_eQTL
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + Combined PD 

```{r R848_PD_combined1, echo = TRUE}
set2 <- PD_genes2
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + HAM_up

```{r R848_HAM_up, echo = TRUE}
set2 <- HAM_up
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + HAM_down

```{r R848_HAM_down, echo = TRUE}
set2 <- HAM_down
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + combined HAM

```{r R848_HAM1, echo = TRUE}
set2 <- HAM2
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + DAM_up

```{r R848_DAM_up, echo = TRUE}
set2 <- DAM_up
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + DAM_down

```{r R848_DAM_down, echo = TRUE}
set2 <- DAM_down
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### Combined DAM 

```{r R848_DAM1, echo = TRUE}
set2 <- DAM2
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

# R848 DOWN FDR 5% logFC < -0.5 

```{r vul_R8482, echo = TRUE}
load("~/Documents/MiGASti/docs/2nd_pass/res_R848_name.Rdata")
res_R848 <- res_R848_name

sign <- subset(res_R848, adj.P.Val < 0.05)
R848_DOWN <- subset(sign, logFC < -0.5)

set1 <- R848_DOWN
set1_v <- set1$symbol
```

### + Patir

```{r R848_Patir2, echo = TRUE}
set2 <- Patir
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + NFkB GO 

```{r R848_NFkB2, echo = TRUE}
set2 <- NFkB
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + INFy GO

```{r R848_INFy2, echo = TRUE}
set2 <- INFy_GO
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + Plaque induced genes 

```{r R848_PIG2, echo = TRUE}
set2 <- PIG
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + Sensome genes 

```{r R848_Sensome2, echo = TRUE}
set2 <- Sensome
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + phagocytosis

```{r R848_phago3, echo = TRUE}
set2 <- Phago
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + homeostatic genes

```{r R848_HOM3, echo = TRUE}
set2 <- HOM
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + AD genes human snRNAseq 

```{r R848_AD2, echo = TRUE}
set2 <- AD
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + AD genes bulk brain 

```{r R848_AD_bulk2, echo = TRUE}
set2 <- AD_bulk
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + TWAS AD genes

```{r R848_AD_TWAS2, echo = TRUE}
set2 <- TWAS_AD
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + AD eQTL

```{r R848_AD_eQTL2, echo = TRUE}
set2 <- AD_eQTL
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + AD GWAS Bellenguez

```{r R848_AD_GWAS_22, echo = TRUE}
set2 <- AD_GWAS_21
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + Combined AD 

```{r R848_AD_combined2, echo = TRUE}
set2 <- AD_genes2
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + MS human snRNAseq

```{r R848_MS2, echo = TRUE}
set2 <- MS
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + TWAS MS

```{r R848_MS_twas2, echo = TRUE}
set2 <- MS_genes
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + TWAS PD

```{r R848_TWAS_PD2, echo = TRUE}
set2 <- TWAS_PD
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + Combined PD 

```{r R848_PD_combined2, echo = TRUE}
set2 <- PD_genes2
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + HAM_up

```{r R848_HAM_up2, echo = TRUE}
set2 <- HAM_up
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + HAM_down

```{r R848_HAM_down2, echo = TRUE}
set2 <- HAM_down
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### Combined HAM 

```{r R848_HAM2, echo = TRUE}
set2 <- HAM2
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + DAM_up

```{r R848_DAM_up2, echo = TRUE}
set2 <- DAM_up
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### + DAM_down

```{r R848_DAM_down2, echo = TRUE}
set2 <- DAM_down
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```

### Combined DAM 

```{r R848_DAM2, echo = TRUE}
set2 <- DAM2
set2_v <- set2$symbol
setEnrichment( set1 = set1_v, set2 = set2_v)
overlap <- merge(set1, set2, by = "symbol")
list(overlap$symbol)
```
