---
title: "heatmap_kmeans"
output: 
 rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
   
---

```{r setup, echo=FALSE, results = 'hide'}
library(readxl)
library(edgeR)
library(limma)
library(DESeq2)
library(tidyverse)
library(data.table)
library(pheatmap)
library(readxl) 
library(ComplexHeatmap)
library(RColorBrewer)
library(dplyr)
library(tidyverse)
library(ggsci)
library(factoextra)
library(ggsci)
library(amap)
library(scales)
library(ggeasy)
library(ggtext)
```

# Treshold FDR 5% log FC > 1 or < -1 (210 genes)

```{r data_load, echo=TRUE}
load("~/df_num_LPS_IFNy_LOGFC1.Rdata")
load("~/Documents/MiGASti/docs/Kmeans210gsupport_LPS_IFNy.Rdata")
df <- df_num
df3 <- scale(df)
```

###  4 clusters

```{r gap100, echo=TRUE}
set.seed(123)
k <- kmeans(df3, centers = 4, nstart = 25)
kmeans <- k$cluster
kmean_download <- as.data.frame(kmeans)
kmeanscor <- cbind(kmean_download, DF_log2FC)
gencode_30 <- read.delim("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
colnames(gencode_30) = c("ensembl","symbol")
kmeanscor1 = merge(kmeanscor, gencode_30, by="ensembl")

cluster1 <- kmeanscor1[kmeanscor1$kmeans == 1,]
cluster2 <- kmeanscor1[kmeanscor1$kmeans == 2,]
cluster3 <- kmeanscor1[kmeanscor1$kmeans == 3,]
cluster4 <- kmeanscor1[kmeanscor1$kmeans == 4,]

list(cluster4$symbol)
list(cluster3$symbol)
list(cluster1$symbol)
list(cluster2$symbol)
```

# Heatmap 

```{r _heatmap, echo = TRUE, fig.width=10, fig.height=6, dpi=300}
library(dplyr)
#plot the genes based on samples 
rownames(df3) = kmeanscor1$symbol
kmeanscor2 <- kmeanscor1[order(kmeanscor1$kmeans),]
dataframe <- kmeanscor2[,3:4]
df2 <- as.matrix(dataframe)
rownames(df2) = kmeanscor2$symbol

ann <- data.frame(kmeanscor2$kmeans)
colnames(ann) <- c('cluster')
colours <- list('cluster' = c('1' = 'red', '2' = 'royalblue',  '3'  = 'navy', '4'='green'))
row_ha <- rowAnnotation(df = ann)
as.factor(row_ha)

colPalette <- colorRampPalette(rev(brewer.pal(10, "RdYlBu")))(400)
Heatmap(df2,
        col = colPalette,
        name = "logFC", # Legend title
        cluster_rows = F,
        cluster_columns = T,
        #  row_names_gp = gpar(fontsize = 6), # Text size for row names)
        show_column_dend = F, 
        show_column_names = T,
        show_row_names = F,
        right_annotation = row_ha)
```


```{r _gprofiler, echo = TRUE}
library(gProfileR)
go_res <- gprofiler(cluster1, organism = "hsapiens", hier_filtering= "strong", min_set_size=5,  src_filter="GO" )
go_res %>% select( p.value, term.size, overlap.size, domain, term.name) %>% as_tibble()  %>% arrange(p.value)
go_res

go_res <- gprofiler(cluster2, organism = "hsapiens", hier_filtering= "strong", min_set_size=5,  src_filter="GO" )
go_res %>% select( p.value, term.size, overlap.size, domain, term.name) %>% as_tibble()  %>% arrange(p.value)

go_res <- gprofiler(cluster3, organism = "hsapiens", hier_filtering= "strong", min_set_size=5,  src_filter="GO" )
go_res %>% select( p.value, term.size, overlap.size, domain, term.name) %>% as_tibble()  %>% arrange(p.value)

go_res <- gprofiler(cluster4, organism = "hsapiens", hier_filtering= "strong", min_set_size=5,  src_filter="GO" )
go_res %>% select( p.value, term.size, overlap.size, domain, term.name) %>% as_tibble()  %>% arrange(p.value)

 






