---
title: "Differential expression analysis of all stimuli compared to unstimulated in 1 region only"
author: "Gijsje"
date: "2/16/2021"
output: 
 rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---

```{r load.packages, echo=FALSE, message=FALSE, results='hide'}
library(readxl)
library(edgeR)
library(limma)
library(DESeq2)
library(tidyverse)
library(data.table)
library(shiny)
library(DT)


knitr::opts_chunk$set( 
  warning=FALSE,
  message=FALSE,
  results = 'asis',
  error = FALSE,
  tidy = FALSE,
  fig.show = "hold")
```

```{r Helper Functions, echo=FALSE}
createDT <- function(DF, caption="", scrollY=500){
data <- DT::datatable(DF, caption=caption,
extensions = 'Buttons',
options = list( dom = 'Bfrtip',
buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
scrollY = scrollY, scrollX=T, scrollCollapse = T, paging = F,
columnDefs = list(list(className = 'dt-center', targets = "_all"))
)
)
return(data)
}
```

```{r data_load, echo = TRUE, results = 'hide'}
load("~/Documents/MiGASti/Databases/gene_matrix.RData")
metadata <- read.table("~/Documents/MiGASti/Databases/metadata2.txt")
#set rownames to Sample
row.names(metadata) <- metadata$Sample 
setwd("~/Documents/MiGASti/Databases")
#exclude samples that did not pass QC filtering
exclude <- read.table("samples2remove2.txt")
exclude <- exclude$x
genes_counts_filt = genes_counts[, !colnames(genes_counts) %in% exclude] 
#Excludes the samples from filters. 
#dim(genes_counts_filt)
metadata_filt = metadata[ !(rownames(metadata) %in% exclude), ]
length(metadata_filt)
gencode_30 = read.table("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
colnames(gencode_30) = c("ensembl","symbol")
#order metadata and genes counts
genes_counts_ordered <- genes_counts_filt[,rownames(metadata_filt)]
#head(genes_counts_ordered)
all(rownames(metadata_filt) == colnames (genes_counts_ordered)) #TRUE
```

# Preparing the samples for DEG

```{r preparing, echo = TRUE}
#remove uncultured samples
metadata_cultured <- metadata_filt[metadata_filt$Stimulation != "ununstim",]
#remove uncultured samples
metadata_cultured_subset <- metadata_cultured[metadata_cultured$Stimulation != "TNFa",]
#check numbers
#dim(metadata_cultured)
#check numbers per stimulation
#table(metadata_filt$Stimulation)
#select only GFM samples in metadata
metadata_GFM = metadata_cultured_subset[metadata_cultured_subset$Region=="GFM",]
#select only GFM samples in genes counts
genes_counts_GFM <- genes_counts_ordered[,metadata_GFM$Sample]
#order metadata and genes_counts
genes_counts_GFM_ordered <- genes_counts_GFM[,rownames(metadata_GFM)]
#check ordering
all(rownames(metadata_GFM) == colnames (genes_counts_GFM_ordered)) #TRUE
#check numbers
#dim(metadata_cultured)
#check numbers per stimulation
#table(metadata_filt$Stimulation)
#select only GFM samples in metadata
metadata_GFM = metadata_cultured[metadata_cultured$Region=="GFM",]
#select only GFM samples in genes counts
genes_counts_GFM <- genes_counts_ordered[,metadata_GFM$Sample]
#order metadata and genes_counts
genes_counts_GFM_ordered <- genes_counts_GFM[,rownames(metadata_GFM)]
#check ordering
all(rownames(metadata_GFM) == colnames (genes_counts_GFM_ordered)) #TRUE

#round counts; deseq2 can only handle integers
genes_counts_GFM_ordered <- round(genes_counts_GFM_ordered, digits=0)

#make sure covariate variables are the right format 
#cannot create dds object with numeric values
metadata_GFM$Donor_id <- as.factor(metadata_GFM$Donor_id)
metadata_GFM$age <- as.integer(metadata_GFM$age)
metadata_GFM$sex <- as.factor(metadata_GFM$sex)
metadata_GFM$Stimulation <- as.factor(metadata_GFM$Stimulation)
metadata_GFM$picard_pct_ribosomal_bases = scale(metadata_GFM$picard_pct_ribosomal_bases)
metadata_GFM$picard_pct_mrna_bases = scale(metadata_GFM$picard_pct_mrna_bases)
metadata_GFM$picard_pct_pf_reads_aligned = scale(metadata_GFM$picard_pct_pf_reads_aligned)
metadata_GFM$picard_pct_percent_duplication = scale(metadata_GFM$picard_pct_percent_duplication)

#adjust for: ~ age + (1|donor_id) + picard_pct_ribosomal_bases + picard_pct_mrna_bases +   picard_pct_percent_duplication + picard_pct_pf_reads_aligned 
```

# DESeq2 of GFM samples 

```{r data_loadings, echo = TRUE}
#createDeSEQ2 object for LPS
dds <- DESeqDataSetFromMatrix(countData = genes_counts_GFM_ordered,
                              colData = metadata_GFM,
                              design = ~ age + sex + picard_pct_ribosomal_bases + picard_pct_mrna_bases + picard_pct_percent_duplication + picard_pct_pf_reads_aligned + Stimulation) 
#variable of interest at end of the formula

#Make sure that control group is set as the reference group
dds$Stimulation <- relevel(dds$Stimulation, ref="unstim")
table(dds$Stimulation) 
#head(dds)

#filter: CPM > 1 in 50% of the samples 
keep.exp = rowSums(cpm(genes_counts_GFM_ordered) > 1) >= 0.5*ncol(genes_counts_GFM_ordered)
dds = dds[keep.exp,]

#Run differential expression 
dds <- DESeq(dds, betaPrior = FALSE)
resultsNames(dds)
```

# DESeq2: LPS vs unstim
### Number of differentially expressed genes 

```{r data_LPS, echo = TRUE}
# generate results table for LPS vs unstim
res_LPS <- results(dds, name="Stimulation_LPS_vs_unstim")
sum(res_LPS$padj < 0.05, na.rm=TRUE)
resOrdered_LPS <- res_LPS[order(res_LPS$pvalue),] 
resOrdered_LPS <- as.data.frame(resOrdered_LPS)
```

### Volcano plot LPS vs unstim

```{r vulcano_LPS, echo = TRUE}
head(res_LPS)
with(res_LPS, plot(log2FoldChange, -log10(pvalue), pch=20, main="Volcano plot", xlim=c(-2.5,2)))
with(subset(res_LPS, padj<.05 ), points(log2FoldChange, -log10(pvalue), pch=20, col="red"))
```

### MA plot LPS vs unstim
#The function plotMA shows the log2 fold changes attributable to a given variable over the mean of normalized counts for all the samples in the DESeqDataSet. Points will be colored if the adjusted p value is less than 0.1. Points which fall out of the window are plotted as open triangles pointing either up or down.

```{r MA_LPS, echo = TRUE}
plotMA(res_LPS, ylim=c(-2,2))
```

### TOP differentially expressed genes LPS vs unstim

```{r diff_LPS, echo = TRUE}
setDT(resOrdered_LPS, keep.rownames = "ensembl")
resOrdered_LPS <- left_join(resOrdered_LPS, gencode_30, by = "ensembl")
resOrdered_LPS_top = resOrdered_LPS[order(resOrdered_LPS$padj) ,]
setDT(resOrdered_LPS_top, keep.rownames = "ensembl")
resOrdered_LPS_top = resOrdered_LPS_top[, c("ensembl", "symbol", "baseMean", "log2FoldChange", "lfcSE", "stat", "pvalue", "padj")]
createDT(resOrdered_LPS_top)
write.table(resOrdered_LPS_top, "DEG_LPS_GFM.txt")
```

# DESeq2: IFNy vs unstim
### Number of differentially expressed genes 

```{r IFNy, echo = TRUE}
# generate results table for IFNy vs unstim
res_IFNy <- results(dds, name="Stimulation_IFNy_vs_unstim")
sum(res_IFNy$padj < 0.05, na.rm=TRUE)
resOrdered_IFNy <- res_IFNy[order(res_IFNy$pvalue),] 
resOrdered_IFNy <- as.data.frame(resOrdered_IFNy)
```

### Volcano plot IFNy vs unstim

```{r vulcano_IFNy, echo = TRUE}
head(res_IFNy)
with(res_IFNy, plot(log2FoldChange, -log10(pvalue), pch=20, main="Volcano plot", xlim=c(-10,10)))
with(subset(res_IFNy, padj<.05 ), points(log2FoldChange, -log10(pvalue), pch=20, col="red"))
```

### MA plot IFNy vs unstim

```{r MA_IFNy, echo = TRUE}
plotMA(res_IFNy, ylim=c(-2,2))
```

### TOP differentially expressed genes IFNy vs unstim

```{r diff_IFNy, echo = TRUE}
setDT(resOrdered_IFNy, keep.rownames = "ensembl")
resOrdered_IFNy <- left_join(resOrdered_IFNy, gencode_30, by = "ensembl")
resOrdered_IFNy_top = resOrdered_IFNy[order(resOrdered_IFNy$padj) ,]
resOrdered_IFNy_top = resOrdered_IFNy_top[, c("ensembl", "symbol", "baseMean", "log2FoldChange", "lfcSE", "stat", "pvalue", "padj")]
createDT(resOrdered_IFNy_top)
write.table(resOrdered_IFNy_top, "DEG_IFNy_GFM.txt")
```

# DESeq2: R848 vs unstim
### Number of differentially expressed genes

```{r R848, echo = TRUE}
# generate results table for R848 vs unstim
res_R848 <- results(dds, name="Stimulation_R848_vs_unstim")
sum(res_R848$padj < 0.05, na.rm=TRUE)
resOrdered_R848 <- res_R848[order(res_R848$pvalue),] 
resOrdered_R848 <- as.data.frame(resOrdered_R848)
```

### Volcano plot R848 vs unstim

```{r vulcano_R848, echo = TRUE}
head(res_R848)
with(res_R848, plot(log2FoldChange, -log10(pvalue), pch=20, main="Volcano plot", xlim=c(-10,10)))
with(subset(res_R848, padj<.05 ), points(log2FoldChange, -log10(pvalue), pch=20, col="red"))
```

### MA plot R848 vs unstim

```{r MA_R848, echo = TRUE}
plotMA(res_R848, ylim=c(-2,2))
```

### TOP differentially expressed genes R848 vs unstim

```{r diff_R848, echo = TRUE}
setDT(resOrdered_R848, keep.rownames = "ensembl")
resOrdered_R848 <- left_join(resOrdered_R848, gencode_30, by = "ensembl")
resOrdered_R848_top = resOrdered_R848[order(resOrdered_R848$padj) ,]
setDT(resOrdered_R848_top, keep.rownames = "ensembl")
resOrdered_R848_top = resOrdered_R848_top[, c("ensembl", "symbol", "baseMean", "log2FoldChange", "lfcSE", "stat", "pvalue", "padj")]
createDT(resOrdered_R848_top)
```

# DESeq2: DEX vs unstim
### Number of differentially expressed genes

```{r DEX, echo = TRUE}
# generate results table for DEX vs unstim
res_DEX <- results(dds, name="Stimulation_DEX_vs_unstim")
sum(res_DEX$padj < 0.05, na.rm=TRUE)
resOrdered_DEX <- res_DEX[order(res_DEX$pvalue),] 
resOrdered_DEX <- as.data.frame(resOrdered_DEX)
```

### Volcano plot DEX vs unstim

```{r vulcano_DEX, echo = TRUE}
head(res_DEX)
with(res_DEX, plot(log2FoldChange, -log10(pvalue), pch=20, main="Volcano plot", xlim=c(-10,10)))
with(subset(res_DEX, padj<.05 ), points(log2FoldChange, -log10(pvalue), pch=20, col="red"))
```

### MA plot DEX vs unstim

```{r MA_DEX, echo = TRUE}
plotMA(res_DEX, ylim=c(-2,2))
```

### TOP differentially expressed genes DEX vs unstim

```{r diff_DEX, echo = TRUE}
setDT(resOrdered_DEX, keep.rownames = "ensembl")
resOrdered_DEX <- left_join(resOrdered_DEX, gencode_30, by = "ensembl")
resOrdered_DEX_top = resOrdered_DEX[order(resOrdered_DEX$padj) ,]
setDT(resOrdered_DEX_top, keep.rownames = "ensembl")
resOrdered_DEX_top = resOrdered_DEX_top[, c("ensembl", "symbol", "baseMean", "log2FoldChange", "lfcSE", "stat", "pvalue", "padj")]
createDT(resOrdered_DEX_top)
#write.table(resOrdered_DEX_top, "DEG_DEX_GFM.txt")
```

# DESeq2: Il4 vs unstim
### Number of differentially expressed genes

```{r IL4, echo = TRUE}
res_IL4 <- results(dds, name="Stimulation_IL4_vs_unstim")
sum(res_IL4$padj < 0.05, na.rm=TRUE)
resOrdered_IL4 <- res_IL4[order(res_IL4$pvalue),] 
resOrdered_IL4 <- as.data.frame(resOrdered_IL4)
```


### Volcano plot IL4 vs unstim

```{r vulcano_IL4, echo = TRUE}
head(res_IL4)
with(res_IL4, plot(log2FoldChange, -log10(pvalue), pch=20, main="Volcano plot", xlim=c(-10,10)))
with(subset(res_IL4, padj<.05 ), points(log2FoldChange, -log10(pvalue), pch=20, col="red"))
```


### MA plot IL4 vs unstim

```{r MA_IL4, echo = TRUE}
plotMA(res_IL4, ylim=c(-2,2))
```

### TOP differentially expressed genes IL4 vs unstim

```{r diff_IL4, echo = TRUE}
setDT(resOrdered_IL4, keep.rownames = "ensembl")
resOrdered_IL4 <- left_join(resOrdered_IL4, gencode_30, by = "ensembl")
resOrdered_IL4_top = resOrdered_IL4[order(resOrdered_IL4$padj) ,]
setDT(resOrdered_IL4_top, keep.rownames = "ensembl")
resOrdered_IL4_top = resOrdered_IL4_top[, c("ensembl", "symbol", "baseMean", "log2FoldChange", "lfcSE", "stat", "pvalue", "padj")]
createDT(resOrdered_IL4_top)
write.table(resOrdered_IL4_top, "DEG_IL4_GFM.txt")
```

# DESeq2: ATP vs unstim
### Number of differentially expressed genes

```{r ATP, echo = TRUE}
# generate results table for ATP vs unstim
res_ATP <- results(dds, name="Stimulation_ATP_vs_unstim")
sum(res_ATP$padj < 0.05, na.rm=TRUE)
resOrdered_ATP <- res_ATP[order(res_ATP$pvalue),] 
resOrdered_ATP <- as.data.frame(resOrdered_ATP)
```

### Volcano plot ATP vs unstim

```{r vulcano_ATP, echo = TRUE}
head(res_ATP)
with(res_ATP, plot(log2FoldChange, -log10(pvalue), pch=20, main="Volcano plot", xlim=c(-10,10)))
with(subset(res_ATP, padj<.05 ), points(log2FoldChange, -log10(pvalue), pch=20, col="red"))
```

### MA plot ATP vs unstim

```{r MA_ATP, echo = TRUE}
plotMA(res_ATP, ylim=c(-2,2))
```

### TOP differentially expressed genes ATP vs unstim

```{r diff_ATP, echo = TRUE}
setDT(resOrdered_ATP, keep.rownames = "ensembl")
resOrdered_ATP <- left_join(resOrdered_ATP, gencode_30, by = "ensembl")
resOrdered_ATP_top = resOrdered_ATP[order(resOrdered_ATP$padj) ,]
setDT(resOrdered_ATP_top, keep.rownames = "ensembl")
resOrdered_ATP_top = resOrdered_ATP_top[, c("ensembl", "symbol", "baseMean", "log2FoldChange", "lfcSE", "stat", "pvalue", "padj")]
createDT(resOrdered_ATP_top)
write.table(resOrdered_ATP_top, "DEG_ATP_GFM.txt")
```

#Create genelists with Log2FC < 1 and < -1 for different stimuli

```{r lists, echo = TRUE}
resOrdered_LPS_p <- subset(resOrdered_LPS_top, padj < 0.05)
resOrdered_LPS_LFC <- subset(resOrdered_LPS_p, log2FoldChange > 1 | log2FoldChange < -1)
write.table(resOrdered_LPS_LFC, "LPS_GFM_FC1.txt")

resOrdered_IFNy_p <- subset(resOrdered_IFNy_top, padj < 0.05)
resOrdered_IFNy_LFC <- subset(resOrdered_IFNy_p, log2FoldChange > 1 | log2FoldChange < -1)
write.table(resOrdered_IFNy_LFC, "IFNy_GFM_FC1.txt")

resOrdered_R848_p <- subset(resOrdered_R848_top, padj < 0.05)
resOrdered_R848_LFC <- subset(resOrdered_R848_p, log2FoldChange > 1 | log2FoldChange < -1)
write.table(resOrdered_R848_LFC, "R848_GFM_FC1.txt")

resOrdered_DEX_p <- subset(resOrdered_DEX_top, padj < 0.05)
resOrdered_DEX_LFC <- subset(resOrdered_DEX_p, log2FoldChange > 1 | log2FoldChange < -1)
write.table(resOrdered_DEX_LFC, "DEX_GFM_FC1.txt")

resOrdered_IL4_p <- subset(resOrdered_IL4_top, padj < 0.05)
resOrdered_IL4_LFC <- subset(resOrdered_IL4_p, log2FoldChange > 1 | log2FoldChange < -1)
write.table(resOrdered_IL4_LFC, "IL4_GFM_FC1.txt")

resOrdered_ATP_p <- subset(resOrdered_ATP_top, padj < 0.05)
resOrdered_ATP_LFC <- subset(resOrdered_ATP_p, log2FoldChange > 1 | log2FoldChange < -1)
write.table(resOrdered_ATP_LFC, "ATP_GFM_FC1.txt")
```
 










