---
title: "Markers TNF"
author: "Gijsje"
date: "2/25/2021"
output: 
 rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---

```{r load.packages, echo=FALSE, message=FALSE, results='hide'}
library(readxl)
library(edgeR)
library(limma)
library(DESeq2)
library(tidyverse)
library(data.table)
library(pheatmap)


knitr::opts_chunk$set( 
  warning=FALSE,
  message=FALSE,
  results = 'asis',
  error = FALSE,
  tidy = FALSE,
  fig.show = "hold")
```


```{r check, echo = TRUE, results = 'hide'}
load("~/Documents/MiGASti/Databases/gene_matrix.RData")
metadata <- read.table("~/Documents/MiGASti/Databases/metadata.txt")
#set rownames to Sample
row.names(metadata) <- metadata$Sample 
setwd("~/Documents/MiGASti/Databases")
exclude <- read.table("samples2remove.txt")
exclude <- exclude$x
genes_tpm_filt = genes_tpm[, !colnames(genes_tpm) %in% exclude] 
#Excludes the samples from filters. 
#dim(genes_tpm_filt)
metadata_filt = metadata[ !(rownames(metadata) %in% exclude), ]
length(metadata_filt)
#check for expression of markers in dataset
markers = "~/Documents/MiGASti/Databases/Markers_TNFa.xlsx"
markers = read_excel(markers, col_names = TRUE) 
markers = as.data.frame(markers)
gencode_30 = read.table("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
colnames(gencode_30) = c("ensembl","symbol")
genes_tpm_filt = log2((genes_tpm_filt) + 1)
genes_tpm_filt <- as.data.frame(genes_tpm_filt)
setDT(genes_tpm_filt, keep.rownames = "ensembl")
genes_tpm_filt$ensembl
res_name = merge(genes_tpm_filt, gencode_30, by="ensembl")
rownames(res_name) = res_name$ensembl
```


```{r data_load_markers, echo = TRUE, results = 'hide'}
marker_expression = merge(res_name, markers, by ="symbol")
dim (marker_expression)
head (marker_expression)
metadata_filt_ordered <- metadata_filt[colnames(genes_tpm_filt),]
metadata_filt_ordered = metadata_filt_ordered[-1,]
all(rownames(metadata_filt_ordered) == colnames (genes_tpm_filt)) #TRUE
Stimulation <- metadata_filt_ordered$Stimulation
```

# TNF a markers
### CD300E
#expected to be increased in TNFa: based on Park et al. 2017 (monocytes)

```{r data_load_IL1B, echo = TRUE}
CD300E <- marker_expression[1, ] 
CD300E$ensembl = NULL
CD300E$symbol = NULL
CD300E_new<-as.data.frame(t(CD300E))
df = data.frame(Stimulation, CD300E_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("CD300E") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### FTH1
#expected to be increased in TNFa

```{r data_load_IL6, echo = TRUE}
FTH1 <- marker_expression[2, ] 
FTH1$ensembl = NULL
FTH1$symbol = NULL
FTH1_new<-as.data.frame(t(FTH1))
df = data.frame(Stimulation, FTH1_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("FTH1") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### MADD
#expected to be increased in TNFa

```{r data_load_TNF, echo = TRUE}
MADD <- marker_expression[3, ] 
MADD$ensembl = NULL
MADD$symbol = NULL
MADD_new<-as.data.frame(t(MADD))
df = data.frame(Stimulation, MADD_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot() + 
  ggtitle("MADD") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### MT2A 
#expected to be increased in TNFa

```{r data_load_CXCL9, echo = TRUE}
MT2A <- marker_expression[4, ] 
MT2A$ensembl = NULL
MT2A$symbol = NULL
MT2A_new<-as.data.frame(t(MT2A))
df = data.frame(Stimulation, MT2A_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot() + 
  ggtitle("MT2A") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### S100A8
#expected to be increased in TNFa

```{r data_load_CXCL10, echo = TRUE}
S100A8 <- marker_expression[5, ] 
S100A8$ensembl = NULL
S100A8$symbol = NULL
S100A8_new<-as.data.frame(t(S100A8))
df = data.frame(Stimulation, S100A8_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot() +
ggtitle("S100A8") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### S100A9
#expected to be increased in TNFa

```{r data_load_CXCL11, echo = TRUE}
S100A9 <- marker_expression[6, ] 
S100A9$ensembl = NULL
S100A9$symbol = NULL
S100A9_new<-as.data.frame(t(S100A9))
df = data.frame(Stimulation, S100A9_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot() + 
  ggtitle("S100A9") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### SOD2
#expected to be increased in TNFa

```{r data_load_CCL2, echo = TRUE}
SOD2 <- marker_expression[7, ] 
SOD2$ensembl = NULL
SOD2$symbol = NULL
SOD2_new<-as.data.frame(t(SOD2))
df = data.frame(Stimulation, SOD2_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("SOD2") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### SRGN
#expected to be increased in TNFa

```{r data_load_VCAM1, echo = TRUE}
SRGN <- marker_expression[9, ] 
SRGN$ensembl = NULL
SRGN$symbol = NULL
SRGN_new<-as.data.frame(t(SRGN))
df = data.frame(Stimulation, SRGN_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot() + 
  ggtitle("SRGN") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```
### TIMP1
##expected to be increased in TNFa

```{r data_load_SAA3P, echo = TRUE}
TIMP1 <- marker_expression[10, ] 
TIMP1$ensembl = NULL
TIMP1$symbol = NULL
TIMP1_new<-as.data.frame(t(TIMP1))
df = data.frame(Stimulation, TIMP1_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("TIMP1") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### TRAF1
#expected to be increased in TNFa

```{r data_load_SLC13A3, echo = TRUE}
TRAF1 <- marker_expression[11, ] 
TRAF1$ensembl = NULL
TRAF1$symbol = NULL
TRAF1_new<-as.data.frame(t(TRAF1))
df = data.frame(Stimulation, TRAF1_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("TRAF1") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### TRAF2
#expected to be increased in TNFa

```{r data_load_CXCL1, echo = TRUE}
TRAF2 <- marker_expression[12, ] 
TRAF2$ensembl = NULL
TRAF2$symbol = NULL
TRAF2_new<-as.data.frame(t(TRAF2))
df = data.frame(Stimulation, TRAF2_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("TRAF2") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```










