---
title: "Differential expression analysis LPS and IFNy"
author: "Gijsje"
date: "2/16/2021"
output: 
 rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---

```{r load.packages, echo=FALSE, message=FALSE, results='hide'}
library(readxl)
library(edgeR)
library(limma)
library(DESeq2)
library(tidyverse)
library(data.table)


knitr::opts_chunk$set( 
  warning=FALSE,
  message=FALSE,
  results = 'asis',
  error = FALSE,
  tidy = FALSE,
  fig.show = "hold")
```


```{r data_load, echo = TRUE, results = 'hide'}
load("~/Documents/MiGASti/Databases/gene_matrix.RData")
metadata <- read.table("~/Documents/MiGASti/Databases/metadata.txt")
#set rownames to Sample
row.names(metadata) <- metadata$Sample 
setwd("~/Documents/MiGASti/Databases")
#exclude samples that did not pass QC filtering
exclude <- read.table("samples2remove.txt")
exclude <- exclude$x
genes_counts_filt = genes_counts[, !colnames(genes_counts) %in% exclude] 
#Excludes the samples from filters. 
#dim(genes_counts_filt)
metadata_filt = metadata[ !(rownames(metadata) %in% exclude), ]
length(metadata_filt)
```


```{r data_load_outline, echo = TRUE}
# Counts, TPM, voom and metadata filtered 
# 496 samples!  
# Filter for genes not expressed: 50 % of the samples! 
# dim(genes_counts_ordered) #18997   496
# dim(metadata_filt) #496  38
```

# Preparing the samples for DEG

```{r preparing, echo = TRUE}
#remove uncultured samples
metadata_cultured <- metadata_filt[metadata_filt$Stimulation != "ununstim",]
#check numbers
#dim(metadata_cultured)
#check numbers per stimulation
#table(metadata_filt$Stimulation)
#select only GFM samples in metadata
metadata_unstim = metadata_cultured[metadata_cultured$Stimulation=="unstim"]
metadata_LPS = metadata_cultured[metadata_cultured$Stimulation=="LPS",]
metadata_IFNy = metadata_cultured[metadata_cultured$Stimulation=="IFNy",]


