---
title: "02092021_PCAs_heatmap"
author: "Gijsje"
date: "2/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r echo=FALSE}
#==========================================================================================================#
#                                              required packages                                           #
#==========================================================================================================#
library(devtools)
library(factoextra)
library(ggfortify)
library(ggplot2)
library(dplyr)
library(broom)
library(variancePartition)
library(pheatmap)
library(edgeR)
library(limma)
library(DESeq2)
library(Hmisc)
library(matrixStats)
```

```{r echo=FALSE}
load("~/Documents/MiGASti/Databases/gene_matrix.RData")
metadata <- read.table("~/Documents/MiGASti/Databases/metadata.txt")
#set rownames to Sample
row.names(metadata) <- metadata$Sample 
#remove low count genes
cpm <- cpm(genes_counts) 
# CPM >= 1 in at least 50% of the samples
keep.exp <- rowSums(cpm > 1) >= (0.5 * ncol(genes_counts) )
genes_counts5 <- genes_counts[ keep.exp, ] #18625 genes 
#performing voom normalisation on data
counts_voom <- limma::voom(genes_counts5)
genes_counts_voom <- counts_voom$E
#order metadata and genes counts
rownames(metadata)
colnames(genes_counts_voom)
genes_counts_ordered <- genes_counts_voom[,rownames(metadata)]
head(genes_counts_ordered)
all(rownames(metadata) == colnames (genes_counts_ordered)) #TRUE
```

## PCA with no scaling and centering of all 533 samples (also uncultured samples)

```{r echo=FALSE}
res.pca = prcomp(t(genes_counts_ordered))
autoplot(res.pca, colour = "Stimulation", data = metadata)
```

## PCA with scaling and centering of all 533 samples (also uncultured samples)

```{r echo=FALSE}
pca = prcomp(t(genes_counts_ordered), scale. = TRUE, center = TRUE)
autoplot(pca, colour = "Stimulation", data = metadata)
```

#PCA with scaling and centering + labeling of all 533 samples

```{r echo=FALSE}
pca = prcomp(t(genes_counts_ordered), scale. = TRUE, center = TRUE)
autoplot(pca, data= metadata, colour = 'Stimulation', shape = FALSE, label.size = 2)
```

#Variances per PC of all samples 

```{r echo=FALSE}
plot(pca)
```


## Boxplots of PC1 of all 533 samples

```{r echo=FALSE}
PC1 <- pca$x[,1]
Stimulation <- metadata$Stimulation
df = data.frame(Stimulation, PC1)
ggplot(data = df, mapping = aes(x = Stimulation, y = PC1)) +
geom_boxplot()
```

# Boxplots of PC2 of all 533 samples 

```{r echo=FALSE}
PC2 <- pca$x[,2]
Stimulation <- metadata$Stimulation
df = data.frame(Stimulation, PC2)
ggplot(data = df, mapping = aes(x = Stimulation, y = PC2)) +
geom_boxplot()
```


#removal of uncultured samples (ununstim) 

```{r echo=FALSE}
#label ununstim samples in metadata
metadata_cultured <- metadata[metadata$Stimulation != "ununstim",]
#check numbers
dim(metadata_cultured)
#check numbers per stimulation
table(metadata$Stimulation)
#remove samples in genes counts datafile 
genes_counts_cultured <- genes_counts_ordered[,metadata_cultured$Sample]
```

#PCA of cultured samples with centering and scaling

```{r echo=FALSE}
pca = prcomp(t(genes_counts_cultured), scale. = TRUE, center = TRUE)
autoplot(pca, colour = "Stimulation", data = metadata_cultured)
```

#PCA of cultured samples with labeling of samples 

```{r echo=FALSE}
autoplot(pca, data= metadata_cultured, colour = 'Stimulation', shape = FALSE, label.size = 2)
```

#Variances per PC of cultured samples with centering and scaling

```{r echo=FALSE}
plot(pca)
```

#Boxplot PC1

```{r echo=FALSE}
PC1 <- pca$x[,1]
Stimulation <- metadata_cultured$Stimulation
df = data.frame(Stimulation, PC1)
ggplot(data = df, mapping = aes(x = Stimulation, y = PC1)) +
  geom_boxplot()
```

#Boxplot PC2 

```{r echo=FALSE}
PC2 <- pca$x[,2]
Stimulation <- metadata_cultured$Stimulation
df = data.frame(Stimulation, PC2)
ggplot(data = df, mapping = aes(x = Stimulation, y = PC2)) +
  geom_boxplot()
```

#Boxplot PC3

```{r echo=FALSE}
PC3 <- pca$x[,3]
Stimulation <- metadata_cultured$Stimulation
df = data.frame(Stimulation, PC3)
ggplot(data = df, mapping = aes(x = Stimulation, y = PC3)) +
  geom_boxplot()
```

#Scatterplot PC1 + PC2 

```{r echo=FALSE}
PCdf <- data.frame(sample = rownames(pca$x), pc1 = pca$x[,1], pc2 = pca$x[,2], Stimulation = metadata_cultured$Stimulation)
ggplot(data = PCdf, mapping = aes(x = PC1, y = PC2, colour = Stimulation)) +
  geom_point()
```

#create heatmap first 10 PCs + variables of metadata_cultured

```{r echo=FALSE}
#set metadata to no capitals
names(metadata_cultured) = tolower(names(metadata_cultured))
indx <- sapply(metadata, is.character)
metadata[indx] <- lapply(metadata[indx], function(x) as.factor(x))

#include covariates
covariates = c( "donor_id", 
                "region",
                "stimulation",
                "number",
                "average_library_size",
                "qubit_conc",
                "diagnosis",
                "main_diagnosis",
                "sex",
                "age",
                "pmd_minutes",
                "ph",
                "cause_of_death_categories",
                "smoking",
                "alcohol_dependence_daily_use",
                "autoimmune_diseases",
                "infection_2weeks",
                "opiates",
                "benzodiazepines",
                "psychopharmaca",
                "donor_baseline_pmg",
                "trimmomatic_dropped_pct", 
                "picard_pct_mrna_bases",
                "picard_pct_ribosomal_bases",
                "picard_pct_intronic_bases",
                "picard_pct_intergenic_bases",
                "picard_pct_pf_reads_aligned",
                "picard_summed_median",
                "picard_summed_mean",
                "picard_pct_percent_duplication",
                "star_total_reads",
                "star_uniquely_mapped", 
                "star_uniquely_mapped_percent",
                "featurecounts_assigned")

#create format matrix 
matrix_rsquared = matrix(NA, nrow = length(covariates), ncol = 10) #Number of factors
matrix_pvalue = matrix(NA, nrow = length(covariates), ncol = 10)

#lineair model function 
for (x in 1:length(covariates)){
  for (y in 1:10){
    matrix_rsquared[x,y] <- summary( lm(pca$x[,y] ~ metadata_cultured[,covariates[x]]) )$adj.r.squared
    matrix_pvalue[x,y] <- tidy( lm(pca$x[,y] ~ metadata_cultured[,covariates[x]]) )$p.value[2] #To insert pvalues in the heatmap
  }
}

#fill matrix with values 
matrix_rsquared <- as.data.frame(matrix_rsquared)
matrix_pvalue <- as.data.frame(matrix_pvalue)
rownames(matrix_rsquared) <- covariates
rownames(matrix_pvalue) <- covariates 

#create heatmap with rsquared values 
pheatmap(matrix_rsquared)
#create heatmap with pvalue 
pheatmap(matrix_pvalue)
```

# Create heatmap with first 15 PCs


#Boxplot PC5 

```{r echo=FALSE}
PC5 <- pca$x[,5]
Stimulation <- metadata_cultured$stimulation
df = data.frame(Stimulation, PC5)
ggplot(data = df, mapping = aes(x = Stimulation, y = PC5)) +
  geom_boxplot()
```

#Boxplot PC7

```{r echo=FALSE}
PC7 <- pca$x[,7]
Stimulation <- metadata_cultured$stimulation
df = data.frame(Stimulation, PC7)
ggplot(data = df, mapping = aes(x = Stimulation, y = PC7)) +
  geom_boxplot()
```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
