---
title: "DREAM_age"
author: "Gijsje"
date: "5/13/2021"
output: 
 rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---


```{r setup, include = FALSE}
library('variancePartition')
library('edgeR')
library('BiocParallel')
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggfortify)
library(limma)
library(data.table)
library(readxl)
```

```{r Helper Functions, echo = FALSE}
createDT <- function(DF, caption="", scrollY=500){
data <- DT::datatable(DF, caption=caption,
extensions = 'Buttons',
options = list( dom = 'Bfrtip',
buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
scrollY = scrollY, scrollX=T, scrollCollapse = T, paging = F,
columnDefs = list(list(className = 'dt-center', targets = "_all"))
)
)
return(data)
}
```

# Load data

```{r load, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
#load gene matrix
load("~/Documents/MiGASti/Databases/gene_matrix.RData")
#import dataset metadata
metadata <- read.table("~/Documents/MiGASti/Databases/metadata.txt", quote="\"", comment.char="")
gencode <- rtracklayer::import("gencode.v30.genes.gtf", format = "GTF")
#set rownames to Sample
row.names(metadata) <- metadata$Sample
#exclude samples that did not pass QC
exclude <- read.csv("~/Documents/MiGASti/Databases/samples2remove.txt", sep="")
exclude <- exclude$x
genes_counts_filt = genes_counts[, !colnames(genes_counts) %in% exclude]
#Excludes the samples from filters.
#dim(genes_counts_filt)
metadata_filt = metadata[ !(rownames(metadata) %in% exclude), ]
length(metadata_filt)
#remove low count genes
cpm <- cpm(genes_counts_filt)
# CPM >= 1 in at least 50% of the samples
keep.exp <- rowSums(cpm > 1) >= (0.5 * ncol(genes_counts_filt) )
genes_counts_filt1 <- genes_counts_filt[ keep.exp, ] #18997 genes
gene_id <- gencode$gene_id
#order metadata and genes counts
genes_counts_ordered <- genes_counts_filt1[,rownames(metadata_filt)]
#head(genes_counts_ordered)
all(rownames(metadata_filt) == colnames (genes_counts_ordered)) #TRUE

#remove ununstim samples in metadata
metadata_cultured <- metadata_filt[metadata_filt$Stimulation != "ununstim",]
#check numbers
dim(metadata_cultured)
#check numbers per stimulation
table(metadata_filt$Stimulation)
#remove samples in genes counts datafile
#genes_counts_cultured <- genes_counts_ordered2[,metadata_cultured$Sample]
```

# DREAM model 

```{r DREAM, echo = TRUE}
# The variable to be tested must be a fixed effect
names(metadata_cultured) = tolower(names(metadata_cultured))

form <- ~ age + (1|donor_id) + stimulation + picard_pct_ribosomal_bases + region  + picard_pct_mrna_bases + age + picard_pct_percent_duplication + picard_pct_pf_reads_aligned 

# estimate weights using linear mixed model of dream
#vobjDream = voomWithDreamWeights( genes_counts_cultured, form, metadata_cultured )

# Fit the dream model on each gene
#fit = (dream( vobjDream, form, metadata_cultured ))

#res_age <- data.frame(topTable(fit, coef='age', number=nrow(genes_counts_cultured), sort.by = "p"), check.names = F)
#male age as coefficient
#save(res_age, file ="res_age.Rdata")
```

### DE genes 15% 

```{r DE_15, echo = TRUE}
load("/Users/gijsjesnijders/Documents/MiGASti/Databases/res_age.Rdata")
load("~/Downloads/genes_counts_cultured.Rdata")
gencode_30 = read.table("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
colnames(gencode_30) = c("ensembl","symbol")
res_age <- tibble::rownames_to_column(res_age, "ensembl")
res_name_age <- merge(res_age, gencode_30, by = "ensembl")
sign_age <- subset(res_age, adj.P.Val < 0.15)
length(rownames(sign_age))
```

### DE genes 10%

```{r DE10 , echo = TRUE}
sign_age <- subset(res_name_age, adj.P.Val < 0.10)
length(rownames(sign_age))
```

### DE genes 5%

```{r DE5, echo = TRUE}
sign_age <- subset(res_name_age, adj.P.Val < 0.05)
length(rownames(sign_age))
```


# Plots
### FDR distibution

```{r FDRdistribution, echo = TRUE}
res = res_name_age
p = ggplot(res, aes(P.Value))
p + geom_density(color="darkblue", fill="lightblue") +
theme_classic() +
ggtitle("FDR Distribution")
```

### Fold change distribution

```{r foldchangedistribution, echo = TRUE}
p = ggplot(res, aes(logFC))
p + geom_density(color = "darkblue", fill = "lightblue") +
theme_classic() +
ggtitle("Fold Change Distribution")
```


### MA plot 

```{r MAplot, echo = TRUE}
plot.data = res
plot.data$id = rownames(plot.data)
data = data.frame(plot.data)
data$P.Value = -log10(data$P.Value)
data$fifteen = as.factor(abs(data$adj.P.Val < 0.05))
ma = ggplot(data, aes(AveExpr, logFC, color = fifteen))
ma + geom_point() +
scale_color_manual(values = c("black", "red"), labels = c ("> 0.05", "< 0.05")) +
labs(title = "MA plot", color = "labels") +
theme_classic()
#theme(plot.title = element_text(hjust = 0.5)) + ylim (-10,10) + xlim(-4,22)
```


### Volcano plot 

```{r Volcano, echo = TRUE}
vp = ggplot(data, aes(logFC, P.Value, color = fifteen))
vp + geom_point() +
scale_color_manual(values = c("black", "red"), labels = c("> 0.05", "< 0.05")) +
labs(title = "Gene Level Volcano Plot", color = "FDR") +
#theme(plot.title = element_text(hjust = 0.5)) +
theme_classic() +
xlim(-1,1) + ylim(0, 20) + ylab("-log10 pvalue")
```

# Overlap with MiGA aging genes (input: 1693 genes) 
### list of genes

```{r MiGA, echo = TRUE}
setwd("~/Documents/MiGA/Revision")
MIGA = "~/Documents/MiGA/Revision/AgeMIGA.xlsx"
MIGA = read_excel(MIGA, col_names = TRUE) 
MIGA = as.data.frame(MIGA)
overlap <- merge(sign_age, MIGA, by = "symbol")
list(overlap$symbol)
```

### Is the overlap significant?

```{r MiGA_enriched, echo = TRUE}
set1 <- sign_age
set2 <- MIGA

set1_v <- set1$symbol
set2_v <- set2$symbol

setEnrichment <- function(set1, set2, universe = 20000){

  a = sum(set1 %in% set2)

  c = length(set1) - a

  b = length(set2) - a

  d = universe - length(set2) - c

  contingency_table = matrix(c(a, c, b, d), nrow = 2)

  # one-tailed test for enrichment only

  fisher_results = fisher.test(contingency_table, alternative = "greater")

  # returns data.frame containing the lengths of the two sets, the overlap, the enrichment ratio (odds ratio) and P value

  df <- tibble::tibble( set1_length = length(set1), set2_length = length(set2), overlap = a, ratio = fisher_results$estimate, p.value = fisher_results$p.value)

  return(df)
}

setEnrichment( set1 = set1_v, set2 = set2_v)
```

# Boxplot top 6 genes age per stimulation 
#note. pvalues based on wilcox test of log2(tpm)+1 without additional correction(not based on pvalue DEGs with DREAM)

```{r load2, echo = TRUE, warning=FALSE}
metadata <- read.table("~/Documents/MiGASti/Databases/metadata.txt", quote="\"", comment.char="")
#set rownames to Sample
row.names(metadata) <- metadata$Sample 
#exclude samples that did not pass QC
exclude <- read.csv("~/Documents/MiGASti/Databases/samples2remove.txt", sep="")
exclude <- exclude$x
genes_tpm_filt = genes_tpm[, !colnames(genes_tpm) %in% exclude] 
#Excludes the samples from filters. 
#dim(genes_tpm_filt)
metadata_filt = metadata[ !(rownames(metadata) %in% exclude), ]
length(metadata_filt)
genes_tpm_filt = log2((genes_tpm_filt) + 1)
genes_tpm_filt <- as.data.frame(genes_tpm_filt)
genes_tpm_ordered <- genes_tpm_filt[,rownames(metadata_filt)]
#head(genes_tpm_ordered)
all(rownames(metadata_filt) == colnames (genes_tpm_ordered)) #TRUE
```

### Boxplot top 6 genes LPS

```{r boxplot_age, echo=TRUE, warning=FALSE}
samples_baseline = metadata_filt$Sample[metadata_filt$Stimulation %in% "unstim"] # Includes same donor 
samples_condition = metadata_filt$Sample[metadata_filt$Stimulation %in% "LPS"] # Includes same donor  

metadata4deg = metadata_filt[rownames(metadata_filt) %in% c(samples_baseline,samples_condition) ,]

genes_voom_baseline_condition = as.data.frame(genes_tpm_ordered[, colnames(genes_tpm_ordered) %in% c(samples_baseline, samples_condition)])
# dim(genes_voom_baseline_condition)
# dim(metadata4deg)

deg_lists = res_name_age[order(res_name_age$adj.P.Val) ,]
top_6 = head(deg_lists)
data.table(top_6)

gene2check = as.data.frame(genes_voom_baseline_condition[top_6$ensembl, ])

gene2check$ensembl = rownames(gene2check)
gene2check = merge(gene2check, top_6[, c("symbol", "ensembl")], by = "ensembl")

names(metadata4deg) = tolower(names(metadata4deg))

gene2check_m = melt(gene2check, id.vars = c("ensembl", "symbol"))
gene2check_charac = merge(gene2check_m, metadata4deg, by.x = "variable", by.y = "sample")

ggplot(gene2check_charac, aes(x = stimulation, y = value, fill = stimulation)) +
  geom_boxplot(notch = F, na.rm = T) +
  geom_jitter() +
  theme_bw() + facet_wrap(~symbol, scales = "free_y") +
  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "centre", method = "wilcox.test")
```

### Boxplot top 6 genes IFNy

```{r boxplot_IFNy, echo=TRUE, warning=FALSE}
samples_baseline = metadata_filt$Sample[metadata_filt$Stimulation %in% "unstim"] # Includes same donor 
samples_condition = metadata_filt$Sample[metadata_filt$Stimulation %in% "IFNy"] # Includes same donor  

metadata4deg = metadata_filt[rownames(metadata_filt) %in% c(samples_baseline,samples_condition) ,]

genes_voom_baseline_condition = as.data.frame(genes_tpm_ordered[, colnames(genes_tpm_ordered) %in% c(samples_baseline, samples_condition)])
# dim(genes_voom_baseline_condition)
# dim(metadata4deg)


deg_lists = res_name_age[order(res_name_age$adj.P.Val) ,]
top_6 = head(deg_lists)
data.table(top_6)

gene2check = as.data.frame(genes_voom_baseline_condition[top_6$ensembl, ])

gene2check$ensembl = rownames(gene2check)
gene2check = merge(gene2check, top_6[, c("symbol", "ensembl")], by = "ensembl")

names(metadata4deg) = tolower(names(metadata4deg))

gene2check_m = melt(gene2check, id.vars = c("ensembl", "symbol"))
gene2check_charac = merge(gene2check_m, metadata4deg, by.x = "variable", by.y = "sample")

ggplot(gene2check_charac, aes(x = stimulation, y = value, fill = stimulation)) +
  geom_boxplot(notch = F, na.rm = T) +
  geom_jitter() +
  theme_bw() + facet_wrap(~symbol, scales = "free_y") +
  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "centre", method = "wilcox.test")
```

### Boxplot top 6 genes R848

```{r boxplot_R848, echo=TRUE, warning=FALSE}
samples_baseline = metadata_filt$Sample[metadata_filt$Stimulation %in% "unstim"] # Includes same donor 
samples_condition = metadata_filt$Sample[metadata_filt$Stimulation %in% "R848"] # Includes same donor  

metadata4deg = metadata_filt[rownames(metadata_filt) %in% c(samples_baseline,samples_condition) ,]

genes_voom_baseline_condition = as.data.frame(genes_tpm_ordered[, colnames(genes_tpm_ordered) %in% c(samples_baseline, samples_condition)])
# dim(genes_voom_baseline_condition)
# dim(metadata4deg)


deg_lists = res_name_age[order(res_name_age$adj.P.Val) ,]
top_6 = head(deg_lists)
data.table(top_6)

gene2check = as.data.frame(genes_voom_baseline_condition[top_6$ensembl, ])

gene2check$ensembl = rownames(gene2check)
gene2check = merge(gene2check, top_6[, c("symbol", "ensembl")], by = "ensembl")

names(metadata4deg) = tolower(names(metadata4deg))

gene2check_m = melt(gene2check, id.vars = c("ensembl", "symbol"))
gene2check_charac = merge(gene2check_m, metadata4deg, by.x = "variable", by.y = "sample")

ggplot(gene2check_charac, aes(x = stimulation, y = value, fill = stimulation)) +
  geom_boxplot(notch = F, na.rm = T) +
  geom_jitter() +
  theme_bw() + facet_wrap(~symbol, scales = "free_y") +
  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "centre", method = "wilcox.test")
```


### Boxplot top 6 genes TNFa

```{r boxplot_TNFa, echo=TRUE, warning=FALSE}
samples_baseline = metadata_filt$Sample[metadata_filt$Stimulation %in% "unstim"] # Includes same donor 
samples_condition = metadata_filt$Sample[metadata_filt$Stimulation %in% "TNFa"] # Includes same donor  

metadata4deg = metadata_filt[rownames(metadata_filt) %in% c(samples_baseline,samples_condition) ,]

genes_voom_baseline_condition = as.data.frame(genes_tpm_ordered[, colnames(genes_tpm_ordered) %in% c(samples_baseline, samples_condition)])
# dim(genes_voom_baseline_condition)
# dim(metadata4deg)


deg_lists = res_name_age[order(res_name_age$adj.P.Val) ,]
top_6 = head(deg_lists)
data.table(top_6)

gene2check = as.data.frame(genes_voom_baseline_condition[top_6$ensembl, ])

gene2check$ensembl = rownames(gene2check)
gene2check = merge(gene2check, top_6[, c("symbol", "ensembl")], by = "ensembl")

names(metadata4deg) = tolower(names(metadata4deg))

gene2check_m = melt(gene2check, id.vars = c("ensembl", "symbol"))
gene2check_charac = merge(gene2check_m, metadata4deg, by.x = "variable", by.y = "sample")

ggplot(gene2check_charac, aes(x = stimulation, y = value, fill = stimulation)) +
  geom_boxplot(notch = F, na.rm = T) +
  geom_jitter() +
  theme_bw() + facet_wrap(~symbol, scales = "free_y") +
  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "centre", method = "wilcox.test")
```


### Boxplot top 6 genes DEX

```{r boxplot_DEX, echo=TRUE, warning=FALSE}
samples_baseline = metadata_filt$Sample[metadata_filt$Stimulation %in% "unstim"] # Includes same donor 
samples_condition = metadata_filt$Sample[metadata_filt$Stimulation %in% "DEX"] # Includes same donor  

metadata4deg = metadata_filt[rownames(metadata_filt) %in% c(samples_baseline,samples_condition) ,]

genes_voom_baseline_condition = as.data.frame(genes_tpm_ordered[, colnames(genes_tpm_ordered) %in% c(samples_baseline, samples_condition)])
# dim(genes_voom_baseline_condition)
# dim(metadata4deg)


deg_lists = res_name_age[order(res_name_age$adj.P.Val) ,]
top_6 = head(deg_lists)
data.table(top_6)

gene2check = as.data.frame(genes_voom_baseline_condition[top_6$ensembl, ])

gene2check$ensembl = rownames(gene2check)
gene2check = merge(gene2check, top_6[, c("symbol", "ensembl")], by = "ensembl")

names(metadata4deg) = tolower(names(metadata4deg))

gene2check_m = melt(gene2check, id.vars = c("ensembl", "symbol"))
gene2check_charac = merge(gene2check_m, metadata4deg, by.x = "variable", by.y = "sample")

ggplot(gene2check_charac, aes(x = stimulation, y = value, fill = stimulation)) +
  geom_boxplot(notch = F, na.rm = T) +
  geom_jitter() +
  theme_bw() + facet_wrap(~symbol, scales = "free_y") +
  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "centre", method = "wilcox.test")
```

### Boxplot top 6 genes IL4

```{r boxplot_IL4, echo=TRUE, warning=FALSE}
samples_baseline = metadata_filt$Sample[metadata_filt$Stimulation %in% "unstim"] # Includes same donor 
samples_condition = metadata_filt$Sample[metadata_filt$Stimulation %in% "IL4"] # Includes same donor  

metadata4deg = metadata_filt[rownames(metadata_filt) %in% c(samples_baseline,samples_condition) ,]

genes_voom_baseline_condition = as.data.frame(genes_tpm_ordered[, colnames(genes_tpm_ordered) %in% c(samples_baseline, samples_condition)])
# dim(genes_voom_baseline_condition)
# dim(metadata4deg)


deg_lists = res_name_age[order(res_name_age$adj.P.Val) ,]
top_6 = head(deg_lists)
data.table(top_6)

gene2check = as.data.frame(genes_voom_baseline_condition[top_6$ensembl, ])

gene2check$ensembl = rownames(gene2check)
gene2check = merge(gene2check, top_6[, c("symbol", "ensembl")], by = "ensembl")

names(metadata4deg) = tolower(names(metadata4deg))

gene2check_m = melt(gene2check, id.vars = c("ensembl", "symbol"))
gene2check_charac = merge(gene2check_m, metadata4deg, by.x = "variable", by.y = "sample")

ggplot(gene2check_charac, aes(x = stimulation, y = value, fill = stimulation)) +
  geom_boxplot(notch = F, na.rm = T) +
  geom_jitter() +
  theme_bw() + facet_wrap(~symbol, scales = "free_y") +
  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "centre", method = "wilcox.test")
```


### Boxplot top 6 genes ATP

```{r boxplot_ATP, echo=TRUE, warning=FALSE}
samples_baseline = metadata_filt$Sample[metadata_filt$Stimulation %in% "unstim"] # Includes same donor 
samples_condition = metadata_filt$Sample[metadata_filt$Stimulation %in% "ATP"] # Includes same donor  

metadata4deg = metadata_filt[rownames(metadata_filt) %in% c(samples_baseline,samples_condition) ,]

genes_voom_baseline_condition = as.data.frame(genes_tpm_ordered[, colnames(genes_tpm_ordered) %in% c(samples_baseline, samples_condition)])
# dim(genes_voom_baseline_condition)
# dim(metadata4deg)


deg_lists = res_name_age[order(res_name_age$adj.P.Val) ,]
top_6 = head(deg_lists)
data.table(top_6)

gene2check = as.data.frame(genes_voom_baseline_condition[top_6$ensembl, ])

gene2check$ensembl = rownames(gene2check)
gene2check = merge(gene2check, top_6[, c("symbol", "ensembl")], by = "ensembl")

names(metadata4deg) = tolower(names(metadata4deg))

gene2check_m = melt(gene2check, id.vars = c("ensembl", "symbol"))
gene2check_charac = merge(gene2check_m, metadata4deg, by.x = "variable", by.y = "sample")

ggplot(gene2check_charac, aes(x = stimulation, y = value, fill = stimulation)) +
  geom_boxplot(notch = F, na.rm = T) +
  geom_jitter() +
  theme_bw() + facet_wrap(~symbol, scales = "free_y") +
  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "centre", method = "wilcox.test")
```

# Data table for download 

```{r download, echo = TRUE}
res_age_diff_top = res_name_age[, c("ensembl", "symbol", "logFC", "AveExpr", "t", "P.Value", "adj.P.Val", "z.std")]
createDT(res_age_diff_top)
```

