---
title: "20210224_DEG_scatterplot"
author: "Gijsje"
date: "03/03/2021"
output: 
  rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---


```{r packages, echo=FALSE, message=FALSE, results='hide'}
library(readxl)
library(edgeR)
library(limma)
library(DESeq2)
library(tidyverse)
library(data.table)
library(shiny)
library(pheatmap)
library(ggfortify)
library(ggplot2)
library(UpSetR)
library(VennDiagram)
library(knitr) 
library(tidyverse)
library(dplyr)
library(tidyr)
library(tibble)
library(purrr)
library(corrplot)
library(ggplot2)
library(ggpubr)


knitr::opts_chunk$set( 
  warning=FALSE,
  message=FALSE,
  results = 'asis',
  error = FALSE,
  tidy = FALSE,
  fig.show = "hold")
```

# Do regions respond in the same way to different stimulations?
#Number of samples per stimulation:
#GFM: unstim 38, LPS 35, IFNy 23, R848 17, TNFa 23, ATP 4, DEX 10, IL4 8
#SVZ: unstim 29, LPS 29, IFNy 24, R848 24, TNFa 24, ATP 1 
#GTS: unstim 30, LPS 25, IFNy 12
#THA: unstim 22, LPS 22, IFNy 9
#CC:  unstim 15, LPS 16, IFNy 10

# LPS
### GFM x SVZ
#16108 genes 

```{r scatter_:PS_GFM_SVZ, echo = TRUE}
setwd("~/Documents/MiGASti/Databases")
DEG_LPS_GFM <- read.table("DEG_LPS_GFM.txt")
DEG_LPS_SVZ <- read.table("DEG_LPS_SVZ.txt")
DEG_genes_LPS_GFM_SVZ = merge(DEG_LPS_GFM, DEG_LPS_SVZ, by ="symbol")

GFM <- DEG_genes_LPS_GFM_SVZ[,4]
SVZ <- DEG_genes_LPS_GFM_SVZ[,11]
df = data.frame(GFM, SVZ)
p <- ggplot(df, aes(x=GFM, y=SVZ) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  scale_fill_continuous(type = "viridis") +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  theme_bw()
p + stat_cor(method = "spearman")
```

### CC x SVZ
#16020 genes 

```{r scatter_LPS_CC_SVZ, echo = TRUE}
setwd("~/Documents/MiGASti/Databases")
DEG_LPS_CC <- read.table("DEG_LPS_CC.txt")
DEG_LPS_SVZ <- read.table("DEG_LPS_SVZ.txt")
DEG_genes_LPS_CC_SVZ = merge(DEG_LPS_CC, DEG_LPS_SVZ, by ="symbol")

CC <- DEG_genes_LPS_CC_SVZ[,4]
SVZ <- DEG_genes_LPS_CC_SVZ[,11]
df = data.frame(CC, SVZ)
p <- ggplot(df, aes(x=CC, y=SVZ) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  scale_fill_continuous(type = "viridis") +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  theme_bw()
p + stat_cor(method = "spearman")
```

### GTS x THA
#21558 genes 

```{r scatter_LPS_GTS_THA, echo = TRUE}
setwd("~/Documents/MiGASti/Databases")
DEG_LPS_THA <- read.table("DEG_LPS_THA.txt")
DEG_LPS_GTS <- read.table("DEG_LPS_GTS.txt")
DEG_genes_LPS_GTS_THA = merge(DEG_LPS_THA, DEG_LPS_GTS, by ="symbol")

THA <- DEG_genes_LPS_GTS_THA[,4]
GTS <- DEG_genes_LPS_GTS_THA[,11]
df = data.frame(THA, GTS)
p <- ggplot(df, aes(x=GTS, y=THA) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  scale_fill_continuous(type = "viridis") +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  theme_bw()
p + stat_cor(method = "spearman")
```

### GFM x CC 
#20493 genes 

```{r scatter_LPS_GFM_CC, echo = TRUE}
setwd("~/Documents/MiGASti/Databases")
DEG_LPS_GFM <- read.table("DEG_LPS_GFM.txt")
DEG_LPS_CC <- read.table("DEG_LPS_CC.txt")
DEG_genes_LPS_GFM_CC = merge(DEG_LPS_GFM, DEG_LPS_CC, by ="symbol")

GFM <- DEG_genes_LPS_GFM_CC[,4]
CC <- DEG_genes_LPS_GFM_CC[,11]
df = data.frame(GFM, CC)
p <- ggplot(df, aes(x=GFM, y=CC) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  scale_fill_continuous(type = "viridis") +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  theme_bw()
p + stat_cor(method = "spearman")
```

### GFM x GTS 
#21345 genes 

```{r scatter_LPS_GFM_GTS, echo = TRUE}
setwd("~/Documents/MiGASti/Databases")
DEG_LPS_GFM <- read.table("DEG_LPS_GFM.txt")
DEG_LPS_GTS <- read.table("DEG_LPS_GTS.txt")
DEG_genes_LPS_GFM_GTS = merge(DEG_LPS_GFM, DEG_LPS_GTS, by ="symbol")

GFM <- DEG_genes_LPS_GFM_GTS[,4]
GTS <- DEG_genes_LPS_GFM_GTS[,11]
df = data.frame(GFM, GTS)
p <- ggplot(df, aes(x=GFM, y=GTS) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  scale_fill_continuous(type = "viridis") +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  theme_bw()
p + stat_cor(method = "spearman")
```


### GFM x THA 
#20900 genes 

```{r scatter_LPS_GFM_THA, echo = TRUE}
setwd("~/Documents/MiGASti/Databases")
DEG_LPS_GFM <- read.table("DEG_LPS_GFM.txt")
DEG_LPS_THA <- read.table("DEG_LPS_THA.txt")
DEG_genes_LPS_GFM_THA = merge(DEG_LPS_GFM, DEG_LPS_THA, by ="symbol")

GFM <- DEG_genes_LPS_GFM_THA[,4]
THA <- DEG_genes_LPS_GFM_THA[,11]
df = data.frame(GFM, THA)
p <- ggplot(df, aes(x=GFM, y=THA) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  scale_fill_continuous(type = "viridis") +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  theme_bw()
p + stat_cor(method = "spearman")
```


# IFNy
### GFM x SVZ
#16108 genes 

```{r scatter_IFNy_GFM_SVZ, echo = TRUE, results = 'hide'}
setwd("~/Documents/MiGASti/Databases")
DEG_IFNy_GFM <- read.table("DEG_IFNy_GFM.txt")
DEG_IFNy_SVZ <- read.table("DEG_IFNy_SVZ.txt")
DEG_genes_IFNy_GFM_SVZ = merge(DEG_IFNy_GFM, DEG_IFNy_SVZ, by ="symbol")

GFM <- DEG_genes_IFNy_GFM_SVZ[,4]
SVZ <- DEG_genes_IFNy_GFM_SVZ[,11]
df = data.frame(GFM, SVZ)
p <- ggplot(df, aes(x=GFM, y=SVZ) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  scale_fill_continuous(type = "viridis") +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  theme_bw()
p + stat_cor(method = "spearman")
```

### GFM x GTS 
#21345 genes 

```{r scatter_LPS_GFM_SVZ, echo = TRUE, results = 'hide'}
setwd("~/Documents/MiGASti/Databases")
DEG_IFNy_GFM <- read.table("DEG_IFNy_GFM.txt")
DEG_IFNy_GTS <- read.table("DEG_IFNy_GTS.txt")
DEG_genes_IFNy_GFM_GTS = merge(DEG_IFNy_GFM, DEG_IFNy_GTS, by ="symbol")

GFM <- DEG_genes_IFNy_GFM_GTS[,4]
GTS <- DEG_genes_IFNy_GFM_GTS[,11]
df = data.frame(GFM, GTS)
p <- ggplot(df, aes(x=GFM, y=GTS) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  scale_fill_continuous(type = "viridis") +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  theme_bw()
p + stat_cor(method = "spearman")
```

### GFM x THA 
#20900 genes 

```{r scatter_IFNy_GFM_THA, echo = TRUE}
setwd("~/Documents/MiGASti/Databases")
DEG_IFNy_GFM <- read.table("DEG_IFNy_GFM.txt")
DEG_IFNy_THA <- read.table("DEG_IFNy_THA.txt")
DEG_genes_IFNy_GFM_THA = merge(DEG_IFNy_GFM, DEG_IFNy_THA, by ="symbol")

GFM <- DEG_genes_IFNy_GFM_THA[,4]
THA <- DEG_genes_IFNy_GFM_THA[,11]
df = data.frame(GFM, THA)
p <- ggplot(df, aes(x=GFM, y=THA) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  scale_fill_continuous(type = "viridis") +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  theme_bw()
p + stat_cor(method = "spearman")
```

### GFM x CC 
#20493 genes 

```{r scatter_IFNy_GFM_CC, echo = TRUE}
setwd("~/Documents/MiGASti/Databases")
DEG_IFNy_GFM <- read.table("DEG_IFNy_GFM.txt")
DEG_IFNy_CC <- read.table("DEG_IFNy_CC.txt")
DEG_genes_IFNy_GFM_CC = merge(DEG_IFNy_GFM, DEG_IFNy_CC, by ="symbol")

GFM <- DEG_genes_IFNy_GFM_CC[,4]
CC <- DEG_genes_IFNy_GFM_CC[,11]
df = data.frame(GFM, CC)
p <- ggplot(df, aes(x=GFM, y=CC) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  scale_fill_continuous(type = "viridis") +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  theme_bw()
p + stat_cor(method = "spearman")
```

# TNFa
### GFM x SVZ
#16108 genes 

```{r scatter_TNFa_GFM_SVZ, echo = TRUE, results = 'hide'}
setwd("~/Documents/MiGASti/Databases")
DEG_TNFa_GFM <- read.table("DEG_TNFa_GFM.txt")
DEG_TNFa_SVZ <- read.table("DEG_TNFa_SVZ.txt")
DEG_genes_TNFa_GFM_SVZ = merge(DEG_TNFa_GFM, DEG_TNFa_SVZ, by ="symbol")

GFM <- DEG_genes_TNFa_GFM_SVZ[,4]
SVZ <- DEG_genes_TNFa_GFM_SVZ[,11]
df = data.frame(GFM, SVZ)
p <- ggplot(df, aes(x=GFM, y=SVZ) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  scale_fill_continuous(type = "viridis") +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  theme_bw()
p + stat_cor(method = "spearman")
```

# R848
### GFM x SVZ
#16108 genes

```{r scatter_R848_GFM_SVZ, echo = TRUE, results = 'hide'}
setwd("~/Documents/MiGASti/Databases")
DEG_R848_GFM <- read.table("DEG_R848_GFM.txt")
DEG_R848_SVZ <- read.table("DEG_R848_SVZ.txt")
DEG_genes_R848_GFM_SVZ = merge(DEG_R848_GFM, DEG_R848_SVZ, by ="symbol")

GFM <- DEG_genes_R848_GFM_SVZ[,4]
SVZ <- DEG_genes_R848_GFM_SVZ[,11]
df = data.frame(GFM, SVZ)
p <- ggplot(df, aes(x=GFM, y=SVZ) ) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_bin2d(bins = 100) +
  scale_fill_continuous(type = "viridis") +
  geom_abline(slope = 1, intercept = 0, linetype = 3) +
  theme_bw()
p + stat_cor(method = "spearman")
```















