---
title: "age_stimulation_plots"
author: "Gijsje"
date: "5/13/2021"
output: 
 rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---


```{r setup, include = FALSE}
library('variancePartition')
library('edgeR')
library(dplyr)
library(tidyr)
library(rtracklayer)
library(ggplot2)
library(kableExtra)
library(data.table)
library(ggplot2)
library(tidyverse)
library(readxl)
library(pheatmap)
library(reshape2)
library(kableExtra)
library(RColorBrewer)
require(gridExtra)
library(ggfortify)
library(dplyr)
library(tidyr)
library(tidyverse)
library(circlize)
library(tidyverse)
library(readxl)
library(edgeR)
library(limma)
library(DESeq2)

```

# Load data

```{r data_load, echo=TRUE, results='hide'}
metadata <- read.table("~/Documents/MiGASti/Databases/metadata.txt")
#set rownames to Sample
row.names(metadata) <- metadata$Sample 
setwd("~/Documents/MiGASti/Databases")
#exclude samples that did not pass QC filtering
exclude <- read.table("samples2remove.txt")
exclude <- exclude$x
#Excludes the samples from filters. 
#dim(genes_counts_filt)
metadata_filt = metadata[ !(rownames(metadata) %in% exclude), ]
length(metadata_filt)

#load gene matrix
load("~/Documents/MiGASti/Databases/gene_matrix.RData")
genes_counts_filt = genes_counts[, !colnames(genes_counts) %in% exclude] 
#Excludes the samples from filters. 
#dim(genes_counts_filt)
metadata_filt = metadata[ !(rownames(metadata) %in% exclude), ]
length(metadata_filt)
#remove low count genes
cpm <- cpm(genes_counts_filt) 
# CPM >= 1 in at least 50% of the samples
keep.exp <- rowSums(cpm > 1) >= (0.5 * ncol(genes_counts_filt) )
genes_counts_filt1 <- genes_counts_filt[ keep.exp, ] #18997 genes 
#order metadata and genes counts
genes_counts_ordered <- genes_counts_filt1[,rownames(metadata_filt)]
#head(genes_counts_ordered)
all(rownames(metadata_filt) == colnames (genes_counts_ordered)) #TRUE
```


# Metadata: Distribution of age in LPS and IFNy


```{r distribution, echo=TRUE}
#remove uncultured samples
metadata_cultured <- metadata_filt[metadata_filt$Stimulation != "ununstim",]
#check numbers
#dim(metadata_cultured)
#check numbers per stimulation
#table(metadata_filt$Stimulation)
#select only GFM samples in metadata
metadata_unstim = metadata_cultured[metadata_cultured$Stimulation=="unstim",]
metadata_LPS1 = metadata_cultured[metadata_cultured$Stimulation=="LPS",]
metadata_IFNy1 = metadata_cultured[metadata_cultured$Stimulation=="IFNy",]

set.seed(1)
LPS=metadata_LPS1$age
IFNy=metadata_IFNy1$age
 
# LPS distribution
hist(LPS, breaks=30, xlim=c(0,120), col=rgb(1,0,0,0.5), xlab="age", 
     ylab="samples", main="Distribution of age" )
# Add legend
legend("topright", legend=c("LPS"), col=c(rgb(1,0,0,0.5), 
     rgb(0,0,1,0.5)), pt.cex=2, pch=15 )

# IFNy
hist(IFNy, breaks=30, xlim=c(0,120), col=rgb(0,0,1,0.5), xlab="age", 
     ylab="samples", main="Distribution of age" )
# Add legend
legend("topright", legend=c("IFNy"), col=c(rgb(0,0,1,0.5), 
     rgb(0,0,1,0.5)), pt.cex=2, pch=15 )


# Add legend
legend("topright", legend=c("LPS","IFNy"), col=c(rgb(1,0,0,0.5), 
     rgb(0,0,1,0.5)), pt.cex=2, pch=15 )
```

# Aging in samples that are stimulated with LPS/IFNy? Does the immune response change in aging? 

### Input: colMeans of the 6270 genes that respond to LPS in 128 LPS stimulated samples across aging

```{r LPS, echo=TRUE, results='hide'}
#remove ununstim samples in metadata
metadata_cultured <- metadata_filt[metadata_filt$Stimulation != "ununstim",]
#check numbers
dim(metadata_cultured)
#check numbers per stimulation
table(metadata_filt$Stimulation)
#remove samples in genes counts datafile 
genes_counts_cultured <- genes_counts_ordered[,metadata_cultured$Sample]
voom(genes_counts_cultured)
#table(metadata_filt$stimulation)
#length(metadata_cultured)

genes_counts_cultured <- cbind(ensembl = rownames(genes_counts_cultured), genes_counts_cultured)
genes_counts_cultured
rownames(genes_counts_cultured) <- NULL
```

```{r LPS2, echo = TRUE, results = 'hide'}
#data load 
load("~/Documents/MiGASti/docs/res_name_LPS2.Rdata")
load("~/Documents/MiGASti/docs/res_name_IFNy2.Rdata")
DEG_LPS <- subset(res_name_LPS2, adj.P.Val < 0.05)
DEG_IFNy <- subset(res_name_IFNy2, adj.P.Val < 0.05)

# select only sign LPS genes 
LPS <- genes_counts_cultured[genes_counts_cultured$ensembl %in% DEG_LPS$ensembl, ]
#select only LPS samples 
metadata_LPS = metadata_cultured[metadata_cultured$Stimulation=="LPS",]
#select only LPS samples in genes counts
LPS2 <-LPS[,metadata_LPS$Sample]
#order metadata and genes_counts
LPS_ordered <- LPS2[,rownames(metadata_LPS)]

#select only sign IFNy genes
IFNy <- genes_counts_cultured[genes_counts_cultured$ensembl %in% DEG_IFNy$ensembl, ]
#select only IFNy samples
metadata_IFNy = metadata_cultured[metadata_cultured$Stimulation=="IFNy",]
#select only IFNy samples in genes counts
IFNy2 <-IFNy[,metadata_IFNy$Sample]
#order metadata and genes_counts
IFNy_ordered <- IFNy[,rownames(metadata_IFNy)]

#create plot for LPS
LPS.mic <- data.frame(cbind(LPS = scale(colMeans(LPS_ordered)), Age = metadata_LPS$age))
colnames(LPS.mic) <- c("LPS", "Age")
library(ggpmisc)
formula <- y ~ x
ggplot(LPS.mic, aes_string(y = as.matrix(LPS.mic["LPS"]), x = as.matrix(LPS.mic["Age"]))) + 
    geom_point() + scale_colour_gradient(low = "gray90", high = "black") + stat_smooth(method = lm) + 
    theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_smooth(method = lm, 
    linetype = "dashed") + labs(y = "Average expression LPS response genes", x = "Age") + 
    stat_poly_eq(aes(label = c(paste(..adj.rr.label..))), label.x = "right", label.y = 0.37, 
        formula = formula, parse = TRUE, size = 3) + # geom_cor(method='pearson')+
stat_fit_glance(method = "lm", method.args = list(formula = formula), geom = "text", 
    aes(label = paste("P-value = ", signif(..p.value.., digits = 3), sep = "")), 
    label.x = "right", label.y = 0.3, size = 3)
```


### Input: colMeans of the 658 genes that respond to IFNy in 79 IFNy stimulated samples across aging

```{r IFNy, echo = TRUE}
IFNy.mic <- data.frame(cbind(IFNy = scale(colMeans(IFNy_ordered)), Age = metadata_IFNy$age))
colnames(IFNy.mic) <- c("IFNy", "Age")
library(ggpmisc)
formula <- y ~ x
ggplot(IFNy.mic, aes_string(y = as.matrix(IFNy.mic["IFNy"]), x = as.matrix(IFNy.mic["Age"]))) + 
    geom_point() + scale_colour_gradient(low = "gray90", high = "black") + stat_smooth(method = lm) + 
    theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_smooth(method = lm, 
    linetype = "dashed") + labs(y = "Average expression IFNy response genes", x = "Age") + 
    stat_poly_eq(aes(label = c(paste(..adj.rr.label..))), label.x = "right", label.y = 0.37, 
        formula = formula, parse = TRUE, size = 3) + # geom_cor(method='pearson')+
stat_fit_glance(method = "lm", method.args = list(formula = formula), geom = "text", 
    aes(label = paste("P-value = ", signif(..p.value.., digits = 3), sep = "")), 
    label.x = "right", label.y = 0.3, size = 3)
```

# How many of the age related genes are in the stimulation genes? 
#151 genes are both in aging dataset and LPS stimulation dataset
#17 genes are both in aging dataset and IFNy stimulation dataset

```{r overlapbetweensets, echo = TRUE}
load("/Users/gijsjesnijders/Documents/MiGASti/Databases/res_age.Rdata")
load("~/Downloads/genes_counts_cultured.Rdata")
gencode_30 = read.table("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
colnames(gencode_30) = c("ensembl","symbol")
res_age <- tibble::rownames_to_column(res_age, "ensembl")
res_name_age <- merge(res_age, gencode_30, by = "ensembl")
sign_age <- subset(res_age, adj.P.Val < 0.05)
length(rownames(sign_age))
load("~/Documents/MiGASti/docs/res_name_LPS2.Rdata")
load("~/Documents/MiGASti/docs/res_name_IFNy2.Rdata")
LPS2_overlap <- merge(DEG_LPS, sign_age, by = "ensembl")
length(LPS2_overlap$ensembl)
IFNy2_overlap <- merge(DEG_IFNy, sign_age, by = "ensembl")
length(IFNy2_overlap$ensembl)
list(LPS2_overlap$symbol)
list(IFNy2_overlap$symbol)
```


### Aging in samples that are stimulated with LPS and known to show effect in aging and after LPS stimulation? Does the immune response change in aging? 
#Input: colMeans of the 151 genes that respond to LPS and are found to be differentially expressed in aging in 128 LPS stimulated samples across aging
 

```{r LPS_plot1, echo = TRUE, results = 'hide'}
#remove ununstim samples in metadata
metadata_cultured <- metadata_filt[metadata_filt$Stimulation != "ununstim",]
#check numbers
dim(metadata_cultured)
#check numbers per stimulation
table(metadata_filt$Stimulation)
#remove samples in genes counts datafile 
genes_counts_cultured <- genes_counts_ordered[,metadata_cultured$Sample]
voom(genes_counts_cultured)
#table(metadata_filt$stimulation)
#length(metadata_cultured)


genes_counts_cultured <- cbind(ensembl = rownames(genes_counts_cultured), genes_counts_cultured)
genes_counts_cultured
rownames(genes_counts_cultured) <- NULL
```

```{r LPS_plot22, echo = TRUE}
LPS <- genes_counts_cultured[genes_counts_cultured$ensembl %in% LPS2_overlap$ensembl, ]
#select only LPS samples 
metadata_LPS = metadata_cultured[metadata_cultured$Stimulation=="LPS",]
#select only LPS samples in genes counts
LPS2 <-LPS[,metadata_LPS$Sample]
#order metadata and genes_counts
LPS_ordered <- LPS[,rownames(metadata_LPS)]

LPS.mic <- data.frame(cbind(LPS = scale(colMeans(LPS_ordered)), Age = metadata_LPS$age))
colnames(LPS.mic) <- c("LPS", "Age")
library(ggpmisc)
formula <- y ~ x
ggplot(LPS.mic, aes_string(y = as.matrix(LPS.mic["LPS"]), x = as.matrix(LPS.mic["Age"]))) + 
    geom_point() + scale_colour_gradient(low = "gray90", high = "black") + stat_smooth(method = lm) + 
    theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_smooth(method = lm, 
    linetype = "dashed") + labs(y = "Average expression LPS response genes", x = "Age") + 
    stat_poly_eq(aes(label = c(paste(..adj.rr.label..))), label.x = "right", label.y = 0.37, 
        formula = formula, parse = TRUE, size = 3) + # geom_cor(method='pearson')+
stat_fit_glance(method = "lm", method.args = list(formula = formula), geom = "text", 
    aes(label = paste("P-value = ", signif(..p.value.., digits = 3), sep = "")), 
    label.x = "right", label.y = 0.3, size = 3)
```


### Aging in samples that are stimulated with IFNy and known to show effect in aging and after IFNy stimulation? Does the immune response change in aging? 
#Input: colMeans of the 17 genes that respond to IFNy and are found to be differentially expressed in aging in 79 LPS stimulated samples across aging

```{r 58genes_plot, echo = TRUE, results = 'hide'}
#remove ununstim samples in metadata
metadata_cultured <- metadata_filt[metadata_filt$Stimulation != "ununstim",]
#check numbers
dim(metadata_cultured)
#check numbers per stimulation
table(metadata_filt$Stimulation)
#remove samples in genes counts datafile 
genes_counts_cultured <- genes_counts_ordered[,metadata_cultured$Sample]
voom(genes_counts_cultured)
#table(metadata_filt$stimulation)
#length(metadata_cultured)


genes_counts_cultured <- cbind(ensembl = rownames(genes_counts_cultured), genes_counts_cultured)
genes_counts_cultured
rownames(genes_counts_cultured) <- NULL
```

```{r 17genes_plot, echo = TRUE, results = 'hide'}
IFNy <- genes_counts_cultured[genes_counts_cultured$ensembl %in% IFNy2_overlap$ensembl, ]
#select only IFNy samples
metadata_IFNy = metadata_cultured[metadata_cultured$Stimulation=="IFNy",]
#select only IFNy samples in genes counts
IFNy2 <-IFNy[,metadata_IFNy$Sample]
#order metadata and genes_counts
IFNy_ordered <- IFNy[,rownames(metadata_IFNy)]

IFNy.mic <- data.frame(cbind(IFNy = scale(colMeans(IFNy_ordered)), Age = metadata_IFNy$age))
colnames(IFNy.mic) <- c("IFNy", "Age")
library(ggpmisc)
formula <- y ~ x
ggplot(IFNy.mic, aes_string(y = as.matrix(IFNy.mic["IFNy"]), x = as.matrix(IFNy.mic["Age"]))) + 
    geom_point() + scale_colour_gradient(low = "gray90", high = "black") + stat_smooth(method = lm) + 
    theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_smooth(method = lm, 
    linetype = "dashed") + labs(y = "Average expression IFNy response genes", x = "Age") + 
    stat_poly_eq(aes(label = c(paste(..adj.rr.label..))), label.x = "right", label.y = 0.37, 
        formula = formula, parse = TRUE, size = 3) + # geom_cor(method='pearson')+
stat_fit_glance(method = "lm", method.args = list(formula = formula), geom = "text", 
    aes(label = paste("P-value = ", signif(..p.value.., digits = 3), sep = "")), 
    label.x = "right", label.y = 0.3, size = 3)
```


