---
title: "Markers stimulation"
author: "Gijsje"
date: "2/16/2021"
output: 
 rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---

```{r load.packages, echo=FALSE, message=FALSE, results='hide'}
library(readxl)
library(edgeR)
library(limma)
library(DESeq2)
library(tidyverse)
library(data.table)


knitr::opts_chunk$set( 
  warning=FALSE,
  message=FALSE,
  results = 'asis',
  error = FALSE,
  tidy = FALSE,
  fig.show = "hold")
```


```{r check, echo = TRUE, results = 'hide'}
load("~/Documents/MiGASti/Databases/gene_matrix.RData")
metadata <- read.table("~/Documents/MiGASti/Databases/metadata.txt")
#set rownames to Sample
row.names(metadata) <- metadata$Sample 
setwd("~/Documents/MiGASti/Databases")
exclude <- read.table("samples2remove.txt")
exclude <- exclude$x
genes_tpm_filt = genes_tpm[, !colnames(genes_tpm) %in% exclude] 
#Excludes the samples from filters. 
#dim(genes_tpm_filt)
metadata_filt = metadata[ !(rownames(metadata) %in% exclude), ]
length(metadata_filt)
#check for expression of markers in dataset
markers = "~/Documents/MiGASti/Databases/Markers_stim.xlsx"
markers = read_excel(markers, col_names = TRUE) 
markers = as.data.frame(markers)
gencode_30 = read.table("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
colnames(gencode_30) = c("ensembl","symbol")
genes_tpm_filt = log2((genes_tpm_filt) + 1)
genes_tpm_filt <- as.data.frame(genes_tpm_filt)
setDT(genes_tpm_filt, keep.rownames = "ensembl")
genes_tpm_filt$ensembl
res_name = merge(genes_tpm_filt, gencode_30, by="ensembl")
rownames(res_name) = res_name$ensembl
```


```{r data_load_markers, echo = TRUE, results = 'hide'}
marker_expression = merge(res_name, markers, by ="symbol")
dim (marker_expression)
head (marker_expression)
metadata_filt_ordered <- metadata_filt[colnames(genes_tpm_filt),]
metadata_filt_ordered = metadata_filt_ordered[-1,]
all(rownames(metadata_filt_ordered) == colnames (genes_tpm_filt)) #TRUE
Stimulation <- metadata_filt_ordered$Stimulation
```

# Inflammatory markers (LPS)
### IL1B 
#expected to be increased in LPS

```{r data_load_IL1B, echo = TRUE}
IL1B <- marker_expression[13, ] 
IL1B$ensembl = NULL
IL1B$symbol = NULL
IL1B_new<-as.data.frame(t(IL1B))
df = data.frame(Stimulation, IL1B_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("IL1B") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### IL6
#expected to be increased in LPS

```{r data_load_IL6, echo = TRUE}
IL6 <- marker_expression[14, ] 
IL6$ensembl = NULL
IL6$symbol = NULL
IL6_new<-as.data.frame(t(IL6))
df = data.frame(Stimulation, IL6_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("IL6") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### TNF
#expected to be increased in LPS

```{r data_load_TNF, echo = TRUE}
TNF <- marker_expression[25, ] 
TNF$ensembl = NULL
TNF$symbol = NULL
TNF_new<-as.data.frame(t(TNF))
df = data.frame(Stimulation, TNF_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot() + 
  ggtitle("TNF") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

# Inflammatory markers (IFNy)

### CXCL9 
#expected to be increased in IFNy

```{r data_load_CXCL9, echo = TRUE}
CXCL9 <- marker_expression[9, ] 
CXCL9$ensembl = NULL
CXCL9$symbol = NULL
CXCL9_new<-as.data.frame(t(CXCL9))
df = data.frame(Stimulation, CXCL9_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot() + 
  ggtitle("CXCL9") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### CXCL10
#expected to be increased in IFNy

```{r data_load_CXCL10, echo = TRUE}
CXCL10 <- marker_expression[7, ] 
CXCL10$ensembl = NULL
CXCL10$symbol = NULL
CXCL10_new<-as.data.frame(t(CXCL10))
df = data.frame(Stimulation, CXCL10_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot() +
ggtitle("CXCL10") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### CXCL11
#expected to be increased in IFNy

```{r data_load_CXCL11, echo = TRUE}
CXCL11 <- marker_expression[8, ] 
CXCL11$ensembl = NULL
CXCL11$symbol = NULL
CXCL11_new<-as.data.frame(t(CXCL11))
df = data.frame(Stimulation, CXCL11_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot() + 
  ggtitle("CXCL11") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

# Inflammatory markers (TNF)

### CCL2
#expected to be increased in TNF

```{r data_load_CCL2, echo = TRUE}
CCL2 <- marker_expression[3, ] 
CCL2$ensembl = NULL
CCL2$symbol = NULL
CCL2_new<-as.data.frame(t(CCL2))
df = data.frame(Stimulation, CCL2_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("CCL2") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### VCAM1
#expected to be increased in TNF

```{r data_load_VCAM1, echo = TRUE}
VCAM1 <- marker_expression[26, ] 
VCAM1$ensembl = NULL
VCAM1$symbol = NULL
VCAM1_new<-as.data.frame(t(VCAM1))
df = data.frame(Stimulation, VCAM1_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot() + 
  ggtitle("VCAM1") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

# Inflammatory markers (R848)

### SAA3P
##expected to be MyD88 dependent based on Lin et al. 2017

```{r data_load_SAA3P, echo = TRUE}
SAA3P <- marker_expression[22, ] 
SAA3P$ensembl = NULL
SAA3P$symbol = NULL
SAA3P_new<-as.data.frame(t(SAA3P))
df = data.frame(Stimulation, SAA3P_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("SAA3P") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### SLC13A3
#expected to be MyD88 dependent based on Lin et al. 2017

```{r data_load_SLC13A3, echo = TRUE}
SLC13A3 <- marker_expression[23, ] 
SLC13A3$ensembl = NULL
SLC13A3$symbol = NULL
SLC13A3_new<-as.data.frame(t(SLC13A3))
df = data.frame(Stimulation, SLC13A3_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("SLC13A3") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### CXCL1
#expected to be MyD88 dependent based on Lin et al. 2017

```{r data_load_CXCL1, echo = TRUE}
CXCL1 <- marker_expression[6, ] 
CXCL1$ensembl = NULL
CXCL1$symbol = NULL
CXCL1_new<-as.data.frame(t(CXCL1))
df = data.frame(Stimulation, CXCL1_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("CXCL1") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

# Anti-inflammatory markers (IL-4/dexamethasone)

### NR3C1 
#expected to be reduced: Breen et al. 

```{r data_load_NR3C1, echo = TRUE}
NR3C1 <- marker_expression[17, ] 
NR3C1$ensembl = NULL
NR3C1$symbol = NULL
NR3C1_new<-as.data.frame(t(NR3C1))
df = data.frame(Stimulation, NR3C1_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot() +
  ggtitle("NR3C1") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### NR3C2 
#expected to be reduced: Breen et al. 

```{r data_load_NR3C2, echo = TRUE}
NR3C2 <- marker_expression[18, ] 
NR3C2$ensembl = NULL
NR3C2$symbol = NULL
NR3C2_new<-as.data.frame(t(NR3C2))
df = data.frame(Stimulation, NR3C2_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot() + 
  ggtitle("NR3C2") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

# Anti-inflammatory (dexamethasone) 
### FKBP5
#expected to be increased in Dex

```{r data_load_FKBP5, echo = FALSE}
FKBP5 <- marker_expression[11, ] 
FKBP5$ensembl = NULL
FKBP5$symbol = NULL
FKBP5_new<-as.data.frame(t(FKBP5))
df = data.frame(Stimulation, FKBP5_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("FKBP5") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### CD163
#expected to be increased in Dex/Il4

```{r data_load_CD163, echo = TRUE}
CD163 <- marker_expression[4, ] 
CD163$ensembl = NULL
CD163$symbol = NULL
CD163_new<-as.data.frame(t(CD163))
df = data.frame(Stimulation, CD163_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot() +
  ggtitle("CD163") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### CD206
#expected to be increased in dex/il4

```{r data_load_CD206, echo = TRUE}
CD206 <- marker_expression[16, ] 
CD206$ensembl = NULL
CD206$symbol = NULL
CD206_new<-as.data.frame(t(CD206))
df = data.frame(Stimulation, CD206_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot() +
  ggtitle("CD206") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

# Anti-inflammatory (IL-4)

### IL10
#expected to be increased in IL4 based on Li et al. 2018 

```{r data_load_IL10, echo = TRUE}
IL10 <- marker_expression[12, ] 
IL10$ensembl = NULL
IL10$symbol = NULL
IL10_new<-as.data.frame(t(IL10))
df = data.frame(Stimulation, IL10_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot() +
  ggtitle("IL10") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### RETNLB (FIZZ1)
#expected to be increased in IL4 based on Li et al. 2018

```{r data_load_FIZZ1, echo = TRUE}
RETNLB <- marker_expression[21, ] 
RETNLB$ensembl = NULL
RETNLB$symbol = NULL
RETNLB_new<-as.data.frame(t(RETNLB))
df = data.frame(Stimulation, RETNLB_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("RETNLB") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### ARG1
#expected to be increased in IL4 based on Li et al. 2018 

```{r data_load_ARG1, echo = TRUE, results = 'hide'}
ARG1 <- marker_expression[1, ] 
ARG1$ensembl = NULL
ARG1$symbol = NULL
ARG1_new<-as.data.frame(t(ARG1))
df = data.frame(Stimulation, ARG1_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("ARG1") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### ENTPD1

```{r data_load_ENTPD1, echo = TRUE}
ENTPD1 <- marker_expression[10, ] 
ENTPD1$ensembl = NULL
ENTPD1$symbol = NULL
ENTPD1_new<-as.data.frame(t(ENTPD1))
df = data.frame(Stimulation, ENTPD1_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot() +
  ggtitle("ENTPD1") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```


# Other markers (ATP)

### P2RY4
#purinergic receptors increase after ATP stimulation

```{r data_load_ATP, echo = TRUE}
P2RY4 <- marker_expression[20, ] 
P2RY4$ensembl = NULL
P2RY4$symbol = NULL
P2RY4_new<-as.data.frame(t(P2RY4))
df = data.frame(Stimulation, P2RY4_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("P2RY4") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### P2RX7
#purinergic receptors increase after ATP stimulation

```{r data_load_P2RX7, echo = TRUE}
P2RX7 <- marker_expression[19, ] 
P2RX7$ensembl = NULL
P2RX7$symbol = NULL
P2RX7_new<-as.data.frame(t(P2RX7))
df = data.frame(Stimulation, P2RX7_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("P2RX7") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### TMEM163 
#expected to be increased after ATP stimulation 

```{r data_load_TMEM163, echo = TRUE}
TMEM163 <- marker_expression[24, ] 
TMEM163$ensembl = NULL
TMEM163$symbol = NULL
TMEM163_new<-as.data.frame(t(TMEM163))
df = data.frame(Stimulation, TMEM163_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("TMEM163") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

# Homeostatic markers 

###  MAFB: 
#expected to be increased in culture based on Gosselin 2017

```{r data_load_MAFB, echo = TRUE}
MAFB <- marker_expression[15, ] 
MAFB$ensembl = NULL
MAFB$symbol = NULL
MAFB_new<-as.data.frame(t(MAFB))
df = data.frame(Stimulation, MAFB_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("MAFB") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### CX3CR1 
#expected to be reduced in culture based on review Roy/Frederieke

```{r data_load_CX3CR1, echo = TRUE}
CX3CR1 <- marker_expression[5, ] 
CX3CR1$ensembl = NULL
CX3CR1$symbol = NULL
CX3CR1_new<-as.data.frame(t(CX3CR1))
df = data.frame(Stimulation, CX3CR1_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("CX3CR1") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

### C3 (down)
#expected to be reduced in culture based on review Roy/Frederieke

```{r data_load_C3, echo = TRUE}
C3 <- marker_expression[2, ] 
C3$ensembl = NULL
C3$symbol = NULL
C3_new<-as.data.frame(t(C3))
df = data.frame(Stimulation, C3_new)
ggplot(data = df, mapping = aes(x = Stimulation, y = V1)) +
  geom_boxplot()+ 
  ggtitle("C3") +
  xlab("Stimulation") + ylab("log2((TPM)+1)")
```

# Heatmap with all markers 


```{r data_load_heatmap, echo = TRUE, fig.width=12, fig.height=8, dpi=300}
marker_expression$ensembl = NULL
marker_expression$symbol = NULL
marker_expression = as.matrix(marker_expression)
heatmap(marker_expression)
```










