---
title: "20210224_DEG_heatmap"
author: "Gijsje"
date: "2/24/2021"
output: 
  rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---


```{r packages, echo=FALSE, message=FALSE, results='hide'}
library(readxl)
library(edgeR)
library(limma)
library(DESeq2)
library(tidyverse)
library(data.table)
library(shiny)
library(DT)
library(pheatmap)
library(ggfortify)
library(ggplot2)
library(UpSetR)
library(VennDiagram)
library(knitr) 
library(tidyverse)
library(dplyr)
library(tidyr)
library(tibble)
library(purrr)
library(corrplot)
library(ggplot2)


knitr::opts_chunk$set( 
  warning=FALSE,
  message=FALSE,
  results = 'asis',
  error = FALSE,
  tidy = FALSE,
  fig.show = "hold")
```

# DGE across stimulations in GFM 

```{r load tables, echo = TRUE, results = 'hide'}
setwd("~/Documents/MiGASti/Databases")
LPS <- read.table("LPS_GFM_FC1.txt")
ATP <- read.table("ATP_GFM_FC1.txt")
R848 <- read.table("R848_GFM_FC1.txt")
IFNy <- read.table("IFNy_GFM_FC1.txt")
DEX <- read.table("DEX_GFM_FC1.txt")
IL4 <- read.table("IL4_GFM_FC1.txt")

genes_LPS_GFM <- LPS$symbol
genes_ATP_GFM <- ATP$symbol
genes_R848_GFM <- R848$symbol
genes_IFNy_GFM <- IFNy$symbol
genes_DEX_GFM <- DEX$symbol
genes_IL4_GFM <- IL4$symbol

    
# create a vector with all genes listed
x <- c(genes_LPS_GFM, genes_ATP_GFM, genes_R848_GFM, genes_IFNy_GFM, genes_DEX_GFM, genes_IL4_GFM)
# make a dataframe
as.data.frame(x) #1201 genes 
# check if there are genes duplicate 
duplicated(x) 
unique(x)
#remove duplicated
symbol <- x[!duplicated(x)] #1201 genes 

genelist_GFM <- as.data.frame(symbol)

DEG_LPS_GFM <- read.table("DEG_LPS_GFM.txt")
DEG_IFNy_GFM <- read.table("DEG_IFNy_GFM.txt")
DEG_IL4_GFM <- read.table("DEG_IL4_GFM.txt")
DEG_R848_GFM <- read.table("DEG_R848_GFM.txt")
DEG_ATP_GFM <- read.table("DEG_ATP_GFM.txt")
DEG_DEX_GFM <- read.table("DEG_DEX_GFM.txt")
DEG_TNFa_GFM <- read.table("DEG_TNFa_GFM.txt")

DEG_genes_LPS_GFM = merge(DEG_LPS_GFM, genelist_GFM, by ="symbol")
DEG_genes_IFNy_GFM = merge(DEG_IFNy_GFM, genelist_GFM, by = "symbol")
DEG_genes_IL4_GFM = merge(DEG_IL4_GFM, genelist_GFM, by = "symbol")
DEG_genes_R848_GFM = merge(DEG_R848_GFM, genelist_GFM, by = "symbol")
DEG_genes_ATP_GFM = merge(DEG_ATP_GFM, genelist_GFM, by = "symbol")
DEG_genes_DEX_GFM = merge(DEG_DEX_GFM, genelist_GFM, by = "symbol")
DEG_genes_TNFa_GFM = merge(DEG_TNFa_GFM, genelist_GFM, by = "symbol")

all(colnames(DEG_genes_LPS_GFM) == colnames (DEG_genes_IFNy_GFM))#TRUE
Expression_stimulations_GFM <- cbind(DEG_genes_LPS_GFM, DEG_genes_IFNy_GFM, DEG_genes_IL4_GFM, DEG_genes_R848_GFM, DEG_genes_ATP_GFM, DEG_genes_DEX_GFM, DEG_genes_TNFa_GFM)
#write.table(Expression_stimulations_GFM, "Expression_stimulations_GFM.txt")

as.data.frame(Expression_stimulations_GFM) 
# changes names in data frame
names(Expression_stimulations_GFM)[4] <- "LPS"
names(Expression_stimulations_GFM)[12] <- "IFNy"
names(Expression_stimulations_GFM)[20] <- "IL4"
names(Expression_stimulations_GFM)[28] <- "R848"
names(Expression_stimulations_GFM)[36] <- "ATP"
names(Expression_stimulations_GFM)[44] <- "DEX"
names(Expression_stimulations_GFM)[52] <- "TNFa"

DF_log2FC <- Expression_stimulations_GFM %>% select(1, 4, 12, 20, 28, 36, 44, 52)
DF <- as.data.frame(DF_log2FC)
DF <- t(DF)
head( DF )[,1:10]

# fill NA with 0
DF_log2FC[is.na(DF_log2FC)] = 0
# remove the first column
df_num = as.matrix(DF_log2FC[,2:8])
# scale the data 
df_num_scale = scale(df_num)
# make rownames as function
rownames(df_num) = sapply(DF_log2FC$symbol,function(x) 
strsplit(as.character(x),split = "\\\\")[[1]][1])
```

### Heatmap GFM DGE
#1201 genes

```{r heatmap_GFM_cor, echo=TRUE, message=FALSE}
pheatmap(df_num, scale = "row", show_rownames = F, main = "Differentially expressed genes across stimulations in GFM")
```


### Correlation plot GFM DGE (Spearman)
#1201 genes

```{r corrplot_GFM_cor, echo=TRUE, message=FALSE}
com = DF_log2FC[,2:8]
cc = cor(com, method = "spearman")
```




### PCA across stimulations GFM

```{r PCA, echo=TRUE, message=FALSE}
dat <- data.frame(df_num_scale)
cov <- colnames(dat)
as.data.frame(cov)
dat <- data.frame(dat)
pca <- prcomp(t(dat), scale=TRUE)
PoV <- round((pca$sdev^2/sum(pca$sdev^2))*100)
#Grab covariates to plot:
library(ggplot2)
library(ggfortify)


#Plot:
ggplot(pca, aes(PC1, PC2, color=cov)) +
  geom_point(size=2) +
  xlab(paste0("PC1: ",PoV[1],"% variance")) +
  ylab(paste0("PC2: ",PoV[2],"% variance")) +
  theme_classic()

```


# DGE across stimulations in SVZ

```{r heatmap_SVZ, echo = TRUE, results = 'hide'}
setwd("~/Documents/MiGASti/Databases")
LPS <- read.table("LPS_SVZ_FC1.txt")
R848 <- read.table("R848_SVZ_FC1.txt")
IFNy <- read.table("IFNy_SVZ_FC1.txt")
TNFa <- read.table("TNFa_SVZ_FC1.txt")

genes_LPS_SVZ <- LPS$symbol
genes_R848_SVZ <- R848$symbol
genes_IFNy_SVZ <- IFNy$symbol
genes_TNFa_SVZ <- TNFa$symbol

# create a vector with all genes listed
x <- c(genes_LPS_SVZ, genes_R848_SVZ, genes_IFNy_SVZ, genes_TNFa_SVZ)
# make a dataframe
as.data.frame(x) #
# check if there are genes duplicate 
duplicated(x) 
unique(x) 
#remove duplicated
symbol <- x[!duplicated(x)]  

genelist_SVZ <- as.data.frame(symbol)

DEG_LPS_SVZ <- read.table("DEG_LPS_SVZ.txt")
DEG_IFNy_SVZ <- read.table("DEG_IFNy_SVZ.txt")
DEG_R848_SVZ <- read.table("DEG_R848_SVZ.txt")
DEG_ATP_SVZ <- read.table("DEG_ATP_SVZ.txt")
DEG_TNFa_SVZ <- read.table("DEG_TNFa_SVZ.txt")

DEG_genes_LPS_SVZ = merge(DEG_LPS_SVZ, genelist_SVZ, by ="symbol")
DEG_genes_IFNy_SVZ = merge(DEG_IFNy_SVZ, genelist_SVZ, by = "symbol")
DEG_genes_R848_SVZ = merge(DEG_R848_SVZ, genelist_SVZ, by = "symbol")
DEG_genes_ATP_SVZ = merge(DEG_ATP_SVZ, genelist_SVZ, by = "symbol")
DEG_genes_TNFa_SVZ = merge(DEG_TNFa_SVZ, genelist_SVZ, by = "symbol")

all(colnames(DEG_genes_LPS_SVZ) == colnames (DEG_genes_IFNy_SVZ))#TRUE
Expression_stimulations_SVZ <- cbind(DEG_genes_LPS_SVZ, DEG_genes_IFNy_SVZ, DEG_genes_R848_SVZ, DEG_genes_ATP_SVZ, DEG_genes_TNFa_SVZ)
write.table(Expression_stimulations_SVZ, "Expression_stimulations_SVZ.txt")

as.data.frame(Expression_stimulations_SVZ) 
# changes names in data frame
names(Expression_stimulations_SVZ)[4] <- "LPS"
names(Expression_stimulations_SVZ)[12] <- "IFNy"
names(Expression_stimulations_SVZ)[20] <- "R848"
names(Expression_stimulations_SVZ)[28] <- "ATP"
names(Expression_stimulations_SVZ)[36] <- "TNFa"

DF_log2FC <- Expression_stimulations_SVZ %>% select(1, 4, 12, 20, 28, 36)
DF <- as.data.frame(DF_log2FC)
DF <- t(DF)
head( DF )[,1:10]

# fill NA with 0
DF_log2FC[is.na(DF_log2FC)] = 0
# remove the first column
df_num = as.matrix(DF_log2FC[,2:6])
# scale the data 
df_num_scale = scale(df_num)
# make rownames as function
rownames(df_num) = sapply(DF_log2FC$symbol,function(x) 
strsplit(as.character(x),split = "\\\\")[[1]][1])
```
### Heatmap DGE SVZ
#650 genes 

```{r Heatmap_SVZ2, echo=TRUE, message=FALSE}
# heatmap on row scaling
pheatmap(df_num_scale,scale = "row",main = "Differentially expressed genes across stimulations in SVZ")
```

### PCA across stimulations SVZ

```{r PCA_SVZ, echo=TRUE, message=FALSE}
dat <- data.frame(df_num_scale)
cov <- colnames(dat)
as.data.frame(cov)
dat <- data.frame(dat)
pca <- prcomp(t(dat), scale=TRUE)
PoV <- round((pca$sdev^2/sum(pca$sdev^2))*100)
#Grab covariates to plot:
library(ggplot2)
library(ggfortify)


#Plot:
ggplot(pca, aes(PC1, PC2, color=cov)) +
  geom_point(size=2) +
  xlab(paste0("PC1: ",PoV[1],"% variance")) +
  ylab(paste0("PC2: ",PoV[2],"% variance")) +
  theme_classic()

```


# DGE across stimulations GTS

```{r heatmap_GTS, echo = TRUE, results = 'hide'}
setwd("~/Documents/MiGASti/Databases")
LPS <- read.table("LPS_GTS_FC1.txt")
IFNy <- read.table("IFNy_GTS_FC1.txt")

genes_LPS_GTS <- LPS$symbol
genes_IFNy_GTS <- IFNy$symbol

# create a vector with all genes listed
x <- c(genes_LPS_GTS, genes_IFNy_GTS)
# make a dataframe
as.data.frame(x) #
# check if there are genes duplicate 
duplicated(x) 
unique(x) 
#remove duplicated
symbol <- x[!duplicated(x)] # 115 genes 

genelist_GTS <- as.data.frame(symbol)

DEG_LPS_GTS <- read.table("DEG_LPS_GTS.txt")
DEG_IFNy_GTS <- read.table("DEG_IFNy_GTS.txt")
DEG_R848_GTS <- read.table("DEG_R848_GTS.txt")

DEG_genes_LPS_GTS = merge(DEG_LPS_GTS, genelist_GTS, by ="symbol")
DEG_genes_IFNy_GTS = merge(DEG_IFNy_GTS, genelist_GTS, by = "symbol")
DEG_genes_R848_GTS = merge(DEG_R848_GTS, genelist_GTS, by = "symbol")

all(colnames(DEG_genes_LPS_GTS) == colnames (DEG_genes_IFNy_GTS))#TRUE
Expression_stimulations_GTS <- cbind(DEG_genes_LPS_GTS, DEG_genes_IFNy_GTS, DEG_genes_R848_GTS)
write.table(Expression_stimulations_GTS, "Expression_stimulations_GTS.txt")

as.data.frame(Expression_stimulations_GTS) 
# changes names in data frame
names(Expression_stimulations_GTS)[4] <- "LPS"
names(Expression_stimulations_GTS)[12] <- "IFNy"
names(Expression_stimulations_GTS)[20] <- "R848"

DF_log2FC <- Expression_stimulations_GTS %>% select(1, 4, 12, 20)
DF <- as.data.frame(DF_log2FC)
DF <- t(DF)
head( DF )[,1:10]

# fill NA with 0
DF_log2FC[is.na(DF_log2FC)] = 0
# remove the first column
df_num = as.matrix(DF_log2FC[,2:4])
# scale the data 
df_num_scale = scale(df_num)
# make rownames as function
rownames(df_num) = sapply(DF_log2FC$symbol,function(x) 
strsplit(as.character(x),split = "\\\\")[[1]][1])
```
### Heatmap GTS DGE
#115 genes 

```{r Heatmap_GTS2, echo=TRUE, message=FALSE}
# heatmap on row scaling
pheatmap(df_num_scale, scale = 'row', main = "Differentially expressed genes across stimulations in GTS")
```


# DGE across stimulations THA 

```{r heatmap_THA, echo = TRUE, results = 'hide'}
setwd("~/Documents/MiGASti/Databases")
LPS <- read.table("LPS_THA_FC1.txt")
IFNy <- read.table("IFNy_THA_FC1.txt")

genes_LPS_THA <- LPS$symbol
genes_IFNy_THA <- IFNy$symbol

# create a vector with all genes listed
x <- c(genes_LPS_THA, genes_IFNy_THA)
# make a dataframe
as.data.frame(x) #
# check if there are genes duplicate 
duplicated(x) 
unique(x) 
#remove duplicated
symbol <- x[!duplicated(x)] 

genelist_THA <- as.data.frame(symbol)

DEG_LPS_THA <- read.table("DEG_LPS_THA.txt")
DEG_IFNy_THA <- read.table("DEG_IFNy_THA.txt")

DEG_genes_LPS_THA = merge(DEG_LPS_THA, genelist_THA, by ="symbol")
DEG_genes_IFNy_THA = merge(DEG_IFNy_THA, genelist_THA, by = "symbol")

all(colnames(DEG_genes_LPS_THA) == colnames (DEG_genes_IFNy_THA))#TRUE
Expression_stimulations_THA <- cbind(DEG_genes_LPS_THA, DEG_genes_IFNy_THA)
write.table(Expression_stimulations_THA, "Expression_stimulations_THA.txt")

as.data.frame(Expression_stimulations_THA) 
# changes names in data frame
names(Expression_stimulations_THA)[4] <- "LPS"
names(Expression_stimulations_THA)[12] <- "IFNy"

DF_log2FC <- Expression_stimulations_THA %>% select(1, 4, 12)
DF <- as.data.frame(DF_log2FC)
DF <- t(DF)
head( DF )[,1:10]

# fill NA with 0
DF_log2FC[is.na(DF_log2FC)] = 0
# remove the first column
df_num = as.matrix(DF_log2FC[,2:3])
# scale the data 
df_num_scale = scale(df_num)
# make rownames as function
rownames(df_num) = sapply(DF_log2FC$symbol,function(x) 
strsplit(as.character(x),split = "\\\\")[[1]][1])
```
### Heatmap THA DGE
#126 genes 

```{r Heatmap_THA2, echo=TRUE, message=FALSE}
# heatmap on row scaling
pheatmap(df_num_scale, main = "Differentially expressed genes across stimulations in THA")
```

# DGE across stimulations in CC 

```{r heatmap_CC, echo = TRUE, results = 'hide'}
setwd("~/Documents/MiGASti/Databases")
LPS <- read.table("LPS_CC_FC1.txt")
IFNy <- read.table("IFNy_CC_FC1.txt")

genes_LPS_CC <- LPS$symbol
genes_IFNy_CC <- IFNy$symbol

# create a vector with all genes listed
x <- c(genes_LPS_CC, genes_IFNy_CC)
# make a dataframe
as.data.frame(x) #
# check if there are genes duplicate 
duplicated(x) 
unique(x) 
#remove duplicated
symbol <- x[!duplicated(x)] 

genelist_CC <- as.data.frame(symbol)

DEG_LPS_CC <- read.table("DEG_LPS_CC.txt")
DEG_IFNy_CC <- read.table("DEG_IFNy_CC.txt")

DEG_genes_LPS_CC = merge(DEG_LPS_CC, genelist_CC, by ="symbol")
DEG_genes_IFNy_CC = merge(DEG_IFNy_CC, genelist_CC, by = "symbol")

all(colnames(DEG_genes_LPS_CC) == colnames (DEG_genes_IFNy_CC))#TRUE
Expression_stimulations_CC <- cbind(DEG_genes_LPS_CC, DEG_genes_IFNy_CC)
#write.table(Expression_stimulations_CC, "Expression_stimulations_CC.txt")

as.data.frame(Expression_stimulations_CC) 
# changes names in data frame
names(Expression_stimulations_CC)[4] <- "LPS"
names(Expression_stimulations_CC)[12] <- "IFNy"

DF_log2FC <- Expression_stimulations_CC %>% select(1, 4, 12)
DF <- as.data.frame(DF_log2FC)
DF <- t(DF)
head( DF )[,1:10]

# fill NA with 0
DF_log2FC[is.na(DF_log2FC)] = 0
# remove the first column
df_num = as.matrix(DF_log2FC[,2:3])
# scale the data 
df_num_scale = scale(df_num)
# make rownames as function
rownames(df_num) = sapply(DF_log2FC$symbol,function(x) 
strsplit(as.character(x),split = "\\\\")[[1]][1])
```

### Heatmap CC DGE
#120 genes 

```{r Heatmap_CC2, echo=TRUE, message=FALSE}
# heatmap on row scaling
pheatmap(df_num_scale, main = "Differentially expressed genes across stimulations in CC")
```

# DGE LPS across regions

```{r upset_LPS1, echo=TRUE}
setwd("~/Documents/MiGASti/Databases")
list_filters = list(GFM = genes_LPS_GFM, GTS = genes_LPS_GTS, THA = genes_LPS_THA, SVZ = genes_LPS_SVZ, CC = genes_LPS_CC)
#list_filters
LPS_overlap = unlist(list_filters)
```

### Upset plot
 
```{r upset_LPS2, echo=TRUE}
upset(fromList(list_filters), order.by = "freq")
```

# DGE IFNy across regions

```{r upset_IFNy1, echo=TRUE}
setwd("~/Documents/MiGASti/Databases")
list_filters = list(GFM = genes_IFNy_GFM, GTS = genes_IFNy_GTS, THA = genes_IFNy_THA, SVZ = genes_IFNy_SVZ, CC = genes_IFNy_CC)
#list_filters
LPS_overlap = unlist(list_filters)
```

### Upset plot
 
```{r upset_IFNy2, echo=TRUE}
upset(fromList(list_filters), order.by = "freq")
```

