---
title: "Microglia_monocyte_comparisons"
author: "Gijsje"
date: "2/24/2021"
output: 
 rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---

```{r load.packages, echo=FALSE, message=FALSE, results='hide'}
library(readxl)
library(edgeR)
library(limma)
library(DESeq2)
library(tidyverse)
library(data.table)
library(devtools)
library(factoextra)
library(ggfortify)
library(ggplot2)
library(dplyr)
library(broom)
library(variancePartition)
library(pheatmap)
library(edgeR)
library(limma)
library(DESeq2)
library(Hmisc)
library(matrixStats)


knitr::opts_chunk$set( 
  warning=FALSE,
  message=FALSE,
  results = 'asis',
  error = FALSE,
  tidy = FALSE,
  fig.show = "hold")
```

# Load MiGASTi
#metadata + genes_counts

```{r data_load, echo = TRUE}
metadata <- read.table("~/Documents/MiGASti/Databases/metadata.txt")
#set rownames to Sample
row.names(metadata) <- metadata$Sample 
setwd("~/Documents/MiGASti/Databases")
#exclude samples that did not pass QC filtering
exclude <- read.table("samples2remove.txt")
exclude <- exclude$x
#Excludes the samples from filters. 
#dim(genes_counts_filt)
metadata_filt = metadata[ !(rownames(metadata) %in% exclude), ]
length(metadata_filt)

#load gene matrix
load("~/Documents/MiGASti/Databases/gene_matrix.RData")
genes_counts_filt = genes_counts[, !colnames(genes_counts) %in% exclude] 
#Excludes the samples from filters. 

metadata <- subset(metadata_filt, Stimulation == "unstim" | Stimulation == "ununstim")
genes_counts_filt2 <- genes_counts_filt[, colnames(genes_counts_filt) %in% metadata$Sample]
head(rownames(genes_counts_filt2))
genes_counts_migasti <- genes_counts_filt2
metadata_migasti <- metadata
#check ordering
genes_counts_migasti <- genes_counts_migasti[,rownames(metadata_migasti)]
all(rownames(metadata_migasti) == colnames (genes_counts_migasti)) #TRUE
```

### number of samples Migasti cultured 30 hours 

```{r no_migasti_cultured, echo = TRUE}
cultured <-subset(metadata_migasti, Stimulation == "unstim")
length(cultured$Sample)
list(cultured$Sample)
```

### number of samples Migasti baseline (pMG t = 0)

```{r no_migasti_uncultured, echo = TRUE}
uncultured <-subset(metadata_migasti, Stimulation == "ununstim")
length(uncultured$Sample)
list(uncultured$Sample)
```

# Load MiGA dataset 
#genes_counts + metadata

```{r data_loadmiga, echo = TRUE}
load("~/Documents/MiGA/gene_matrix.RData")
load("~/Documents/MiGA/metadata_255filt_eng_29jun2020.Rdata")
genes_counts_miga <- genes_counts
metadata_miga <- metadata3rd_pass
genes_counts_miga <- genes_counts_miga[, colnames(genes_counts_miga) %in% metadata_miga$donor_tissue]
rownames(metadata_miga) <- metadata_miga$donor_tissue
genes_counts_miga <- genes_counts_miga[,rownames(metadata_miga)]
all(rownames(metadata_miga) == colnames (genes_counts_miga)) #TRUE
```

### number of samples Miga baseline (pMG t = 0)

```{r no_miga, echo = TRUE}
length(metadata_miga$donor_tissue)
list(metadata_miga$donor_tissue)
```

# Load monocyte dataset 
#genes_counts + metadata
#only monocytes baseline cultured

```{r data_loadmonocyte, echo = TRUE, results = 'hide'}
load("~/Documents/Monocytes/gene_matrix.RData")
genes_counts_monocytes <- genes_counts
metadata_monocytes <- read.csv("~/Documents/Monocytes/merged_meta_lrrk2_105.csv", header=TRUE)

metadata_monocytes <- subset(metadata_monocytes, treatment == "Basal")
genes_counts_monocytes <- genes_counts_monocytes[, colnames(genes_counts_monocytes) %in% metadata_monocytes$sample_id]
rownames(metadata_monocytes) = metadata_monocytes$sample_id
# check ordering of data 
genes_counts_monocytes <- genes_counts_monocytes[,rownames(metadata_monocytes)]
all(rownames(metadata_monocytes) == colnames (genes_counts_monocytes)) #TRUE
```

### number of samples monocytes cultured 24hours

```{r no_monocytes, echo = TRUE}
length(metadata_monocytes$sample_id)
list(metadata_monocytes$sample_id)
```

# Load macrophage dataset

```{r data_loadmacrophage, echo = TRUE}
load("~/Documents/MiGASti/Databases/macrophage_data.Rdata")
macrophage_data$V2 <- as.numeric(macrophage_data$V2)
macrophage_data$V2.1 <- as.numeric(macrophage_data$V2.1)
macrophage_data$V2.2 <- as.numeric(macrophage_data$V2.2)
macrophage_data$V2.3 <- as.numeric(macrophage_data$V2.3)
macrophage_data$V2.4 <- as.numeric(macrophage_data$V2.4)
macrophage_data$V2.5 <- as.numeric(macrophage_data$V2.5)
macrophage_data$V2.6 <- as.numeric(macrophage_data$V2.6)
macrophage_data$V2.7 <- as.numeric(macrophage_data$V2.7)
```

### number of samples macrophage data 

```{r macrophage_data, echo = TRUE}
data <- macrophage_data[,3:10]
length(data)
colnames(data)
```

# Load ex vivo Gosselin

```{r EV_gos, echo = TRUE}
EV = "~/Documents/MiGASti/Databases/Gosselin_ex_vivo.xlsx"
EV = read_excel(EV, sheet = "Read Counts", col_names = TRUE) 
EV = as.data.frame(EV)
```

### number of samples ex vivo Gosselin data 

```{r no_gosselinev, echo = TRUE}
data <- EV[,2:10]
length(data)
colnames(data)
```

# Load in vivo Gosselin

```{r iv_gos, echo = TRUE, results = 'hide'}
IV = "~/Documents/MiGASti/Databases/Gosselin_invitro.xlsx"
IV = read_excel(IV, sheet = "Read Counts", col_names = TRUE) 
IV = as.data.frame(IV)
```

### number of samples ex vivo Gosselin data 

```{r no_gosseliniv, echo = TRUE}
data <- IV[,2:23]
length(data)
colnames(data)
```


# Load monocytes Gosselin

```{r mono_gos, echo = TRUE}
mono = "~/Documents/MiGASti/Databases/Gosselin_monocytes.xlsx"
mono = read_excel(mono, sheet = "Read Counts", col_names = TRUE) 
mono = as.data.frame(mono)
```

### number of samples monocytes Gosselin data 

```{r no_gosselinmono, echo = TRUE}
data <- mono[,2:11]
length(data)
colnames(data)
```

# Create one dataframe for all samples 

```{r dataframe_one, echo = TRUE, results = 'hide'}
gencode_30 = read.table("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
colnames(gencode_30) = c("ensembl","symbol")
setDT(genes_counts_monocytes, keep.rownames = "ensembl")
genes_counts_monocytes$ensembl
res_name_monocytes = merge(genes_counts_monocytes, gencode_30, by="ensembl")
rownames(res_name_monocytes) = res_name_monocytes$ensembl

setDT(genes_counts_miga, keep.rownames = "ensembl")
genes_counts_miga$ensembl
res_name_miga = merge(genes_counts_miga, gencode_30, by="ensembl")
rownames(res_name_miga) = res_name_miga$ensembl

setDT(genes_counts_migasti, keep.rownames = "ensembl")
genes_counts_migasti$ensembl
res_name_migasti = merge(genes_counts_migasti, gencode_30, by="ensembl")
rownames(res_name_migasti) = res_name_migasti$ensembl

genes_counts_total <- merge(macrophage_data, res_name_monocytes, by = "symbol")
genes_counts_total <- merge(genes_counts_total, res_name_miga, by = "symbol")
genes_counts_total <- merge(genes_counts_total, res_name_migasti, by = "symbol")
genes_counts_total <- merge(genes_counts_total, EV, by = "symbol")
genes_counts_total <- merge(genes_counts_total, IV, by = "symbol")
genes_counts_total <- merge(genes_counts_total, mono, by = "symbol")
genes_counts_total$symbol <- NULL
genes_counts_total$ensembl.x <- NULL
genes_counts_total$ensembl.y <- NULL
genes_counts_total$ensembl.x <- NULL
genes_counts_total$symbol <- NULL
genes_counts_total$ensembl.y <- NULL
genes_counts_total$ensembl.x <- NULL

#21229 genes 

df <- colnames(genes_counts_total)
metadata_total <- as.data.frame(df)
metadata_total[1:8, 1] <- "macrophages_exvivo"
metadata_total[9:43, 1] <- "cultured_monocytes"
metadata_total[44:298, 1] <- "miga_exvivo"
metadata_total[299:433, 1] <- "migasti_invitro"
metadata_total[434:475, 1] <- "migasti_exvivo"
metadata_total[476:499, 1] <- "gosselin_exvivo"
metadata_total[500:521, 1] <- "gosselin_invitro"
metadata_total[522:531, 1] <- "gosselin_monocytes"

colnames(metadata_total)[1] <- "experiment"
metadata_total$experiment <- as.factor(metadata_total$experiment)

#round counts: DESeq can only handle integers
genes_counts_total <- round(genes_counts_total, digits=0)

dds <- DESeqDataSetFromMatrix(countData = genes_counts_total,
                              colData = metadata_total,
                              design= ~ experiment)

vsd <- vst(dds, blind=FALSE)

plotPCA(vsd, intgroup="experiment")
```

### PCA 

```{r plot_pca_1, echo = TRUE}
PCA <- plotPCA(vsd, intgroup="experiment")
plot = as.data.frame(PCA$data)
library(plyr)
cell_type <- revalue(plot$group, c("cultured_monocytes"="macrophages/monocytes", "gosselin_exvivo"="microglia_exvivo", "gosselin_invitro"="microglia_invitro", "gosselin_monocytes"="macrophages/monocytes", "macrophages_exvivo"="macrophages/monocytes", "miga_exvivo"="microglia_exvivo", "migasti_exvivo"="microglia_exvivo", "migasti_invitro"="microglia_invitro"))
ggplot(data = plot, mapping = aes(x = PC1, y = PC2, colour = experiment)) +
  geom_point(aes(shape=cell_type))
```







