---
title: "Microglia_monocyte_comparisons_ranking"
author: "Gijsje"
date: "2/24/2021"
output: 
 rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---

```{r load.packages, echo=FALSE, message=FALSE, results='hide'}
library(readxl)
library(edgeR)
library(limma)
library(DESeq2)
library(tidyverse)
library(data.table)
library(devtools)
library(factoextra)
library(ggfortify)
library(ggplot2)
library(dplyr)
library(broom)
library(variancePartition)
library(pheatmap)
library(edgeR)
library(limma)
library(DESeq2)
library(Hmisc)
library(matrixStats)


knitr::opts_chunk$set( 
  warning=FALSE,
  message=FALSE,
  results = 'asis',
  error = FALSE,
  tidy = FALSE,
  fig.show = "hold")
```

# Load microglia specific genes and monocytes/macrophages dataset based on Patir et al and Haage et al. 2019

```{r data_load1, echo = TRUE}
Patir = "~/Documents/MiGASti/Databases/Patir_2sets.xlsx"
Patir = read_excel(Patir, col_names = TRUE) 
Patir = as.data.frame(Patir)
```

# Load MiGASTi unstim
#metadata + genes_counts

```{r data_load, echo = TRUE, results = 'hide'}
metadata <- read.table("~/Documents/MiGASti/Databases/metadata.txt")
#set rownames to Sample
row.names(metadata) <- metadata$Sample 
setwd("~/Documents/MiGASti/Databases")
#exclude samples that did not pass QC filtering
exclude <- read.table("samples2remove.txt")
exclude <- exclude$x
#Excludes the samples from filters. 
#dim(genes_counts_filt)
metadata_filt = metadata[ !(rownames(metadata) %in% exclude), ]
length(metadata_filt)

#load gene matrix
load("~/Documents/MiGASti/Databases/gene_matrix.RData")
genes_counts_filt = genes_counts[, !colnames(genes_counts) %in% exclude] 
#Excludes the samples from filters. 

metadata <- subset(metadata_filt, Stimulation == "unstim")
genes_counts_filt2 <- genes_counts_filt[, colnames(genes_counts_filt) %in% metadata$Sample]
head(rownames(genes_counts_filt2))
genes_counts_migasti <- genes_counts_filt2
metadata_migasti <- metadata
#check ordering
genes_counts_migasti <- genes_counts_migasti[,rownames(metadata_migasti)]
all(rownames(metadata_migasti) == colnames (genes_counts_migasti)) #TRUE
gencode_30 = read.table("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
colnames(gencode_30) = c("ensembl","symbol")
setDT(genes_counts_migasti, keep.rownames = "ensembl")
genes_counts_migasti$ensembl
res_name = merge(genes_counts_migasti, gencode_30, by="ensembl")
rownames(res_name) = res_name$ensembl
```

### Top 50 genes migasti unstim

```{r data_load_migasti2, echo = TRUE}
res_name2 <- res_name
genes_v <- res_name$ensembl
genes <- as.data.frame(genes_v)
res_name2$ensembl <- NULL
res_name2$symbol <- NULL
sum <- rowMeans(res_name2)
top <- cbind(sum, genes)
expr3 <- top[order(top$sum, decreasing=TRUE),]
colnames(expr3)[2] <- "ensembl"
merged_migasti_unstim <- merge(gencode_30, expr3, by = "ensembl")
merged_migasti_unstim <- merged_migasti_unstim[order(merged_migasti_unstim$sum, decreasing =T),]
merged_migasti_unstim[1:50, ]
migasti_unstim <- merged_migasti_unstim[1:50, ]
```

### Selected genes based on Patir and Haage et al.

```{r data_load_migasti, echo = TRUE}
marker_expression_MiGAsti <- merge(res_name, Patir, by = "symbol")
marker_expression_migasti <- marker_expression_MiGAsti[!duplicated(marker_expression_MiGAsti$symbol), ]
genes_v <- marker_expression_migasti$symbol
marker_expression_migasti$symbol <-NULL
marker_expression_migasti$ensembl <- NULL
marker_expression_cor <- rowMeans(marker_expression_migasti)
expr <- as.data.frame(marker_expression_cor) 
expr2 <- cbind(expr, genes_v)

expr3 <- expr2[order(expr2$marker_expression_cor, decreasing = T),]
expr3$rank <- 1:nrow(expr3) 
data.table(expr3)
```

# Load MiGASTi ununstim

```{r no_migasti_cultured, echo = TRUE, results = 'hide'}
metadata <- read.table("~/Documents/MiGASti/Databases/metadata.txt")
#set rownames to Sample
row.names(metadata) <- metadata$Sample 
setwd("~/Documents/MiGASti/Databases")
#exclude samples that did not pass QC filtering
exclude <- read.table("samples2remove.txt")
exclude <- exclude$x
#Excludes the samples from filters. 
#dim(genes_counts_filt)
metadata_filt = metadata[ !(rownames(metadata) %in% exclude), ]
length(metadata_filt)

#load gene matrix
load("~/Documents/MiGASti/Databases/gene_matrix.RData")
genes_counts_filt = genes_counts[, !colnames(genes_counts) %in% exclude] 
#Excludes the samples from filters. 

metadata <- subset(metadata_filt, Stimulation == "ununstim")
genes_counts_filt2 <- genes_counts_filt[, colnames(genes_counts_filt) %in% metadata$Sample]
head(rownames(genes_counts_filt2))
genes_counts_migasti <- genes_counts_filt2
metadata_migasti <- metadata
#check ordering
genes_counts_migasti <- genes_counts_migasti[,rownames(metadata_migasti)]
all(rownames(metadata_migasti) == colnames (genes_counts_migasti)) #TRUE
gencode_30 = read.table("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
colnames(gencode_30) = c("ensembl","symbol")
setDT(genes_counts_migasti, keep.rownames = "ensembl")
genes_counts_migasti$ensembl
res_name = merge(genes_counts_migasti, gencode_30, by="ensembl")
rownames(res_name) = res_name$ensembl
expr_migasti_unstim <- expr3
```

### Top 50 genes migasti ununstim

```{r data_load_migastiununstim2, echo = TRUE}
res_name2 <- res_name
genes_v <- res_name2$ensembl
genes <- as.data.frame(genes_v)
res_name2$ensembl <- NULL
res_name2$symbol <- NULL
sum <- rowMeans(res_name2)
top <- cbind(sum, genes)
expr3 <- top[order(top$sum, decreasing = TRUE),]
colnames(expr3)[2] <- "ensembl"
merged_migasti_ununstim <- merge(gencode_30, expr3, by = "ensembl")
merged_migasti_ununstim <- merged_migasti_ununstim[order(merged_migasti_ununstim$sum, decreasing =T),]
merged_migasti_ununstim[1:50, ]
migasti_ununstim <- merged_migasti_ununstim[1:50, ]
```

### Selected genes based on Patir and Haage et al.

```{r no_migasti_cultured2, echo = TRUE}
marker_expression_MiGAsti = merge(res_name, Patir, by ="symbol")

marker_expression_migasti <- marker_expression_MiGAsti[!duplicated(marker_expression_MiGAsti$symbol), ]
genes_v <- marker_expression_migasti$symbol
marker_expression_migasti$symbol <-NULL
marker_expression_migasti$ensembl <- NULL
marker_expression_cor <- rowMeans(marker_expression_migasti)
expr <- as.data.frame(marker_expression_cor) 
expr2 <- cbind(expr, genes_v)

expr3 <- expr2[order(expr2$marker_expression_cor, decreasing = TRUE),]
expr3$rank <- 1:nrow(expr3) 
data.table(expr3)
expr_migasti_ununstim <- expr3
```

# Load MiGA dataset 
#genes_counts + metadata

```{r data_loadmiga2, echo = TRUE, results = 'hide'}
load("~/Documents/MiGA/gene_matrix.RData")
load("~/Documents/MiGA/metadata_255filt_eng_29jun2020.Rdata")
genes_counts_miga <- genes_counts
metadata_miga <- metadata3rd_pass
genes_counts_miga <- genes_counts_miga[, colnames(genes_counts_miga) %in% metadata_miga$donor_tissue]
rownames(metadata_miga) <- metadata_miga$donor_tissue
genes_counts_miga <- genes_counts_miga[,rownames(metadata_miga)]
all(rownames(metadata_miga) == colnames (genes_counts_miga)) #TRUE

gencode_30 = read.table("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
colnames(gencode_30) = c("ensembl","symbol")
setDT(genes_counts_miga, keep.rownames = "ensembl")
genes_counts_miga$ensembl
res_name = merge(genes_counts_miga, gencode_30, by="ensembl")
rownames(res_name) = res_name$ensembl
```

### Top 50 genes migasti ununstim

```{r data_load_migastiununstim, echo = TRUE}
res_name2 <- res_name
genes_v <- res_name$ensembl
genes <- as.data.frame(genes_v)
res_name2$ensembl <- NULL
res_name2$symbol <- NULL
sum <- rowMeans(res_name2)
top <- cbind(sum, genes)
expr3 <- top[order(top$sum, decreasing=TRUE),]
colnames(expr3)[2] <- "ensembl"
merged_miga <- merge(gencode_30, expr3, by = "ensembl")
merged_miga <- merged_miga[order(merged_miga$sum, decreasing =T),]
merged_miga[1:50, ]
miga <- merged_miga[1:50, ]
```

### Selected genes based on Patir and Haage et al.

```{r data_loadmiga1, echo = TRUE}
marker_expression_MiGA = merge(res_name, Patir, by ="symbol")

marker_expression_miga <- marker_expression_MiGA[!duplicated(marker_expression_MiGA$symbol), ]
genes_v <- marker_expression_miga$symbol
marker_expression_miga$ensembl <- NULL
marker_expression_miga$symbol <- NULL
marker_expression_cor <- rowMeans(marker_expression_miga)
expr <- as.data.frame(marker_expression_cor) 
expr2 <- cbind(expr, genes_v)
expr3 <- expr2[order(expr2$marker_expression_cor, decreasing = TRUE),]
expr3$rank <- 1:nrow(expr3) 
data.table(expr3)
expr_miga <- expr3
```

# Load cultured monocyte dataset 
#genes_counts + metadata
#only monocytes baseline cultured

```{r data_loadmonocyte1, echo = TRUE, results = 'hide'}
load("~/Documents/Monocytes/gene_matrix.RData")
genes_counts_monocytes <- genes_counts
metadata_monocytes <- read.csv("~/Documents/Monocytes/merged_meta_lrrk2_105.csv", header=TRUE)

metadata_monocytes <- subset(metadata_monocytes, treatment == "Basal")
genes_counts_monocytes <- genes_counts_monocytes[, colnames(genes_counts_monocytes) %in% metadata_monocytes$sample_id]
rownames(metadata_monocytes) = metadata_monocytes$sample_id
# check ordering of data 
genes_counts_monocytes <- genes_counts_monocytes[,rownames(metadata_monocytes)]
all(rownames(metadata_monocytes) == colnames (genes_counts_monocytes)) #TRUE

gencode_30 = read.table("~/Documents/MiGASti/Databases/ens.geneid.gencode.v30")
colnames(gencode_30) = c("ensembl","symbol")
setDT(genes_counts_monocytes, keep.rownames = "ensembl")
genes_counts_monocytes$ensembl
res_name = merge(genes_counts_monocytes, gencode_30, by="ensembl")
rownames(res_name) = res_name$ensembl
marker_expression_monocytes = merge(res_name, Patir, by ="symbol")
```

### Top 50 genes monocytes 

```{r data_load_monocytes_top50, echo = TRUE}
res_name2 <- res_name
genes_v <- res_name$ensembl
genes <- as.data.frame(genes_v)
res_name2$ensembl <- NULL
res_name2$symbol <- NULL
sum <- rowMeans(res_name2)
top <- cbind(sum, genes)
expr3 <- top[order(top$sum, decreasing=TRUE),]
colnames(expr3)[2] <- "ensembl"
merged_monocytes <- merge(gencode_30, expr3, by = "ensembl")
merged_monocytes <- merged_monocytes[order(merged_monocytes$sum, decreasing =T),]
merged_monocytes[1:50, ]
cultured_monocytes <- merged_monocytes[1:50, ]
```

### Selected genes based on Patir and Haage et al.

```{r data_loadmonocyte2, echo = TRUE}
marker_expression_monocytes <- marker_expression_monocytes[!duplicated(marker_expression_monocytes$symbol), ]
genes_v <- marker_expression_monocytes$symbol
marker_expression_monocytes$ensembl <- NULL
marker_expression_monocytes$symbol <- NULL
marker_expression_cor <- rowMeans(marker_expression_monocytes)
expr <- as.data.frame(marker_expression_cor) 
expr2 <- cbind(expr, genes_v)

expr3 <- expr2[order(expr2$marker_expression_cor, decreasing = TRUE),]
expr3$rank <- 1:nrow(expr3) 
data.table(expr3)
```

# Load macrophage dataset

### Top 50 genes macrophages

```{r data_loadmacrophages12, echo = TRUE, results = 'hide'}
load("~/Documents/MiGASti/Databases/macrophage_data.Rdata")
genes_v <- macrophage_data$symbol
genes <- as.data.frame(genes_v)
macrophage_data$ensembl <- NULL
macrophage_data$symbol <- NULL
macrophage_data$V2 <- as.numeric(macrophage_data$V2)
macrophage_data$V2.1 <- as.numeric(macrophage_data$V2.1)
macrophage_data$V2.2 <- as.numeric(macrophage_data$V2.2)
macrophage_data$V2.3 <- as.numeric(macrophage_data$V2.3)
macrophage_data$V2.4 <- as.numeric(macrophage_data$V2.4)
macrophage_data$V2.5 <- as.numeric(macrophage_data$V2.5)
macrophage_data$V2.6 <- as.numeric(macrophage_data$V2.6)
macrophage_data$V2.7 <- as.numeric(macrophage_data$V2.7)

sum <- rowMeans(macrophage_data)
sum2 <- as.data.frame(sum)
top <- cbind(sum, genes)
merged_macrophages <- top[order(top$sum, decreasing = TRUE),]
data.table(merged_macrophages[1:50, ])
macrophages <- merged_macrophages[1:50, ]
colnames(macrophages)[2] <- "symbol"
```

### Selected genes based on Patir and Haage et al.

```{r data_loadmacrophages1, echo = TRUE}
load("~/Documents/MiGASti/Databases/macrophage_data.Rdata")
marker_expression_macrophages = merge(macrophage_data, Patir, by ="symbol")
marker_expression_macrophages <- marker_expression_macrophages[!duplicated(marker_expression_macrophages$symbol), ]
genes_v <- marker_expression_macrophages$symbol
marker_expression_macrophages$ensembl <- NULL
marker_expression_macrophages$symbol <- NULL
marker_expression_macrophages$V2 <- as.numeric(marker_expression_macrophages$V2)
marker_expression_macrophages$V2.1 <- as.numeric(marker_expression_macrophages$V2.1)
marker_expression_macrophages$V2.2 <- as.numeric(marker_expression_macrophages$V2.2)
marker_expression_macrophages$V2.3 <- as.numeric(marker_expression_macrophages$V2.3)
marker_expression_macrophages$V2.4 <- as.numeric(marker_expression_macrophages$V2.4)
marker_expression_macrophages$V2.5 <- as.numeric(marker_expression_macrophages$V2.5)
marker_expression_macrophages$V2.6 <- as.numeric(marker_expression_macrophages$V2.6)
marker_expression_macrophages$V2.7 <- as.numeric(marker_expression_macrophages$V2.7)
marker_expression_cor <- rowMeans(marker_expression_macrophages)
expr <- as.data.frame(marker_expression_cor) 
expr2 <- cbind(expr, genes_v)

expr3 <- expr2[order(expr2$marker_expression_cor, decreasing = TRUE),]
expr3$rank <- 1:nrow(expr3) 
data.table(expr3)
expr_macrophages <- expr3
```

# Load microglia ex vivo Gosselin

### Top 50 genes microglia ex vivo Gosselin

```{r data_load_Gosselin_EV, echo = TRUE}
EV = "~/Documents/MiGASti/Databases/Gosselin_ex_vivo.xlsx"
EV = read_excel(EV, sheet = "Read Counts", col_names = TRUE) 
EV = as.data.frame(EV)
genes_v <- EV$symbol
genes <- as.data.frame(genes_v)
EV$symbol <- NULL

sum <- rowMeans(EV)
sum2 <- as.data.frame(sum)
top <- cbind(sum, genes)
merged_microglia_EV_Gosselin <- top[order(top$sum, decreasing = TRUE),]
data.table(merged_microglia_EV_Gosselin[1:50, ])
microglia_EV_Gosselin <- merged_microglia_EV_Gosselin[1:50, ]
colnames(microglia_EV_Gosselin)[2] <- "symbol"
```

### Selected genes based on Patir and Haage et al.

```{r data_load_Gosselin_EV2, echo = TRUE}
EV = "~/Documents/MiGASti/Databases/Gosselin_ex_vivo.xlsx"
EV = read_excel(EV, sheet = "Read Counts", col_names = TRUE) 
EV = as.data.frame(EV)
marker_expression_microglia_EV_Gosselin = merge(EV, Patir, by ="symbol")
marker_expression_microglia_ev_Gosselin <- marker_expression_microglia_EV_Gosselin[!duplicated(marker_expression_microglia_EV_Gosselin$symbol), ]
genes_v <- marker_expression_microglia_ev_Gosselin$symbol

marker_expression_microglia_ev_Gosselin$symbol <- NULL

marker_expression_cor <- rowMeans(marker_expression_microglia_ev_Gosselin)
expr <- as.data.frame(marker_expression_cor) 
expr2 <- cbind(expr, genes_v)

expr3 <- expr2[order(expr2$marker_expression_cor, decreasing = TRUE),]
expr3$rank <- 1:nrow(expr3) 
data.table(expr3)
```

# Load microglia in vitro Gosselin

### Top 50 genes microglia in vitro Gosselin

```{r data_load_Gosselin_IV, echo = TRUE}
IV = "~/Documents/MiGASti/Databases/Gosselin_invitro.xlsx"
IV = read_excel(IV, sheet = "Read Counts", col_names = TRUE) 
IV = as.data.frame(IV)
genes_v <- IV$symbol
genes <- as.data.frame(genes_v)
IV$symbol <- NULL

sum <- rowMeans(IV)
sum2 <- as.data.frame(sum)
top <- cbind(sum, genes)
merged_microglia_IV_Gosselin <- top[order(top$sum, decreasing = TRUE),]
data.table(merged_microglia_IV_Gosselin[1:50, ])
microglia_IV_Gosselin <- merged_microglia_IV_Gosselin[1:50, ]
colnames(microglia_IV_Gosselin)[2] <- "symbol"
```

### Selected genes based on Patir and Haage et al.

```{r data_load_Gosselin_IV2, echo = TRUE}
IV = "~/Documents/MiGASti/Databases/Gosselin_invitro.xlsx"
IV = read_excel(IV, sheet = "Read Counts", col_names = TRUE) 
IV = as.data.frame(IV)
marker_expression_microglia_IV_Gosselin = merge(IV, Patir, by ="symbol")
marker_expression_microglia_iv_Gosselin <- marker_expression_microglia_IV_Gosselin[!duplicated(marker_expression_microglia_IV_Gosselin$symbol), ]
genes_v <- marker_expression_microglia_iv_Gosselin$symbol

marker_expression_microglia_iv_Gosselin$symbol <- NULL

marker_expression_cor <- rowMeans(marker_expression_microglia_iv_Gosselin)
expr <- as.data.frame(marker_expression_cor) 
expr2 <- cbind(expr, genes_v)

expr3 <- expr2[order(expr2$marker_expression_cor, decreasing = TRUE),]
expr3$rank <- 1:nrow(expr3) 
data.table(expr3)
```


# Load monocytes Gosselin

### Top 50 genes monocytes Gosselin

```{r data_load_Gosselin_mono, echo = TRUE}
mono = "~/Documents/MiGASti/Databases/Gosselin_monocytes.xlsx"
mono = read_excel(mono, sheet = "Read Counts", col_names = TRUE) 
mono = as.data.frame(mono)
genes_v <- mono$symbol
genes <- as.data.frame(genes_v)
mono$symbol <- NULL

sum <- rowMeans(mono)
sum2 <- as.data.frame(sum)
top <- cbind(sum, genes)
merged_microglia_mono_Gosselin <- top[order(top$sum, decreasing = TRUE),]
data.table(merged_microglia_mono_Gosselin[1:50, ])
monocytes_Gosselin <- merged_microglia_mono_Gosselin[1:50, ]
colnames(monocytes_Gosselin)[2] <- "symbol"
```

### Selected genes based on Patir and Haage et al.

```{r data_load_Gosselin_mono2, echo = TRUE}
mono = "~/Documents/MiGASti/Databases/Gosselin_invitro.xlsx"
mono = read_excel(mono, sheet = "Read Counts", col_names = TRUE) 
mono = as.data.frame(mono)
marker_expression_microglia_mono_Gosselin = merge(mono, Patir, by ="symbol")
marker_expression_microglia_mono_Gosselin <- marker_expression_microglia_mono_Gosselin[!duplicated(marker_expression_microglia_mono_Gosselin$symbol), ]
genes_v <- marker_expression_microglia_mono_Gosselin$symbol

marker_expression_microglia_mono_Gosselin$symbol <- NULL

marker_expression_cor <- rowMeans(marker_expression_microglia_mono_Gosselin)
expr <- as.data.frame(marker_expression_cor) 
expr2 <- cbind(expr, genes_v)

expr3 <- expr2[order(expr2$marker_expression_cor, decreasing = TRUE),]
expr3$rank <- 1:nrow(expr3) 
data.table(expr3)
```

# Percentage overlap top 50 genes between migasti unstimulated and other datasets

### + migasti ununstim

```{r data_overlaps, echo = TRUE}
overlap_migasti <- merge(migasti_unstim, migasti_ununstim, by = "symbol")
percentage_migasti = length(overlap_migasti$symbol)/length(migasti_unstim$symbol)*100
percentage_migasti
list(overlap_migasti$symbol)
```

### + miga 

```{r data_overlaps2, echo = TRUE}
overlap_migasti_miga <- merge(migasti_unstim, miga, by = "symbol")
percentage_migasti_miga = length(overlap_migasti_miga$symbol)/length(migasti_unstim$symbol)*100
percentage_migasti_miga
list(overlap_migasti_miga$symbol)
```

### + cultured monocytes

```{r data_overlaps3, echo = TRUE}
overlap_cultured_monocytes <- merge(migasti_unstim, cultured_monocytes, by = "symbol")
percentage_migasti_cultured_monocytes = length(overlap_cultured_monocytes$symbol)/length(migasti_unstim$symbol)*100
percentage_migasti_cultured_monocytes
list(overlap_cultured_monocytes$symbol)
```

### + macrophages

```{r data_overlaps4, echo = TRUE}
overlap_macrophages <- merge(migasti_unstim, macrophages, by = "symbol")
percentage_migasti_macrophages = length(overlap_macrophages$symbol)/length(migasti_unstim$symbol)*100
percentage_migasti_macrophages
list(overlap_macrophages$symbol)
```

### + ex vivo Gosselin

```{r data_overlaps5, echo = TRUE}
overlap_microglia_EV_Gosselin <- merge(migasti_unstim, microglia_EV_Gosselin, by = "symbol")
percentage_migasti_EV_Gosselin = length(overlap_microglia_EV_Gosselin$symbol)/length(migasti_unstim$symbol)*100
percentage_migasti_EV_Gosselin
list(overlap_microglia_EV_Gosselin$symbol)
```

### + in vitro Gosselin 

```{r data_overlaps6, echo = TRUE}
overlap_microglia_IV_Gosselin <- merge(migasti_unstim, microglia_IV_Gosselin, by = "symbol")
percentage_migasti_IV_Gosselin = length(overlap_microglia_IV_Gosselin$symbol)/length(migasti_unstim$symbol)*100
percentage_migasti_IV_Gosselin
list(overlap_microglia_IV_Gosselin$symbol)
```

### + monocytes Gosselin

```{r data_overlaps7, echo = TRUE}
overlap_microglia_mono_Gosselin <- merge(migasti_unstim, microglia_IV_Gosselin, by = "symbol")
percentage_migasti_mono_Gosselin = length(overlap_microglia_mono_Gosselin$symbol)/length(migasti_unstim$symbol)*100
percentage_migasti_mono_Gosselin
list(overlap_microglia_mono_Gosselin$symbol)
```

# percentage between datasets

```{r data_overlaps8, echo = TRUE}
percentage <- rbind(percentage_migasti, percentage_migasti_cultured_monocytes, percentage_migasti_macrophages, percentage_migasti_EV_Gosselin, percentage_migasti_IV_Gosselin,percentage_migasti_mono_Gosselin) 

datasets <- rownames(percentage)

percentages <- cbind(datasets, percentage)

p<-ggplot(data=percentages, aes(x=percentage, y=datasets)) +
geom_bar(stat="identity")
p
```
