---
title: "DTU"
author: "Gijsje"
date: "2/16/2021"
output: 
 rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---

```{r setup1, include=FALSE}
load("~/Documents/MiGASti/Databases/splicing_dtu_stimulation_microglia.Rdata")

library(tidyverse)
library(satuRn)
library(SummarizedExperiment)
library(edgeR)
library(readr)
library(data.table)
library(tidyverse)
library(SummarizedExperiment)
library(edgeR)
library(readr)
library(data.table)
library(tidyverse)
library(UpSetR)

library(RColorBrewer)
library(ComplexHeatmap)
```

```{r Helper Functions, echo=FALSE}
createDT <- function(DF, caption="", scrollY=500){
data <- DT::datatable(DF, caption=caption,
extensions = 'Buttons',
options = list( dom = 'Bfrtip',
buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
scrollY = scrollY, scrollX=T, scrollCollapse = T, paging = F,
columnDefs = list(list(className = 'dt-center', targets = "_all"))
)
)
return(data)
}
```

# Summary of experiments DTU
#code for DTU analysis can be found in docs DTU.Rmd. Only follow up analyses are showed in html.

### empirical FDR < 0.05

```{r empirical, echo = TRUE}
rm(sumExp)
# summarise

# empirical FDR < 0.05
map_df(all_res, ~{group_by(.x, FDR = empirical_FDR < 0.05) %>% tally() %>% filter(FDR == TRUE)  }, .id = "contrast")
```

### regular FDR < 0.05 

```{r regular, echo = TRUE}
# regular FDR < 0.05
map_df(all_res, ~{group_by(.x, FDR = regular_FDR < 0.05) %>% tally() %>% filter(FDR == TRUE) }, .id = "contrast")
```

### Upsetplot DTU across conditions 

```{r summary, echo = TRUE}
dtu_genes <- map(all_res, ~{ .x %>% filter(empirical_FDR < 0.05) %>% pull(gene) })

#UpSetR::upset(fromList(dtu_genes),nintersects = NA)
UpSetR::upset(fromList(dtu_genes),nintersects = NA,nsets = 7)

sig_res <- map_df(all_res, ~{.x %>% filter(empirical_FDR < 0.05, abs(estimates) > 1)}, .id = "contrast")

age_res <- filter(all_res$age, empirical_FDR < 0.05)

stimulation_res <- map_df(all_res, ~{.x %>% filter(empirical_FDR < 0.05, abs(estimates) > 1)}, .id = "contrast" ) %>% filter(contrast != "age")
```

### Volcano plots stimulation

```{r volcano_all, echo = TRUE}
all_res %>%
  map_df( ~{.x %>% filter(pval < 0.05)}, .id = "contrast") %>%
  filter(!contrast %in% "age") %>%
  ggplot(aes( x = estimates, y = -log10(pval), colour = empirical_FDR < 0.05 )) +
  geom_point(size = 1) + scale_colour_manual(values = c("black", "red")) + 
  facet_wrap(~contrast, scales = "free") +
  xlim(-4,4) +
  theme_bw()
```

### Volcano plot age

```{r volcano_age, echo = TRUE}
# age only
all_res %>%
  map_df( ~{.x %>% filter(pval < 0.05)}, .id = "contrast") %>%
  filter(contrast == "age") %>%
  ggplot(aes( x = estimates, y = -log10(pval), colour = empirical_FDR < 0.5 )) +
  geom_point(size = 1) + scale_colour_manual(values = c("black", "red")) + 
  facet_wrap(~contrast, scales = "free") +
  xlim(-0.1,0.1) +
  theme_bw()
```


# LPS 
### empirical FDR < 0.05 

```{r LPS5, echo=TRUE}
length(which(all_res$LPS$empirical_FDR < 0.05))
```

### empirical FDR < 0.10 

```{r LPS10, echo=TRUE}
length(which(all_res$LPS$empirical_FDR < 0.10))
```

### empirical FDR < 0.15 

```{r LPS15, echo=TRUE}
length(which(all_res$LPS$empirical_FDR < 0.15))
```

### Table for download empirical FDR < 0.05

```{r transcriptsLPS, echo=TRUE}
LPS_t <- subset(all_res$LPS, empirical_FDR < 0.05)
res_name_LPS_top = LPS_t[, c("isoform_id", "gene_id", "gene", "estimates", "se", "df", "pval", "regular_FDR", "empirical_pval", "empirical_FDR")]
createDT(res_name_LPS_top)
```

# IFNy 
### empirical FDR < 0.05 

```{r IFNy5, echo = TRUE}
length(which(all_res$IFNy$empirical_FDR < 0.05))
```

### empirical FDR < 0.10

```{r IFNy1-, echo = TRUE}
length(which(all_res$IFNy$empirical_FDR < 0.10))
```

### empirical FDR < 0.15

```{r IFNy15, echo = TRUE}
length(which(all_res$IFNy$empirical_FDR < 0.15))
```

### Table for download empirical FDR < 0.05

```{r transcriptsIFNy, echo=TRUE}
IFNy_t <- subset(all_res$IFNy, empirical_FDR < 0.05)
res_name_IFNy_top = IFNy_t[, c("isoform_id", "gene_id", "gene", "estimates", "se", "df", "pval", "regular_FDR", "empirical_pval", "empirical_FDR")]
createDT(res_name_IFNy_top)
```

# R848 
###  empirical FDR < 0.05 

```{r R848, echo = TRUE}
length(which(all_res$R848$empirical_FDR < 0.05))
```

###  empirical FDR < 0.10 

```{r R848_10, echo=TRUE}
length(which(all_res$R848$empirical_FDR < 0.10))
```

###  empirical FDR < 0.15 

```{r R848_15, echo=TRUE}
length(which(all_res$R848$empirical_FDR < 0.15))
```

### Table for download empirical FDR < 0.05

```{r transcriptsR848, echo=TRUE}
R848_t <- subset(all_res$R848, empirical_FDR < 0.05)
res_name_R848_top = R848_t[, c("isoform_id", "gene_id", "gene", "estimates", "se", "df", "pval", "regular_FDR", "empirical_pval", "empirical_FDR")]
createDT(res_name_R848_top)
```

# TNFa 
### empirical FDR < 0.05 

```{r TNFa5, echo = TRUE}
length(which(all_res$TNFa$empirical_FDR < 0.05))
```

### empirical FDR < 0.10 

```{r TNFa10, echo = TRUE}
length(which(all_res$TNFa$empirical_FDR < 0.10))
```

### empirical FDR < 0.15 

```{r TNFa15, echo = TRUE}
length(which(all_res$TNFa$empirical_FDR < 0.15))
```

### Table for download empirical FDR < 0.05

```{r transcriptsTNFa, echo=TRUE}
TNFa_t <- subset(all_res$TNFa, empirical_FDR < 0.05)
res_name_TNFa_top = TNFa_t[, c("isoform_id", "gene_id", "gene", "estimates", "se", "df", "pval", "regular_FDR", "empirical_pval", "empirical_FDR")]
createDT(res_name_TNFa_top)
```

# DEX
### empirical FDR < 0.05 

```{r DEX5, echo=TRUE}
length(which(all_res$DEX$empirical_FDR < 0.05))
```

### empirical FDR < 0.10

```{r DEX10, echo=TRUE}
length(which(all_res$DEX$empirical_FDR < 0.10))
```

### empirical FDR < 0.15

```{r DEX15, echo=TRUE}
length(which(all_res$DEX$empirical_FDR < 0.15))
```

### Table for download empirical FDR < 0.05

```{r transcriptsDEX, echo=TRUE}
DEX_t <- subset(all_res$DEX, empirical_FDR < 0.05)
res_name_DEX_top = DEX_t[, c("isoform_id", "gene_id", "gene", "estimates", "se", "df", "pval", "regular_FDR", "empirical_pval", "empirical_FDR")]
createDT(res_name_DEX_top)
```

# IL4 
### empirical FDR < 0.05 

```{r IL4_5, echo=TRUE}
length(which(all_res$IL4$empirical_FDR < 0.05))
```

### empirical FDR < 0.10

```{r IL4_10, echo=TRUE}
length(which(all_res$IL4$empirical_FDR < 0.10))
```

### empirical FDR < 0.15

```{r IL4_15, echo=TRUE}
length(which(all_res$IL4$empirical_FDR < 0.15))
```

### Table for download empirical FDR < 0.05

```{r transcriptsIL4, echo=TRUE}
IL4_t <- subset(all_res$IL4, empirical_FDR < 0.05)
res_name_IL4_top = IL4_t[, c("isoform_id", "gene_id", "gene", "estimates", "se", "df", "pval", "regular_FDR", "empirical_pval", "empirical_FDR")]
createDT(res_name_IL4_top)
```

# ATP 
### empirical FDR < 0.05 

```{r ATP, echo=TRUE}
length(which(all_res$ATP$empirical_FDR < 0.05))
```

### empirical FDR < 0.10 

```{r ATP_10, echo=TRUE}
length(which(all_res$ATP$empirical_FDR < 0.10))
```

### empirical FDR < 0.15 

```{r ATP15, echo = TRUE}
length(which(all_res$ATP$empirical_FDR < 0.15))
```

### Table for download empirical FDR < 0.05

```{r transcriptsATP, echo=TRUE}
ATP_t <- subset(all_res$ATP, empirical_FDR < 0.05)
res_name_ATP_top = ATP_t[, c("isoform_id", "gene_id", "gene", "estimates", "se", "df", "pval", "regular_FDR", "empirical_pval", "empirical_FDR")]
createDT(res_name_ATP_top)
```

# Age 
### empirical FDR < 0.05 

```{r age, echo = TRUE}
length(which(all_res$age$empirical_FDR < 0.05))
```

### empirical FDR < 0.10 

```{r age10, echo = TRUE}
length(which(all_res$age$empirical_FDR < 0.10))
```

### empirical FDR < 0.15 

```{r age15, echo = TRUE}
length(which(all_res$age$empirical_FDR < 0.15))
```

### Table for download empirical FDR < 0.05

```{r transcriptsage, echo=TRUE}
age_t <- subset(all_res$age, empirical_FDR < 0.05)
res_name_age_top = age_t[, c("isoform_id", "gene_id", "gene", "estimates", "se", "df", "pval", "regular_FDR", "empirical_pval", "empirical_FDR")]
createDT(res_name_age_top)
```