---
title: "Microglia_monocyte_comparisons"
author: "Gijsje"
date: "2/24/2021"
output: 
 rmarkdown::html_document:
   theme: united
   highlight: tango
   code_folding: hide
   toc: true
   toc_float: true
   df_print: paged
   smooth_scroll: true
   number_sections: false
   self_contained: true
---

```{r load.packages, echo=FALSE, message=FALSE, results='hide'}
library(readxl)
library(edgeR)
library(limma)
library(DESeq2)
library(tidyverse)
library(data.table)
library(devtools)
library(factoextra)
library(ggfortify)
library(ggplot2)
library(dplyr)
library(broom)
library(variancePartition)
library(pheatmap)
library(edgeR)
library(limma)
library(DESeq2)
library(Hmisc)
library(matrixStats)


knitr::opts_chunk$set( 
  warning=FALSE,
  message=FALSE,
  results = 'asis',
  error = FALSE,
  tidy = FALSE,
  fig.show = "hold")
```

# Load MiGASTi
#metadata + genes_counts

```{r data_load, echo = TRUE}
metadata <- read.table("~/Documents/MiGASti/Databases/metadata.txt")
#set rownames to Sample
row.names(metadata) <- metadata$Sample 
setwd("~/Documents/MiGASti/Databases")
#exclude samples that did not pass QC filtering
exclude <- read.table("samples2remove.txt")
exclude <- exclude$x
#Excludes the samples from filters. 
#dim(genes_counts_filt)
metadata_filt = metadata[ !(rownames(metadata) %in% exclude), ]
length(metadata_filt)

#load gene matrix
load("~/Documents/MiGASti/Databases/gene_matrix.RData")
genes_counts_filt = genes_counts[, !colnames(genes_counts) %in% exclude] 
#Excludes the samples from filters. 

metadata <- subset(metadata_filt, Stimulation == "unstim" | Stimulation == "ununstim")
genes_counts_filt2 <- genes_counts_filt[, colnames(genes_counts_filt) %in% metadata$Sample]
head(rownames(genes_counts_filt2))
genes_counts_migasti <- genes_counts_filt2
metadata_migasti <- metadata
#check ordering
genes_counts_migasti <- genes_counts_migasti[,rownames(metadata_migasti)]
all(rownames(metadata_migasti) == colnames (genes_counts_migasti)) #TRUE
```

### number of samples Migasti cultured 30 hours 

```{r no_migasti_cultured, echo = TRUE}
cultured <-subset(metadata_migasti, Stimulation == "unstim")
length(cultured$Sample)
list(cultured$Sample)
```

### number of samples Migasti baseline (pMG t = 0)

```{r no_migasti_uncultured, echo = TRUE}
uncultured <-subset(metadata_migasti, Stimulation == "ununstim")
length(uncultured$Sample)
list(uncultured$Sample)
```

# Load MiGA dataset 
#genes_counts + metadata

```{r data_loadmiga, echo = TRUE}
load("~/Documents/MiGA/gene_matrix.RData")
load("~/Documents/MiGA/metadata_255filt_eng_29jun2020.Rdata")
genes_counts_miga <- genes_counts
metadata_miga <- metadata3rd_pass
genes_counts_miga <- genes_counts_miga[, colnames(genes_counts_miga) %in% metadata_miga$donor_tissue]
rownames(metadata_miga) <- metadata_miga$donor_tissue
genes_counts_miga <- genes_counts_miga[,rownames(metadata_miga)]
all(rownames(metadata_miga) == colnames (genes_counts_miga)) #TRUE
```

### number of samples Miga baseline (pMG t = 0)

```{r no_miga, echo = TRUE}
length(metadata_miga$donor_tissue)
list(metadata_miga$donor_tissue)
```

# Load monocyte dataset 
#genes_counts + metadata
#only monocytes baseline cultured

```{r data_loadmonocyte, echo = TRUE}
load("~/Documents/Monocytes/gene_matrix.RData")
genes_counts_monocytes <- genes_counts
metadata_monocytes <- read.csv("~/Documents/Monocytes/merged_meta_lrrk2_105.csv", header=TRUE)

metadata_monocytes <- subset(metadata_monocytes, treatment == "Basal")
genes_counts_monocytes <- genes_counts_monocytes[, colnames(genes_counts_monocytes) %in% metadata_monocytes$sample_id]
rownames(metadata_monocytes) = metadata_monocytes$sample_id
# check ordering of data 
genes_counts_monocytes <- genes_counts_monocytes[,rownames(metadata_monocytes)]
all(rownames(metadata_monocytes) == colnames (genes_counts_monocytes)) #TRUE
```

### number of samples monocytes cultured 24hours

```{r no_monocytes, echo = TRUE}
length(metadata_monocytes$sample_id)
list(metadata_monocytes$sample_id)
```

# Create one dataframe of all datasets 

```{r dataframe, echo = TRUE, results = 'hide'}
all(rownames(genes_counts_monocytes) == rownames(genes_counts_miga)) #TRUE
all(rownames(genes_counts_miga) == rownames(genes_counts_migasti)) #TRUE
all(rownames(genes_counts_monocytes) == rownames(genes_counts_migasti)) #TRUE

#create frame for metadata 
metadata_miga_v <-select(metadata_miga, "donor_tissue")
metadata_migasti_v <-select(metadata_migasti, "Sample")
metadata_monocytes_v <- select(metadata_monocytes, "sample_id")

names(metadata_miga_v)[names(metadata_miga_v) == "donor_tissue"] <- "Sample"
names(metadata_monocytes_v)[names(metadata_monocytes_v) == "sample_id"] <- "Sample"

#create groups with type of experiment 
experiment <- c("cultured_monocytes_24hours")
metadata_monocytes_v$experiment <- experiment
experiment <- c("baseline_pMG_uncultured_miga")
metadata_miga_v$experiment <- experiment
experiment <- c("migasti")
metadata_migasti_v$experiment <- experiment

#rename cultured and uncultured samples in migasti dataset
metadata_migasti_v$experiment <- experiment
t <- grepl('RNA', row.names(metadata_migasti_v))
metadata_migasti_v[t,]$experiment <- 'baseline_pMG_uncultured_migasti'
t <- grepl('unstim', row.names(metadata_migasti_v))
metadata_migasti_v[t,]$experiment <- 'migasti_cultured_30_hours'

identical(rownames(metadata_miga_v), colnames(genes_counts_miga))
identical(rownames(metadata_migasti_v), colnames(genes_counts_migasti))
identical(rownames(metadata_monocytes_v), colnames(genes_counts_monocytes))


genes_counts_total <- cbind(genes_counts_migasti, genes_counts_miga, genes_counts_monocytes)

metadata_total <- rbind(metadata_migasti_v, metadata_miga_v, metadata_monocytes_v)
identical(rownames(metadata_total), colnames(genes_counts_total))

genes_counts_total <- genes_counts_total[,rownames(metadata_total)]
all(rownames(metadata_total) == colnames (genes_counts_total)) #TRUE

#round counts: DESeq can only handle integers
genes_counts_total <- round(genes_counts_total, digits=0)

dds <- DESeqDataSetFromMatrix(countData = genes_counts_total,
                              colData = metadata_total,
                              design= ~ experiment)

vsd <- vst(dds, blind=FALSE)
```

### PCA of MiGASti, MiGA and cultured monocytes

```{r plot_pca_1, echo = TRUE}
PCA <- plotPCA(vsd, intgroup="experiment")
plot = as.data.frame(PCA$data)
library(plyr)
cell_type <- revalue(plot$group, c("cultured_monocytes_24hours"="monocytes", "baseline_pMG_uncultured_miga"="microglia", "baseline_pMG_uncultured_migasti"="microglia", "migasti_cultured_30_hours"="microglia"))
ggplot(data = plot, mapping = aes(x = PC1, y = PC2, colour = experiment)) +
  geom_point(aes(shape=cell_type))
```


# Include stimulation data of LPS and IFNy for monocytes and microglia

```{r load_miga2.0, echo = TRUE}
metadata <- subset(metadata_filt, Stimulation == "unstim" | Stimulation == "ununstim"| Stimulation == "LPS" | Stimulation == "IFNy")
genes_counts_filt2 <- genes_counts_filt[, colnames(genes_counts_filt) %in% metadata$Sample]
head(rownames(genes_counts_filt2))
genes_counts_migasti <- genes_counts_filt2
metadata_migasti <- metadata
#check ordering
genes_counts_migasti <- genes_counts_migasti[,rownames(metadata_migasti)]
all(rownames(metadata_migasti) == colnames (genes_counts_migasti)) #TRUE
```

### number of samples MiGASti LPS 6 hours 

```{r migasti2, echo = TRUE}
cultured <-subset(metadata_migasti, Stimulation == "LPS")
length(cultured$Sample)
list(cultured$Sample)
```


### number of samples MiGASti IFNy 6 hours 

```{r no_m, echo = TRUE}
cultured <-subset(metadata_migasti, Stimulation == "IFNy")
length(cultured$Sample)
list(cultured$Sample)
```

# Include stimulation monocyte 24 hours LPS and IFNy

```{r monocytes_lpsIFNy, echo = TRUE}
load("~/Documents/Monocytes/gene_matrix.RData")
genes_counts_monocytes <- genes_counts
metadata_monocytes <- read.csv("~/Documents/Monocytes/merged_meta_lrrk2_105.csv", header=TRUE)
rownames(metadata_monocytes) = metadata_monocytes$sample_id
# check ordering of data
all(rownames(metadata_monocytes) == colnames (genes_counts_monocytes)) #TRUE
```

### number of samples monocytes LPS 24 hours

```{r no_monocytes_extra, echo = TRUE}
cultured <-subset(metadata_monocytes, treatment == "LPS")
length(cultured$sample_id)
list(cultured$sample_id)
```

### number of samples monocytes IFNy 24 hours

```{r no_monocytesIFNy, echo = TRUE}
cultured <-subset(metadata_monocytes, treatment == "IFNg")
length(cultured$sample_id)
list(cultured$sample_id)
```

# Create one dataframe of all datasets including LPS and IFNy 

```{r dataframe_all, echo = TRUE, results = 'hide'}
all(rownames(genes_counts_monocytes) == rownames(genes_counts_miga)) #TRUE
all(rownames(genes_counts_miga) == rownames(genes_counts_migasti)) #TRUE
all(rownames(genes_counts_monocytes) == rownames(genes_counts_migasti)) #TRUE

#create frame for metadata 
metadata_miga_v <-select(metadata_miga, "donor_tissue")
metadata_migasti_v <-select(metadata_migasti, "Sample")
metadata_monocytes_v <- select(metadata_monocytes, "sample_id")

names(metadata_miga_v)[names(metadata_miga_v) == "donor_tissue"] <- "Sample"
names(metadata_monocytes_v)[names(metadata_monocytes_v) == "sample_id"] <- "Sample"

#create groups with type of experiment 
experiment <- c("cultured_monocytes_24hours")
metadata_monocytes_v$experiment <- experiment
experiment <- c("baseline_pMG_MiGA")
metadata_miga_v$experiment <- experiment
experiment <- c("MiGASti")
metadata_migasti_v$experiment <- experiment

#create groups with microglia/monocytes
group <- c("monocytes")
metadata_monocytes_v$group <- group
group <- c("microglia")
metadata_miga_v$group <- group
group <- c("microglia")
metadata_migasti_v$group <- group 

#rename cultured, LPS and IFNy samples in monocytes dataset
t <- grepl('B', row.names(metadata_monocytes_v))
metadata_monocytes_v[t,]$experiment <- 'cultured_monocytes_24hours'

t <- grepl('LPS', row.names(metadata_monocytes_v))
metadata_monocytes_v[t,]$experiment <- 'LPS_monocytes_24hours'

t <- grepl('IFNg', row.names(metadata_monocytes_v))
metadata_monocytes_v[t,]$experiment <- 'IFNy_monocytes_24h'

#rename cultured, LPS, IFNy and baseline in MiGASti dataset

t <- grepl('RNA', row.names(metadata_migasti_v))
metadata_migasti_v[t,]$experiment <- 'baseline_pMG_uncultured_migasti'
t <- grepl('unstim', row.names(metadata_migasti_v))
metadata_migasti_v[t,]$experiment <- 'cultured_microglia_30_hours'
t <- grepl('IFNy', row.names(metadata_migasti_v))
metadata_migasti_v[t,]$experiment <- 'IFNy_microglia_6h'
t <- grepl('LPS', row.names(metadata_migasti_v))
metadata_migasti_v[t,]$experiment <- 'LPS_microglia_6h'

identical(rownames(metadata_miga_v), colnames(genes_counts_miga))
identical(rownames(metadata_migasti_v), colnames(genes_counts_migasti))
identical(rownames(metadata_monocytes_v), colnames(genes_counts_monocytes))


genes_counts_total <- cbind(genes_counts_migasti, genes_counts_miga, genes_counts_monocytes)

metadata_total <- rbind(metadata_migasti_v, metadata_miga_v, metadata_monocytes_v)
identical(rownames(metadata_total), colnames(genes_counts_total))

genes_counts_total <- genes_counts_total[,rownames(metadata_total)]
all(rownames(metadata_total) == colnames (genes_counts_total)) #TRUE

#round counts: DESeq can only handle integers
genes_counts_total <- round(genes_counts_total, digits=0)

dds <- DESeqDataSetFromMatrix(countData = genes_counts_total,
                              colData = metadata_total,
                              design= ~ experiment)

vsd <- vst(dds, blind=FALSE)
```
### PCA of MiGASti, MiGA and cultured monocytes with LPS and IFNy included


```{r plot_pca_all, echo = TRUE}
plotPCA(vsd, intgroup="experiment")
PCA <- plotPCA(vsd, intgroup="experiment")
plot = as.data.frame(PCA$data)
library(plyr)
cell_type <- revalue(plot$group, c("cultured_monocytes_24hours"="monocytes", "baseline_pMG_uncultured_miga"="microglia", "baseline_pMG_uncultured_migasti"="microglia", "migasti_cultured_30_hours"="microglia", "IFNy_microglia_6h"="microglia", "LPS_microglia_6h"="microglia", "LPS_monocytes_24hours"="monocytes", "IFNy_monocytes_24h"="monocytes","cultured_microglia_30_hours"="microglia","baseline_pMG_MiGA"="microglia" ))
ggplot(data = plot, mapping = aes(x = PC1, y = PC2, colour = experiment)) +
  geom_point(aes(shape=cell_type))
```


